### 2019-09-02 icra.tex
Add icra paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

%\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper

\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.

%In case you encounter the following error:
%Error 1010 The PDF file may be corrupt (unable to open PDF file) OR
%Error 1000 An error occurred while parsing a contents stream. Unable to analyze the PDF file.
%This is a known problem with pdfLaTeX conversion filter. The file cannot be opened with acrobat reader
%Please use one of the alternatives below to circumvent this error by uncommenting one or the other
%\pdfobjcompresslevel=0
%\pdfminorversion=4

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}
\usepackage{graphicx}  % assumes amsmath package installed
\usepackage{enumitem}
\usepackage{tikz}
\usepackage{amsfonts}
\usepackage{multirow}
\usepackage{float}
\usetikzlibrary{arrows,automata}
\usepackage{pgfplots}
\usetikzlibrary{intersections}

%\usepackage[utf8]{inputenc}

\title{\LARGE \bf
Multi-Modal Simultaneous Forecasting\\%
of Vehicle Position Sequences using Social Attention
}

\author{Jean Mercat$^{1, 2}$, Nicole El Zoghby$^{1}$, Guillaume Sandou$^{2}$, Thomas Gilles$^{1, 3}$,\\%
Dominique Beauvois$^{2}$, and Guillermo Pita Gil$^{1}$% <-this % stops a space
%\thanks{*This work was made in collaboration with L2S and Renault SAS}% <-this % stops a space
\thanks{$^{1}$ Department of data fusion,
        Technocentre Renault, 78280 Guyancourt, France
        {\tt\Small \{Jean.Mercat, Thomas.Gilles, Nicole.El-Zoghby, Guillermo.Pita-Gil\}@renault.com}}%
\thanks{$^{2}$ Laboratoire des signaux et des syst\`emes,
        Centrale-Sup\'elec, 91192 Gif sur Yvette, France
        {\tt\Small \{Jean.Mercat, Guillaume.Sandou, Dominique.Beauvois\}@centralesupelec.fr}}%
\thanks{$^{3}$ \'Ecole polytechnique,
        Route de Saclay, 91128 Palaiseau
        {\tt\Small Thomas.Gilles@polytechnique.edu}}%
}

\begin{document}

\tikzset{
    state/.style={
           rectangle,
           draw=black, very thick,
           minimum height=2em,
           inner sep=2pt,
           text centered,
           },
    name plot/.style={every path/.style={name path global=#1}}
}

\pgfmathdeclarefunction{dnorm}{2}{%
  \pgfmathparse{1/(#2*sqrt(2*pi))*exp(-((x-#1)^2)/(2*#2^2))}%
}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

Vehicle trajectory forecasting models use a wide variety of frameworks for interaction and multi-modality.
They rely on various representations of the road scene and definitions of maneuvers.
In this paper we present a simple model that
simultaneously forecasts each vehicle position on a road scene as a sequence
of multi-modal probability density functions.
This relies solely on vehicle position tracks and does not define maneuvers.
We produce an easily extendable model that combines these predictive capabilities while surpassing state-of-the-art
results.
Its architecture uses multi-head attention to account for complete interactions between all vehicles,
and long short-term memory (LSTM) layers for encoding and forecasting.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}

Automation of driving tasks aims for safety and comfort improvements.
For that purpose, most Autonomous Driving (AD) system relies on the anticipation of the
traffic scene movements.
%Here we present a forecasting model that produce multimodal stochastic predictions for all vehicles of a partially
%observed road scene.
The AD system uses sensors for perception to produce an interpreted representation of its surroundings.
For this work the surroundings are composed of all tracks of observed vehicle positions for the past
few seconds.
Using this simple representation without constraint on the input data, the produced output is a multimodal stochastic
forecast for each vehicle future positions.

In most forecasting applications, only the ego vehicle trajectory is predicted.
However, avoiding collisions can only be made with forecasts for all participants.
In some applications, to predict the whole scene, each vehicle trajectory can be predicted separately with the same
model.
However, this means that only one-way interactions (others to predicted vehicle) are modeled.
Therefore, simultaneous forecast of all perceived vehicles takes better accounts of interactions.

In a scene observed from an ego vehicle, the quality of the perceived information about the
surroundings of each vehicle is uneven.
The model forecasting a vehicle trajectory at the boundary of the field of view should not assume
that there is nothing where nothing is seen.

Some of the information that determines the future trajectories of the observed vehicle cannot be observed.
Therefore, the predictions are inherently uncertain.
We classify the uncertainties in two kinds: small variations around a mode and modes.
The first kind are the continuous uncertainties that are present at each step of the process such as perception,
estimation, and some model approximations.
The second kind are discrete local maxima of probability density.
They stand for occurrences of choices, for example, a driver chooses a lane,
or the perception system chooses a classification.
The forecasts represent both source of uncertainties with
density probability functions in the form of Gaussian mixtures.

This works combines stochastic multimodal forecasting with simultaneous forecasts of the whole scene
without assuming anything in unobserved areas.
This produces a model that simulates two-way interactions in time between all participants
while accounting for uncertainties from modes, and small variations.
It relies on a very simple representation of the environment that is easily extendable.

\section{Related work}

\textbf{The evolution} of road scene trajectory forecasting methods have progressively switched from kinematic, physics based models to learned
statistical frameworks.
The Kalman filters~\cite{Kalman1960} with various physical model is still widely used for short-term motion
forecast (less than 1 second) in the automotive industry.
However, to forecast with longer time horizon (about 5 seconds), statistical models have been used.
The survey~\cite{Lefevre2014} compares many statistical models for trajectory and maneuver forecast
with the most widely used models at that time being Hidden Markov Models (HMM) and Support Vector Machines (SVM).
Since then, recurrent neural networks (RNN) mostly using the Long Short-Term Memory~\cite{Hochreiter1997} (LSTM)
architecture have become the standard technology for statistical trajectory forecasting.
Its model free versatility has allowed the models to account for interactions between vehicles.
It has been used as a maneuver classifier in~\cite{Khosroshahi2016} and as a trajectory predictor in~\cite{Altche2017}.
The current predictive LSTM-based networks commonly use an encoder/decoder architecture such as~\cite{Hu2018}
that builds a Conditional Variational Autoencoder for vehicle trajectory forecasting.

\textbf{Maneuver prediction} is a field of prediction in itself but is also used in trajectory forecasting
as a solution for multimodal forecasting~\cite{Hu2018, Houenou2013, Deo2018}.
The forecast for each mode is a trajectory conditioned to a maneuver.
The global forecast is made using a second module that estimate each maneuver probability to combine all
the trajectory forecasts.
This produces multi-modal forecasts indeed, however the predicted modes correspond to the predefined maneuvers.
%Therefore, only a careful choice of maneuver definitions could match them to
%the actual modes of the data distribution.
As shown in~\cite{Shouno2018}, the various modes in the trajectory data are very complex and numerous.
As a result, capturing them with a few predefined maneuvers is not likely to be enough.
Moreover, as stated in~\cite{Wissing2018} that addresses this issue, the time at which the maneuver occurs in the sequence
is not always defined.
This makes the predicted modes relative to the whole sequence and not to the position at each time step.

\textbf{Probability density function} express stochastic forecasts and control the
uncertainty without depending on sampling as in generative models~\cite{Kuefler2017} and models based on variational
auto-encoding~\cite{Lenz2017, Hu2018}.
As done in~\cite{Lenz2017, Deo2018}, our model expresses forecasts as Gaussian mixture density functions.
However, no maneuver definition as in~\cite{Deo2018} nor latent sampling is needed for our model to
produce a multimodal distribution.

\textbf{Attention mechanisms} in neural networks have been able to introduce inter-dependencies within a variable
number of inputs.
It has been used for pedestrian trajectory forecasting in~\cite{Vemula2018} with spatiotemporal graphs and
in~\cite{Sadeghian2018a} with spatial and social attention using a generative neural network.
In~\cite{Sadeghian2018b}, attention over top-view road scene images for car trajectory forecasting is used.
Multi-head attention mechanism has been developed in~\cite{Vaswani2017} for sentence translation.
In~\cite{Messaoud2019} a mechanism called non-local  multi-head attention is developed.
However, this is a spatial attention that does not allow vehicle to vehicle attention.

In the present architecture, we use a multi-head social attention mechanism working together with
recurrent neural networks to forecast in the form of sequences of multi-modal position probability density
functions.

\section{Objectives}
\label{sec_obj}

The goal is to build a road scene vehicles trajectory prediction model that is versatile yet accurate.
It relies on vehicle tracks that are produced by a perception, fusion and tracking system.
For versatility, the inputs and the architecture are kept simple.
Therefore, the model is easy to adapt to more complete inputs, so it can be improved for specific applications.
The forecasts are made simultaneously for all vehicles in the road scene and accounts for interactions.
Moreover, it is commonly admitted that future trajectories are stochastic with a multimodal distribution.
The forecasts also account for that without depending on any maneuver supervision nor random sampling.
The model is able to work with road scene observations that contain different numbers of vehicles without ordering.

%\subsection{Global definition of the prediction function}
\label{sec_inout}

\textbf{The inputs} are sequences of all vehicle $(x, y)$ positions in a road scene.
At each time $t_0$, we consider an observation history with a fixed observation frequency and a
fixed number of observations $n_{\text{hist}}$.
The history trajectory is written $\{(x, y)_{k}\}_{k=-n_{\text{hist}}+1, 0}$.
The coordinate system is centered on the ego vehicle position at $t_0$.

\textbf{The outputs} are sequences of $n_{\text{pred}}$ Gaussians mixtures for each vehicle.
It is expressed with a sextuplets $(\hat{x}, \hat{y}, \sigma_x, \sigma_y, \rho, p)$
for each vehicle, at each forecast step and for each mixture component.
It defines a Gaussian component $(\mathcal{N}((\hat{x}, \hat{y}), \Sigma), p)$ with
    \[\Sigma =
    \left( \begin{matrix}
        \sigma_x^2 & \rho \sigma_x \sigma_y \\
        \rho \sigma_x \sigma_y & \sigma_y^2
    \end{matrix}\right)
    \]
the covariance matrix, and $p$ the mixture weight such that for $M$ components, $\sum_{m=1}^M{p_m} = 1$.

\textbf{The forecasting model} is a set of functions $pred_\theta : inputs \rightarrow outputs$.
The inputs and outputs sets are defined with the cartesian products:
\begin{equation*}
    \begin{split}
    \text{inputs} &\in \left( \mathbb{R}^{2} \right)^{n_{\text{hist}} \times n_{\text{veh}}}\\
    \text{outputs} &\in \bigg(\Big( {
    \underbrace{\vphantom{\mathbb{R}_+^{2}}
        \mathbb{R}^{2}}_{\hat{x}, \hat{y}}} \times
    {\underbrace{
        \mathbb{R}_+^{2}}_{\sigma_x, \sigma_y}} \times
    {\underbrace{\vphantom{\mathbb{R}_+^{2}}
        [-1, 1]}_{\rho}}\Big)^{n_{\text{mix}}} \times
    {\underbrace{\vphantom{\mathbb{R}_+^{2}}
    \Delta^{n_{\text{mix}}}}_{p}} \bigg)^{n_{\text{pred}} \times n_{\text{veh}}}
    \end{split}
\end{equation*}

$\Delta^{n_{\text{mix}}}$ is the $n_{\text{mix}}$ elements simplex:
\[\Delta^{n} = \Big\{(p_1,\dots,p_n)\in\mathbb{R}^{n}~\Big|~\sum_{i = 1}^n p_i = 1,~ p_i \ge 1~ \forall i\Big\}\]

The $pred_\theta$ function set is defined as a neural network with weights $\theta$.
It is invariant to permutations along the vehicle axis and it
is defined for all number of vehicles $n_{\text{veh}}$ and forecast steps $n_{\text{pred}}$.

\section{Model architecture}

This model uses an encoder-decoder structure.
It is based on LSTM networks for encoding and forecasting.
Here, we propose to add two multi-head self-attention layers to this architecture for interaction awareness.
The first attention layer is added after encoding to incorporate current interactions in the encoded vector of each vehicle.
The second attention layer is added after forecasting and before decoding.
This allows the forecast position sequences to remain coherent with each other.

The benefit of using self-attention layers for interactions is that the whole model computations are defined for
any number of vehicles and invariant to their ordering in the input tensor.
Since all other computations are the same for each vehicle, this property is kept throughout the whole model.

\subsection{Global architecture}

\begin{figure}[ht]
   \centering
    \begin{tikzpicture}
        \node(X1){input$_1$};
        \node[below of=X1, node distance=0.6cm](X2){$\vdots$};
        \node[below of=X2, node distance=0.7cm](X3){input$_{n_{\text{veh}}}$};

        \node[draw, right of=X1, node distance=1.5cm, rectangle](ENC1){Enc.};
        \node[below of=ENC1, node distance=0.6cm](ENC2){$\vdots$};
        \node[draw, right of=X3, node distance=1.5cm, rectangle](ENC3){Enc.};

        \path (X1) edge (ENC1);
        \path (X3) edge (ENC3);

        \node[draw, rectangle, right of=ENC2, node distance=1cm](TRANS1){\rotatebox{90}{Self-attention}};

        \draw (ENC1.east) -| (TRANS1.west);
        \draw (ENC3.east) -| (TRANS1.west);

        \node[draw, right of=ENC1, node distance=2cm, rectangle](PRED1){Pred.};
        \node[below of=PRED1, node distance=0.6cm](PRED2){$\vdots$};
        \node[draw, below of=PRED2, node distance=0.7cm, rectangle](PRED3){Pred.};

        \draw (TRANS1.east) |- (PRED1.west);
        \draw (TRANS1.east) |- (PRED3.west);

        \node[draw, rectangle, right of=PRED2, node distance=1cm](TRANS2){\rotatebox{90}{Self-attention}};

        \draw (PRED1.east) -| (TRANS2.west);
        \draw (PRED3.east) -| (TRANS2.west);

        \node[draw, right of=PRED1, node distance=2cm, rectangle](DEC1){Dec.};
        \node[below of=DEC1, node distance=0.6cm](DEC2){$\vdots$};
        \node[draw, below of=DEC2, node distance=0.7cm, rectangle](DEC3){Dec.};

        \draw (TRANS2.east) |- (DEC1.west);
        \draw (TRANS2.east) |- (DEC3.west);

        \node[right of=DEC1, node distance=1.5cm](Y1){output$_1$};
        \node[right of=DEC2, node distance=1.5cm](Y2){$\vdots$};
        \node[right of=DEC3, node distance=1.5cm](Y3){output$_{n_{\text{veh}}}$};

        \draw (DEC1.east) -- (Y1.west);
        \draw (DEC3.east) -- (Y3.west);

%        \draw[decorate,decoration={brace,mirror}] (X1.north west) -- node[left=10pt] {$X$} (X3.south west);
    \end{tikzpicture}
    \caption{Bloc representation of our forecasting model.
    Inputs are the sequence of past observations of each vehicle.
    Outputs are the matching forecasts.}
    \label{sch_whole_model}

\end{figure}

The figure~\ref{sch_whole_model} is a bloc representation of our forecasting model.
It is described in four parts: Encoder, Self-Attention, Predictor, and Decoder.
The two self-attention layers are different layers with the same architecture whereas the
encoder, predictor, and decoder are the same for every input.

\subsection{Encoder}

The encoder should act as a current state estimation for each vehicle using the past observation sequences.
This state is an intermediary vector of the neural network and is difficult to interpret.
However, since it should encode the current state with at least the information of position, kinematic state,
and interaction features, it should have a sufficient dimension, we chose 128.
The input $(x, y)$ position sequences are fed to a one dimensional convolutional layer with a kernel
of size 3 sliding over the time dimension that creates sequences of 128 features for each vehicle.
Then each feature sequence is encoded with a Long Short-Term Memory (LSTM)~\cite{Hochreiter1997}
into a vector of 128 features for each vehicle.

\subsection{Self-attention}
\label{sec_self_attention}

The multi-head self-attention layers allow vehicle interactions while keeping independence from their number
and ordering.
This mechanism is described in~\cite{Vaswani2017} where it is applied on sentence translation.
In this section we explain its use for vehicle interactions.
We tried $4$ and $8$ attention heads for both layers and used $4$ in the application.
The computations made by each attention head is represented on figure~\ref{sch_selfattention} and are detailed below.

\begin{figure}[ht]
    \centering
    \begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
        \node(X1){input$_1$};
        \node[below of=X1, node distance=2cm](X2){$\vdots$};
        \node[below of=X2, node distance=2cm](X3){input$_{n_{\text{veh}}}$};

        \coordinate[right of= X1, node distance=1cm](X1b){};

        \draw (X1) -- (X1b);

        \node[draw, right of=X1b, node distance=1cm](LQ1){$L_{q}$};
        \node[draw, above of=LQ1, node distance=1cm](LV1){$L_{v}$};
        \node[draw, below of=LQ1, node distance=1cm](LK1){$L_{k}$};

        \draw (X1b) -- (LV1);
        \draw (X1b) -- (LQ1);
        \draw (X1b) -- (LK1);

        \node[right of=LV1, node distance=1cm](V1){$\mathbf{v}_1$};
        \node[right of=LQ1, node distance=1cm](Q1){$\mathbf{q}_1$};
        \node[right of=LK1, node distance=1cm](K1){$\mathbf{k}_1$};

        \draw (LV1) -- (V1);
        \draw (LQ1) -- (Q1);
        \draw (LK1) -- (K1);

        \coordinate[right of= X3, node distance=1cm](X3b){};

        \draw (X3) -- (X3b);

        \node[draw, right of=X3b, node distance=1cm](LQ3){$L_{q}$};
        \node[draw, above of=LQ3, node distance=1cm](LV3){$L_{v}$};
        \node[draw, below of=LQ3, node distance=1cm](LK3){$L_{ k}$};

        \draw (X3b) -- (LV3);
        \draw (X3b) -- (LQ3);
        \draw (X3b) -- (LK3);

        \node[right of=LV3, node distance=1cm](V3){$\mathbf{v}_{n_{\text{veh}}}$};
        \node[right of=LQ3, node distance=1cm](Q3){$\mathbf{q}_{n_{\text{veh}}}$};
        \node[right of=LK3, node distance=1cm](K3){$\mathbf{k}_{n_{\text{veh}}}$};

        \draw (LV3) -- (V3);
        \draw (LQ3) -- (Q3);
        \draw (LK3) -- (K3);

        \coordinate[right of=V1, node distance=0.3cm](TOP){};
        \coordinate[right of=K3, node distance=0.3cm](BOT){};
        \draw[decorate,decoration={brace}] (TOP) -- node[left=5pt]{} (BOT);

        \node[right of=X2, text width=3cm, node distance=4.5cm](EQ){\footnotesize \[ V = \left( \begin{matrix}
                                                            \mathbf{v}_1 \\
                                                            \vdots \\
                                                            \mathbf{v}_{n_{\text{veh}}}
                                                          \end{matrix}\right) \]
                                                    \\
                                                   \footnotesize \[ Q = \left( \begin{matrix}
                                                            \mathbf{q}_1 \\
                                                            \vdots \\
                                                            \mathbf{q}_{n_{\text{veh}}}
                                                           \end{matrix} \right)\]
                                                    \\
                                                    \footnotesize \[K &= \left( \begin{matrix}
                                                            \mathbf{k}_1 \\
                                                            \vdots \\
                                                            \mathbf{k}_{n_{\text{veh}}}
                                                          \end{matrix} \right)\]
                                                  };

        \node[right of=EQ, node distance=1.5cm, rotate=90](EQ2){
        \footnotesize $\text{output}=\underset{\text{dim=last}}{\operatorname{Softmax}}\left(\frac{QK^T}{\sqrt{d_k}}\right)V$};

        \coordinate[right of =EQ2, node distance=1cm](CO2){};
        \coordinate[above of=CO2, node distance=2cm](CO1){};
        \coordinate[below of=CO2, node distance=2cm](CO3){};

        \node[right of=CO2, node distance=1cm](O2){$\vdots$};
        \node[right of=CO1, node distance=1cm](O1){output$_1$};
        \node[right of=CO3, node distance=1cm](O3){output$_{n_{\text{veh}}}$};

        \draw (CO1) -- (O1);
        \draw (CO3) -- (O3);

        \draw[decorate,decoration={brace, mirror}, xshift=-2cm] (CO1) -- node[left=15pt]{} (CO3);

    \end{tikzpicture}
    \caption{Schematic representation of one attention head computations.
    Blocs $L_{q}$, $L_{v}$, $L_{k}$ are matrix multiplications of the input vectors.}
    \label{sch_selfattention}
\end{figure}

Each vehicle should pay attention to specific features from a selection of the other vehicles.
This is made with four steps: pulling together specific features, identifying these feature collections,
enquiring among identifiers, and gathering the results.
Each head produces a different selection of features using a linear projection of
the input tensor resulting in the value tensor $V$.
To identify these features, a key tensor $K$ is associated to each value.
Then, each vehicle must select which other vehicle to pay attention to.
For that purpose a query $Q$ is produced to find a selection of keys.
The match score between a key and a query is the dot product is scaled with the square root of the key
dimension $\sqrt{d_k}$ and normalized with a softmax.
This produces an attention matrix that contains coefficients close to $1$ for matching queries and keys
and close to $0$ otherwise.
Finally, this matrix is used to gather the values from $V$.
Thus, the self-attention computation for each head is written:
\begin{equation}
    \text{output}=\underbrace{\underset{\text{dim}=\text{last}}{\operatorname{Softmax}}\left(\frac{QK^T}{\sqrt{d_k}}\right)}_{\text{attention matrix}}V
    \label{eq_selfattention}
\end{equation}
The outputs from all heads are concatenated and combined with a linear layer.
The resulting tensor is then added to the input as in residual networks.

\subsection{Predictor}
Tensors produced with the self-attention layer are repeated $n_{\text{pred}}$ times to be fed as time sequences
into a second LSTM layer.
This produces intermediary forecasts with some interaction awareness.
Within the forecast position sequences, vehicle interactions may have changed.
Thus, we placed a second multi-head self-attention layer before feeding the output to the decoder.

\subsection{Decoder}

Feature sequences are decoded with two convolutional layers with kernel size 1 and relu activations.
Finally, a last convolutional layer with kernel size 1 with a specific activation produces
the mixture of Gaussian output.
The output is described in section~\ref{sec_inout}.
Let $o_i$ be the $i^{\text{th}}$ coordinate of the output tensor before the activation function.
To constraint it, a specific activation function is applied on each coordinate at every time steps:
\begin{equation*}
    \begin{split}
        &\text{activation}(\{o_1, o_2, o_3, o_4, o_5, o_6\}_{i=1,n_{\text{mix}}})\\
        &~=\{(\hat{x}, \hat{y}, \sigma_x, \sigma_y, \rho, p)\}_{i=1,n_{\text{mix}}}\\
        &~=\Big\{(o_1, o_2, e^{\frac{o_3}{2}}, e^{\frac{o_4}{2}}, \operatorname{tanh}(o_5), \underset{i \in \text{mix}}{\operatorname{Softmax}}(o_6))_i\Big\}_{i=1,n_{\text{mix}}}
    \end{split}
\end{equation*}

\section{Loss and performance indicators}
\label{sec_loss}

The model is trained with the Adam optimizer~\cite{Kingma2015} that minimizes the negative log-likelihood (NLL) loss.
The usual performance indicators for such forecasting models are root mean squared error (RMSE),
mean absolute error (MAE), and negative log-likelihood (NLL).

\textbf{The RMSE} computation is made with equation~\eqref{eq_rmse} with $(x_k^i, y_k^i)$ the observed positions and
$(\hat{x}_k^i, \hat{y}_k^i)$ the most probable forecast positions of the $i^{th}$ sequence at time $t_k$.
$N$ is the number of sequences in the subset of the database on which the computation is made.

\begin{equation}
    \text{RMSE}(k) = \sqrt{\frac{1}{N}\sum_{i=1}^N{(x_k^i - \hat{x}_k^i)^2 + (y_k^i - \hat{y}_k^i)^2}}
    \label{eq_rmse}
\end{equation}

\textbf{The MAE} computation is made with equations~\eqref{eq_ade}.

\begin{equation}
    \begin{split}
        \text{MAE}_x(k) &=\frac{1}{N}\sum_{i=1}^N{|(x_k^i - \hat{x}_k^i)|}\\
        \text{MAE}_y(k) &=\frac{1}{N}\sum_{i=1}^N{|(y_k^i - \hat{y}_k^i)|}
    \end{split}
    \label{eq_ade}
\end{equation}

\textbf{The NLL} computation, at each forecast time $t_k$, for each Gaussian component centered on $(\hat{x}, \hat{y})$,  with the forecast
error\\
$\mathbf{d} = (d_x, d_y) = (x - \hat{x}, y -\hat{y})$ and the forecast error
covariance defined with $(\sigma_x, \sigma_y, \rho)$ is written:

\begin{equation}
    \begin{split}
        \text{NLL}(d_x, d_y, \Sigma) =
        & \frac{1}{2}\underbrace{\frac{1}{(1 - \rho^2)}\biggl(\frac{d_{x}^2}{\sigma_{x}^2} + \frac{d_{y}^2}{\sigma_{y}^2}
        -2 \rho \frac{d_{x} d_{y}}{\sigma_{x} \sigma_{y}}\biggr)
        }_{\mathbf{d_k}^T\Sigma_k^{-1}\mathbf{d_k}}\\
        & + \underbrace{\ln\biggl(\sigma_x\sigma_y\sqrt{1-\rho^2}\biggr)}_{\ln(\sqrt{|\Sigma_k|})}\\
        & + \ln(2\pi)
    \end{split}
    \label{eq_gauss_nll}
\end{equation}

The time index $k$ is dropped to improve readability.
The computation of the overall NLL value for all mixture components is written:

\begin{equation}
    \text{NLL}(d_x, d_y, \Sigma, p) = -ln\left( \sum_{m=1}^{n_{\text{mix}}} p_m e^{-\text{NLL}(d_{x_m}, d_{y_m}, \Sigma_m)} \right)
    \label{eq_multi_nll}
\end{equation}

The mean NLL (MNLL) over the test set is written:
\begin{equation}
    \text{MNLL} = \frac{1}{N}\sum_{i=1}^N{\text{NLL}_i}
\end{equation}

Minimizing the NLL loss maximizes the likelihood of the data for the forecast.
However, the optimization is not convex and one spurious local minimum that is often obtained is
%a unimodal distribution.
%Another spurious local minimum is
the overfitting of common behaviors on short horizons.

\emph{To avoid short-term overfitting}, the forecast variances are clipped with a 10cm minimum value.
This is an acceptable minimum variance, for long term forecasts that is not restrictive.
The gradient descent optimization with NLL loss tend to favor forecasts with small covariances and errors because the
NLL derivative is larger in those cases.
In~\cite{Lenz2017}, NLL overfitting has been obtained and has degraded the results, making the NLL value unreliable as a
performance indicator.
Clipping the variance values avoids large values of gradient for forecasts that where already well fitted.
This helps the optimization process avoiding this source of overfitting and allows us to rely on the NLL values.

%\emph{To encourage multimodal predictions}, an additional loss that penalize the similarity is added.
%A method using Detrimental Point Process (DPP) adapted from~\cite{Yuan2019} is used.
%The mode distance matrix $D \in \mathbb{R}^{n_{\text{mix}}, n_{\text{mix}}}$ defined as the
%difference between predicted modes positions:
%\begin{equation}
%    D_{i,j} = ||(x_i, y_i) - (x_j, y_j)||_2^2
%\end{equation}
%With $p$ the mixture factor vector
%%with values clipped within $[\epsilon, 1]$
%, and $I$ the identity matrix, the
%DPP kernel \\
%$L(D,p) = \exp(-\alpha~ diag(p) ~D~ diag(p))$ is used to define the similarity penalization:
%\begin{equation}
%    L_{diverse} = -tr(I - (L(D,p) + I)^{-1})
%\end{equation}
%
%In~\cite{Yuan2019}, using the mixing coefficient as a DPP kernel quality factor is said to favor major modes.
%This is detrimental if all modes are to be predicted.
%However, our goal is to predict the most likely future positions and not all possible future modes.
%
%\emph{To encourage multimodal predictions}, an additional loss that penalize the similarity is added.
%\begin{figure}[H]
%    \centering
%    \begin{tikzpicture}
%     \begin{axis}[
%            x=0.8cm,
%            y=5cm,
%            xmin=1,
%            xmax=10,
%            ymin=0,
%            ymax=0.4,
%            axis lines=center,
%            ticks=none,
%    %        xlabel={$k_t$},
%    %        ylabel={$f(k_t)$},
%    %        xlabel style={below right},
%    %        ylabel style={above left},
%            % (moved common `addplot' options here)
%            smooth,
%            domain=0:10,
%            samples=101,
%            no markers,
%        ]
%
%    % Draw curves
%    \addplot [thin, smooth, name path global=first] {dnorm(4,1)};
%    \addplot [thin, smooth, name path global=second] {dnorm(5.7,1.5)};
%
%    % Draw vertical line:
%    \path [name path global=first_line] ({rel axis cs:0,0}-|{axis cs:4,0}) -- ({rel axis cs:0,1}-|{axis cs:4,0});
%    \path [name path global=second_line] ({rel axis cs:0,0}-|{axis cs:5.7,0}) -- ({rel axis cs:0,1}-|{axis cs:5.7,0});
%
%    \path [name intersections={of=first_line and first}] (intersection-1) circle
%                coordinate (m1);
%    \path [name intersections={of=second_line and second}] (intersection-1) circle
%                coordinate (m2);
%
%    \node at ([shift={(6, 4)}]m1) [below right = 1cm of m1] {$\mathcal{N}_1$};
%    \node at ([shift={(6, 5)}]m2) [right = 1cm of m2] {$\mathcal{N}_2$};
%%    \draw [thick](m1) node [dot] {}
%%                -- (m1) node[dot, label:] {};
%%    \draw (m2) node[dot, label:right:$\mathcal{N}_2$] {};
%
%    \path [name intersections={of=first_line and second}] (intersection-1) circle
%                coordinate (p1);
%    \path [name intersections={of=second_line and first}] (intersection-1) circle
%                coordinate (p2);
%
%    \draw [dashed] (0, 0 -|  m1) node [dot]{}
%        -- (m1)                node [dot] {};
%    \draw [dashed] (0, 0 -|  m2) node [dot]{}
%        -- (m2)                node [dot] {};
%    \draw [very thick, <->] (0, 0 -|  p1) node [dot]{}
%        -- (p1)                node [dot, label=left:$p_{1/2}$] {};
%    \draw [very thick, <->] (0, 0 -|  p2) node [dot]{}
%        -- (p2)                node [dot, label=left:$p_{2/1}$] {};
%    %\draw [thick] (p2 -| 0, 0) node [dot, label=p1]{};
%    %\draw [thick, name intersections={of={second_line and first}}] ({rel axis cs:0,0}-|intersection-1) -- ({rel axis cs:0,1}-|intersection-1);
%    \end{axis}
%    \end{tikzpicture}
%
%    \caption{Graph of two Gaussian probability density functions (PDF). $p_{1/2}$ and $p_{2/1}$ are the values of one PDF computed
%    at the mean value of the other.}
%    \label{plt_gauss}
%\end{figure}
%
%Figure~\ref{plt_gauss} is an example representation of the intermediate computations used for
%the similarity penalization.
%For each couple of Gaussian mixture component $i$, and $j$ with $i \neq j$, the probability density function of one component
%is computed at the mean value of the other.
%This produces the two values $p_{i/j}$ and $p_{j/i}$.
%If the variance of one of the component $i$ is lower or equal to the clipping value defined in the previous paragraph,
%$p_{i/j}$ is set to $0$.
%The similarity penalization is defined as the mean value of the pair values product as written in equation~\eqref{eq_sim_loss}.
%\begin{equation}
%    \text{SIM} = \frac{1}{n_{\text{mix}}^2}\sum_{i \neq j} p_{i/j}p_{j/i}
%    \label{eq_sim_loss}
%\end{equation}
%
%For the closest horizon predictions, the variance is equal to the clipping value keeping the
%similarity penalization to $0$.
%This is a good behavior since the prediction at 0.2s does not need to be multimodal.
%As the prediction horizon increase, the variance also increase.
%Then the similarity penalization lowers the variances and pushes the mean values further from each other.
%This pushes predicted mixture of Gaussian toward a multimodal distribution.
%This penalization is on average several order of magnitude lower than the average loss and thus affects only slightly
%the predictions when they are multimodal.

%\subsection{Local attention}
%
%The attention mechanism makes combinations of each vehicles with every other.
%This is too broad and it should focus more on the local vehicles.
%To force this behavior, the equation~\eqref{eq_selfattention} is modified as follow:
%\begin{equation}
%    output=\underset{dim=last}{Softmax}\left(\frac{QK^T}{\sqrt{d_k}} - D\right)V
%    \label{eq_localattention}
%\end{equation}
%The distance tensor $D$ is a tensor whose last two axis form symetric matrices of coefficients
%$d_{i, j} = ||pos_{i} - pos_{j}||_2^2$ with $pos_i$ the position of the $i^{th}$ vehicle.
%It is substracted to the key/query product to lower the value corresponding distant vehicles.
%For the first attention layer, distances are computed directly from the last input positions.
%For the second attention layer, the output layer is used to extract positions from the predicted encoded sequences.
%The distance matrices are computed at every prediction step using these positions.
%Then local attention is computed with the same predicted encoded sequences and the associated distance matrices.
\section{Architecture discussion}

\subsection{Multi-head self-attention}

The general idea of this architecture is to use the good properties of the key-query self-attention layer to account for
interactions.
This offers flexibility to the model allowing powerful LSTM models to compute the features and predictions with
the fixed size inputs it demands while accepting a varying number of interacting vehicles without ordering.
All input vehicles are treated exactly the same way even if the perception of the road scene is only partial.
This allows the simultaneous forecast of each vehicle in the scene with a different and interdependent
social context for each.

This method is sometimes stated as computationally expensive because attention over $n$ objects relies on computations
involving an $n \times n$ attention matrix.
However, in our use cases, the number of objects is lower than $n=30$ and most of the times remains around $10$.
This small number makes global attention affordable in all of our test cases.
If another application uses a larger number of objects, local attention could be considered.
As stated in~\cite{Vaswani2017}, and shown in our application section~\ref{sec_application_attention}, multi-head
attention produces interpretable interactions with each head specializing on different interaction patterns.

\subsection{Hyperparameters}

This model is defined with a few specific hyperparameters that should be tuned: number of encoded features, number of
embedding and decoding layers and their activation functions, the number of heads in each self attention layer,
the number of mixture component in the output distribution and the error covariance clipping value.
Other choices have been made and should be questioned such as the data normalization (none is used),
the use of shortcut connections without layer normalization, the use of LSTM layers and some implicit choices
may have been overlooked.
Optimizing the hyperparameters with a thorough process could bring some improvements and
help understand the model but is not a part of the present study.
In this work, only the general concept was prioritized and the hyperparameters were chosen from experience.

\subsection{Maneuver free multimodal forecast}

With the present model, after defining a constant number of mixture components, they are differentiated solely with
the minimization of the loss.
%The loss is the sum of NLL and similarity penalisation.
The loss is only the NLL value averaged over time.
Thus, minimizing it pushes the predicted modes toward those of the data distribution.
What is forecast is not a mixture of trajectory density functions but a sequence of multimodal position
density functions.
There is a dependency between forecasts at time $t_k$ and at time $t_{k+1}$ but no explicit link between the modes
at those times.
Trajectories can be defined as the optimal transport path between the Gaussian mixture at time $t_k$ and the one at
time $t_{k+1}$.
To simplify, we make the assumption that mixture components centers define local maxima of the probability distributions
and can be tracked in time by matching similar mixture coefficients.

\subsection{How to extend this?}

The simplest extension is to add additional observations on each vehicle such as velocity, orientation, size, blinkers
and any other useful observation.
Since this is a model-free architecture, it can be done without changing it.

Another simple extension is to use a dictionary of encoders and decoders matching an input dictionary of moving
road scene object classes to be forecast.
No change is needed in the attention layer to allow attention between the different class of objects.

In the present architecture, self-attention produces keys, queries, and values from the same input to transform the
input value.
However, the keys and values can be extended with additional inputs with minimal change to the model.
The new input should be encoded with a separate encoder.
It may be for example, the encoding of the lane center line discretized as a sequence of
position points.
A bidirectional LSTM could be a good encoder architecture for this kind of input.
Then, each head of the first attention layer is extended with two additional linear layers $L_{vext}$ and $L_{kext}$,
both producing the same number of outputs and each with the same feature dimension
as the other linear layers from the attention head.
With this new input, extended keys and values are produced :
\begin{equation}
    \begin{split}
        \text{encoded}_{\text{ext}} &= \operatorname{extEncoder}(\text{input}_{\text{ext}})\\
        V_{\text{ext}} &= L_{v_{\text{ext}}}(\text{encoded}_{\text{ext}})\\
        K_{\text{ext}} &= L_{k_{\text{ext}}}(\text{encoded}_{\text{ext}})
    \end{split}
\end{equation}
These new keys and values can simply be concatenated to the head usual keys and values:
\begin{equation}
    \text{output}=\underset{\text{dim}=\text{last}}{\operatorname{Softmax}}\left(\frac{Q[K, K_{\text{ext}}]^T}{\sqrt{d_k}}\right)[V, V_{\text{ext}}]
    \label{eq_selfattention_extended}
\end{equation}
The output matches the vehicle features with an additional attention over the new input.

The ease of extension is one of the best asset of this model.
However, for conciseness and simplicity, only the core model is studied in the application.

\section{Application}

This model was implemented with the Pytorch library.
The NGSIM datasets US-101 and I-80 and its pre-processing were taken from
the published code accompanying the article~\cite{Deo2018}.
This also defines the dataset splitting into training, validation, and test sets.
Thus, a fair comparison with these results is made.
The dataset contains the tracks of all vehicle position on a road segment observed from a camera.
The pre-processing produces data that simulates observations from a given vehicle.
Each vehicle is alternatively chosen as the observing vehicle.
Its surroundings in adjacent lanes and within a given distance are recorded to produce a local road scene.
This road scene is tracked to produce 8 seconds sequences with all positions being recorded at a 5Hz frequency.
The 3 first seconds are used as past observations and the 5 next seconds are used as forecast supervision.

\begin{figure*}
        \centering
%        left btm right top
        \includegraphics[width=0.11\textwidth, clip, trim=0.49cm 5.06cm 43.5cm 0.1cm]{dessins/attention_look_front_131.pdf}
        \includegraphics[width=0.88\textwidth, clip, trim=10cm 4cm 0.1cm 1.55cm]{dessins/attention_look_front_131.pdf}
        \vspace{-10pt}
        \caption{
        A driving scene top view representation with all observed vehicles and the
        attention matrix for one head of the first attention layer.
        The attention that vehicle $i$ is giving to $j$ is drawn as an arrow from $i$ to $j$ with a width proportional
        to the attention coefficient and a color varying with its angle.
        A yellow circle represents the attention of one vehicle on itself (when $i=j$) with a radius proportional to
        the coefficient.
        The attention coefficients are represented as colors in the matrix on the left going from purple at $0$ to
        yellow at $1$.
        Past positions are represented in gray.
        Superposed forecast probability density functions are represented in blue shades in log scale.}
        \label{fig_veh_att}
        \vspace{-10pt}
\end{figure*}

\subsection{Global performance indicators}

\begin{table}
    \centering
    \caption{Comparison of RMSE, NLL and MAE results with baselines using the same dataset.
     *CSP(M) results were recomputed with some minor modifications for a fair comparison.}
    \begin{tabular}{l|l|ccccc}
        \hline
        \multicolumn{2}{l|}{\hspace{-5pt}Time horizon}                & 1s   & 2s   & 3s   & 4s   & 5s   \\
        \hline\hspace{-3pt}
        \multirow{3}{*}{\hspace{-5pt}RMSE\hspace{-3pt}} &\hspace{-3pt}CV~\cite{Mercat2019}      & 0.76 & 1.82 & 3.17 & 4.80 & 6.70 \\
                              &\hspace{-3pt}CSP(M)~\cite{Deo2018}*    & 0.57 & 1.25 & 2.10 & 3.17 & 4.49 \\
                              &\hspace{-3pt}GRIP~\cite{Li2019}        & \textbf{0.37} & \textbf{0.86} & \textbf{1.45} & \textbf{2.21} & \textbf{3.16} \\
                              &\hspace{-3pt}\textit{SAMMP$_4$}          & 0.49 & 1.10 & 1.84 & 2.77 & 3.93\\
                              &\hspace{-3pt}\textit{SAMMP$_8$}          & 0.49 & 1.10 & 1.84 & 2.77 & 3.93\\
%                              &\hspace{-3pt}$SAMMP_{MSE}$             & NA & NA & NA & NA & NA\\
        \hline
        \multirow{3}{*}{\hspace{-5pt}MNLL\hspace{-3pt}} &\hspace{-3pt}CV~\cite{Mercat2019}      & 0.82 & 2.32 & 3.23 & 3.91 & 4.46\\
                              &\hspace{-3pt}CSP(M)~\cite{Deo2018}*    & 1.78 & 3.35 & 4.24 & 4.88 & 5.47\\
                              &\hspace{-3pt}\textit{SAMMP$_4$}          & \textbf{0.45} & \textbf{1.11} & \textbf{1.73} & \textbf{2.26} & \textbf{2.72}\\
                              &\hspace{-3pt}\textit{SAMMP$_8$}          & \textbf{0.45} & \textbf{1.11} & \textbf{1.73} & \textbf{2.26} & \textbf{2.72}\\
%                              &\hspace{-3pt}$SAMMP_{MSE}$             & NA & NA & NA & NA & NA\\
        \hline
        \multirow{2}{*}{\hspace{-5pt}MAEx\hspace{-3pt}}&\hspace{-3pt}CV~\cite{Mercat2019}      & 0.42 & 1.17 & 2.19 & 3.44 & 4.90\\
                              &\hspace{-3pt}\textit{SAMMP$_4$}           & 0.29 & 0.74 & 1.28 & 1.96 & 2.82\\
                              &\hspace{-3pt}\textit{SAMMP$_8$}           & 0.29 & 0.74 & 1.28 & 1.96 & 2.82\\
%                              &\hspace{-3pt}$SAMMP_{MSE}$             & NA & NA & NA & NA & NA\\
        \hline
        \multirow{2}{*}{\hspace{-5pt}MAEy\hspace{-3pt}}&\hspace{-3pt}CV~\cite{Mercat2019}      & 0.11 & 0.21 & 0.31 & 0.39 & 0.48\\
                              &\hspace{-3pt}\textit{SAMMP$_4$}           & 0.06 & 0.11 & 0.15 & 0.19 & 0.22 \\
                              &\hspace{-3pt}\textit{SAMMP$_8$}           & 0.06 & 0.11 & 0.15 & 0.19 & 0.22 \\
%                              &\hspace{-3pt}$SAMMP_{MSE}$             & NA & NA & NA & NA & NA \\
        \hline
    \end{tabular}
    \label{tab_perf}
    \vspace{-10pt}
\end{table}

Table~\ref{tab_perf} reports results using the performance indicators defined in section~\ref{sec_loss}.
%These results are compared with the following baselines from the literature.
All compared models except for GRIP~\cite{Li2019} were trained and computed on the same dataset and evaluated with the same functions.
Since the model from~\cite{Deo2018} only forecasts the observing vehicle trajectory, only the errors for this vehicle are
being compared.

\textbf{Baselines}:

\emph{Constant velocity} (CV): We used a constant velocity Kalman filter with optimized parameters for forecasting
on the same data as described in~\cite{Mercat2019}.

\emph{Convolutional Social Pooling} (CSP(M)): We retrained the model from~\cite{Deo2018}.
It uses a maneuver classifier trained with preprocessed data that conditions a predictor for multimodal forecasts.
A forecast of the center vehicle trajectory is made with information from its social environment using the convolutional social
pooling mechanism.

\emph{Graph-based Interaction-aware Trajectory Prediction} (GRIP): We took the published results from~\cite{Li2019}.
It uses a spatial and temporal graph representation of the scene to make a maximum likelihood trajectory
prediction simultaneously fol all vehicles in the scene.
They obtain the best results in term of RMSE but it does not account for error estimation nor multimodality.

\emph{Social Attention Multi-Modal Prediction} (SAMMP): The model described in this article. SAMMP$_4$ and SAMMP$_4$
are the same model with 4 and 8 mixture components.

Our model forecasts multimodal probability density functions, thus the RMSE and MAE values that only relate to the most
probable forecast give only partial evaluations.
The mean NLL value however is relative to the whole distribution and is the preferred performance indicator.
As discussed in~\cite{Mercat2019}, this value is computed with different definitions in some other publications.
This is why results from~\cite{Deo2018} had to be recomputed.
MAE values are given for further comparisons.
They are less sensitive to very large errors that may be produced by observation errors that exist in the NGSIM dataset.
%Comparing them with RMSE values gives an idea of the error distribution.

\subsection{Attention interpretation}
\label{sec_application_attention}

The attention matrices give insights about the importance of some interactions.
Some of the heads roles can be rationalized by looking at the attention matrix it produces in different contexts.
For example, after every tested trainings of the model, one of the heads had specialized for front vehicle attention.
Figure~\ref{fig_veh_att} shows the attention scores as arrows between vehicles.
The main attention link always goes from one vehicle to the vehicle in front of it, or to itself if there is no
front vehicle.
The attention coefficient decreases with the increasing front vehicle distance.

\subsection{Multi-modal forecasting}

\begin{figure}[ht]
        \centering
%         trim=left botm right top
        \includegraphics[width=\linewidth, clip, trim=13.5cm 2.7cm 3.6cm 0.5cm]{dessins/multimodal_131.pdf}
        \vspace{-10pt}
        \caption{
        Another driving scene top view representation.
        Superposed forecast probability density functions of the ego position
        are represented in blue shades in log scale.
        Ground truth future positions are represented with a green line.
        }
        \label{fig_multimod}
        \vspace{-10pt}
\end{figure}

On the figure~\ref{fig_multimod}, the vehicle 0 aggressively overtakes the vehicle 3.
In this situation, the forecasts are multimodal.
The overtake could be aborted or be made less aggressively, also the last observations of
acceleration and turning could be the results of perception errors.

The NLL loss training is sufficient to produce this multimodal output and for simplicity this is what we used.
However, producing a better mode diversity could avoid mode averaging and mode collapsing.
This can be achieved with the introduction of a loss penalizing the mixture component similarity.
A good example of such a method is presented in~\cite{Yuan2019}.

%The predicted modes are relative to each position and not to the whole trajectory.
%However, it would be inconsistent for a series of modes describing a lane change to significantly change
%its probability over time.
%Thus, it is possible to draw predictions as trajectories by simply matching the modes with similar probabilities.
%This is the representation that is made on figure~\ref{fig_multimod}.
%In this case, the ego vehicle in blue is overtaking the

\section{CONCLUSIONS}

We proposed a road scene forecasting solution that produces multimodal probability function forecast
simultaneously for each vehicle of the scene.
Our method generates interpretable interaction models that could be extended to other road scene observations.
Experiments have surpassed state-of-the-art results with the NLL indicator.
This shows a good forecasting capacity as well as a good error estimation.
Future work will produce some of the proposed extensions and will experiment with the recently published datasets
such as Nuscenes~\cite{Nuscenes2019} and Argoverse~\cite{Chang2019}.

\subsubsection*{Acknowledgements} We are grateful to Edouard Leurent for his comments and corrections.

%The self-attention layers perform a convex combination of the input linear transformations.
%Interactive behaviors are not computed in this layer, it merely shares useful features among vehicles
%for the next layers to compute interactive behaviors.

%DumbPredictor_80_bis
%bias x [[-0.00379273  0.01119626]
% [-0.01110076  0.04440805]
% [-0.01560049  0.05443121]
% [-0.01796214  0.08589871]
% [-0.01867467  0.06826853]]
%bias y [[ 0.04342232 -0.00302752]
% [ 0.05297662  0.01595682]
% [ 0.06110188  0.01036763]
% [ 0.07085429  0.03495284]
% [ 0.07944416  0.0194513 ]]
%bias dist [[0.04421014 0.01260572]
% [0.05681306 0.05820303]
% [0.06842441 0.05716385]
% [0.0816729  0.10115807]
% [0.09105077 0.07447586]]
%bias \% [[9.05066888 2.58063485]
% [5.17432525 5.30091854]
% [3.72854133 3.11493774]
% [2.95013431 3.65396459]
% [2.31402351 1.89277796]]
%ADE [[0.05927011 0.29288597]
% [0.11125454 0.7365424 ]
% [0.15418566 1.28001025]
% [0.19065665 1.96462545]
% [0.22316268 2.82116637]]
%rmse most probable [0.48847378 1.09798007 1.83515226 2.76844685 3.93473846]
%rmse mean prob [0.50515508 1.10888453 1.83993835 2.76945954 3.93497273]
%nll [0.45118189 1.10969029 1.73349189 2.25706048 2.72143882]

%kalmantransformerGRU_8head_nll_46
%RMSE
%0.57939973 1.21238563 1.98451525 2.96286819 4.184847
%NLL
%-0.31314428  1.17847418  2.01719778  2.64240214  3.16598194

%CSLSTM no man:
%NLL:
%-2.0587, -0.3327, 0.7591,  1.4972,  2.0458,
%2.4783,  2.8292,  3.1230,  3.3774,  3.6037,
%3.8075,  3.9930,  4.1645,  4.3241,  4.4758,
%4.6186,  4.7540,  4.8831,  5.0065,  5.1244,
%5.2382,  5.3473,  5.4537,  5.5567,  5.6571
%RMSE:
%0.1052, 0.2111, 0.3223, 0.4436, 0.5680,
%0.6965, 0.8285, 0.9624, 1.1003, 1.2430,
%1.3916, 1.5469, 1.7099, 1.8814, 2.0619,
%2.2514, 2.4500, 2.6580, 2.8755, 3.1034,
%3.3415, 3.5895, 3.8484, 4.1189, 4.4014

%CSlstm_m_nll_3_ft:
%-2.2872, -0.5987,  0.4855,  1.2288,  1.7826,
%2.2160,  2.5693,  2.8672, 3.1257,  3.3550,
%3.5614,  3.7496,  3.9229,  4.0842,  4.2354,
%4.3779, 4.5128,  4.6409,  4.7631,  4.8820,
%4.9997,  5.1173,  5.2349,  5.3523, 5.4717
%
%0.1058, 0.2117, 0.3233, 0.4439, 0.5697,
%0.6978, 0.8292, 0.9651, 1.1060, 1.2528,
%1.4061, 1.5666, 1.7352, 1.9124, 2.0980,
%2.2925, 2.4963, 2.7099, 2.9335, 3.1675,
%3.4118, 3.6671, 3.9319, 4.2058, 4.4899

%NLL:
%-2.4959, -0.7516,  0.3690,  1.1446,  1.7165,
%2.1619,   2.5230,  2.8268,  3.0904,  3.3241,
%3.5533,  3.7383,  3.9091,  4.0685,  4.2180,
%4.3585, 4.4914,  4.6178,  4.7385,  4.8543,
%4.9658,  5.0735,  5.1777,  5.2793, 5.3783
%RMSE:
%0.1064, 0.2109, 0.3271, 0.4482, 0.5738,
%0.7052, 0.8404, 0.9794, 1.1226, 1.2712,
%1.4257, 1.5877, 1.7582, 1.9381, 2.1267,
%2.3232, 2.5289, 2.7452, 2.9715, 3.2069,
%3.4522, 3.7085, 3.9750, 4.2529, 4.5434

%DumbPredictor_local_att_28
%bias [0.01045633 0.05369709 0.09017866 0.10019906 0.06507623]
%bias (\% RMSE) [2.04381547 4.67443955 4.69984225 3.46073344 1.57623962]
%ADE [[0.06369706 0.30824296]
% [0.12183345 0.78371444]
% [0.1686192  1.36039922]
% [0.20843394 2.08387561]
% [0.24398934 3.00214884]]
%rmse [0.51160824 1.14873851 1.91875931 2.89531284 4.12857472]
%nll [-0.56260347  1.08494605  1.97434634  2.62635983  3.13230919]
%

%DumbPredictor_65_ter
%bias x [[ 0.00452968 -0.01127331]
% [-0.00650934  0.04558599]
% [-0.01217984  0.09266021]
% [-0.01260135  0.12747909]
% [-0.01210313  0.09643459]]
%bias y [[-0.00489422 -0.00224755]
% [-0.00967376  0.05130354]
% [ 0.01033081  0.08558198]
% [ 0.02230204  0.11416008]
% [ 0.02578269  0.08270999]]
%bias dist [[0.04983289 0.05494049]
% [0.06659175 0.1251524 ]
% [0.08607097 0.25349735]
% [0.10527049 0.34948659]
% [0.12184918 0.25884173]]
%bias \% [[ 9.6174681  10.6032057 ]
% [ 5.8789682  11.04892035]
% [ 4.56739348 13.45194753]
% [ 3.71077899 12.31938351]
% [ 3.05157971  6.48240857]]
%ADE [[0.07036931 0.34227107]
% [0.11674017 0.78274485]
% [0.15902441 1.33380843]
% [0.19767509 2.02892873]
% [0.23048341 2.87869137]]
%rmse most probable [0.51814981 1.13271159 1.88446582 2.83688376 3.99298696]
%rmse mean prob [0.5218947  1.13865117 1.88975489 2.84144273 3.99702478]
%nll [0.39706671 1.14138537 1.88634527 2.5478212  3.12638663]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{thebibliography}{99}

\bibitem{Kalman1960} R. E. Kalman, ~A New Approach to Linear Filtering and Prediction Problems,
Transactions of the ASME, Journal of Basic Engineering 1960, vol. 82 pp. 35-45.

\bibitem{Lefevre2014} S. Lef\`evre, D. Vasquez, and C. Laugier, ~A survey on motion prediction and risk assessment for intelligent vehicles,
~published in ROBOMECH journal,vol. 1, no. 1, pp. 1-14, 2014.

\bibitem{Hochreiter1997} S. Hochreiter, and J. Schmidhuber, ~Long short-term memory,
Neural computation 9(8), 1997, pp. 1735-1780

\bibitem{Khosroshahi2016} A. Khosroshahi, E. Ohn-Bar, and M. M. Trivedi, ~Surround vehicles trajectory analysis with recurrent neural networks,
IEEE Conference on Intelligent Transportation Systems, Proceedings, ITSC 2016.

\bibitem{Altche2017} F. Altch\'e, A. de la Fortelle, ~An LSTM network for highway trajectory prediction,
~in IEEE 20th Int. Conf. on Intelligent Transportation Systems, 2017.

\bibitem{Hu2018} Y. Hu, W. Zhan, and M. Tomizuka, ~A Framework for Probabilistic Generic Traffic Scene Prediction,
IEEE 21st Int. Conf. on Intelligent Transportation Systems, Nov. 2018, pp. 2790-2796.

    \bibitem{Houenou2013} A. Houenou, P. Bonnifait, V. Cherfaoui, and Y. Wen, ~Vehicle Trajectory Prediction based on Motion Model and Maneuver Recognition,
~presented in IEEE/RSJ International Conference on Intelligent Robots and Systems, Nov.2013, pp.4363-4369.

%\bibitem{Schreier2016} M. Schreier, V. Willert, and J. Adamy, ~An Integrated Approach to Maneuver-Based Trajectory
%Prediction and Criticality Assessment in Arbitrary Road Environments,
%~in IEEE Transactions on Intelligent Transportation Systems, Oct. 2016, vol.17, no. 10, pp. 2751-2766.

%    \bibitem{Barbier2017} M. Barbier, C. Laugier, O. Simonin, and J. Ibanez-Guzman, ~Classification of Drivers Manoeuvre for Road Intersection Crossing with Synthetic and Real Data,
%~in proceedings of the IEEE intelligent Vehicles Symposium, Jun 2017, pp.7.

\bibitem{Deo2018} N. Deo and M. M. Trivedi, ~Multi-Modal Trajectory Prediction of Surrounding Vehicles with Maneuver based LSTMs,
~published in IEEE Intelligent Vehicles Symposium, 2018.

\bibitem{Shouno2018} O. Shouno, ~Deep learning,driving style,personalization,topological map,unsupervised learning,
IEEE Conference on Intelligent Transportation Systems, Proceedings, ITSC, 2018.

\bibitem{Wissing2018} C. Wissing, T. Nattermann, K. H. Glander, T. Bertram,
~Trajectory Prediction for Safety Critical Maneuvers in Automated Highway Driving,
IEEE Conference on Intelligent Transportation Systems, Proceedings, ITSC 2018.

\bibitem{Kuefler2017} A. Kuefler, J. Morton, T. Wheeler, and M. Kochenderfer, ~Imitating Driver Behavior with Generative Adversarial Networks,
IEEE Intelligent Vehicles Symposium (IV), Los Angeles 2017.

\bibitem{Lenz2017} D. Lenz, F. Diehl, M. Truong Le, and A. Knoll, ~Deep Neural Networks for Markovian Interactive Scene Prediction in Highway Scenarios,
~presented in IEEE Intelligent Vehicles Symposium, Jun. 2017, pp. 685-692.

\bibitem{Vemula2018}  A. Vemula, K. Muelling, and J. Oh, ~Social Attention: Modeling Attention in Human Crowds,
IEEE International Conference on Robotics and Automation (ICRA), pp.4601--4607, 2018,.

\bibitem{Sadeghian2018a} A. Sadeghian, V. Kosaraju, A. Sadeghian, N. Hirose, S. H. Rezatofighi, and S. Savarese,
~SoPhie: An Attentive GAN for Predicting Paths Compliant to Social and Physical Constraints,
ArXiv:1806.01482v2, 2018.

\bibitem{Sadeghian2018b} A. Sadeghian, F. Legros, M. Voisin, R. Vesel, A. Alahi, S. Savarese,
~CAR-Net: Clairvoyant Attentive Recurrent Network, Lecture Notes in Computer Science (LNCS), pp. 162-180, 2018.

\bibitem{Vaswani2017} A. Vaswani, N. Shazeer, N. Parmar, J. Uszkoreit, L. Jones, A. Gomez, L. Kaiser, ~Attention Is All You Need,
~31st Conference on Neural Information Processing Systems (NIPS ), Long Beach, CA, USA, 2017

\bibitem{Messaoud2019} K. Massaoud, I. Yahiaoui, A. Verroust-blondet, and F. Nashashibi, ~Non-local Social Pooling for Vehicle Trajectory Prediction,
in proceedings of the IEEE Intelligent Vehicles Symposium (IV), Jun 2019.

\bibitem{Kingma2015} D. P. Kingma, J. Ba, ~Adam: A method for stochastic optimization,
International Conference on Learning Representations (ICLR), 2015.

    \bibitem{Mercat2019} J. Mercat, D. Beauvois, N. El Zoghby, G. Sandou, and G. Pita Gil,
~Inertial single vehicle trajectory prediction baselines with NGSIM dataset,
ArXiv, 2019

\bibitem{Li2019} X. Li, X. Ying, M. C. Chuah, ~GRIP: Graph-based Interaction-aware Trajectory Prediction,
ArXiv:1907.07792v1, 2019.

\bibitem{Yuan2019} Y. Yuan, K. Kitani, ~Diverse Trajectory Forecasting with Determinantal Point Processes,
arXiv:1907.04967, 2019

\bibitem{Nuscenes2019} H. Caesar, V. Bankiti, A. H. Lang, S. Vora,
          V. E. Liong, Q. Xu A. Krishnan, Y. Pan,
          G. Baldan, O. Beijbom, ~nuScenes: A multimodal dataset for autonomous driving,
        arXiv preprint arXiv:1903.11027, 2019.

\bibitem{Chang2019} M. F. Chang, J. Lambert, P. Sangkloy, J. Singh, S. Bak, A. Hartnett, D. Wang, P. Carr, S. Lucey, D. Ramanan, J. Hays,
~Argoverse: 3D Tracking and Forecasting With Rich Maps,
IEEE Conference on Computer Vision and Pattern Recognition (CVPR), 2019.

%\bibitem{Scholler2019} C. Sch{\"{o}}ller, V. Aravantinos, F. Lay and, A. Knoll, ~The Simpler the Better: Constant Velocity for Pedestrian Motion Prediction,
%ArXiv 2019.

%\bibitem{Xu2019} Y. Xu, T. Zhao C. Baker, Y. Zhao, and Y. Wu, ~Learning Trajectory Prediction with Continuous Inverse Optimal Control via Langevin Sampling of Energy-Based Models,
%~in Proc. of the 18th International Conference on Autonomous Agents and Multiagent Systems (AAMAS), May 2019

%\bibitem{Ju2019}, C. Ju, Z. Wang, C. Long, X. Zhang, G. Cong, and D. E. Chang,
%~Interaction-aware Kalman Neural Networks for Trajectory Prediction,
%arXiv, 2019.

%\bibitem{Barbier2017} M. Barbier, C. Laugier, O. Simonin, and J. Ibanez-Guzman, ~Classification of Drivers Manoeuvre for Road Intersection Crossing with Synthetic and Real Data,
%~in proceedings of the IEEE intelligent Vehicles Symposium, Jun 2017, pp.7.

%\bibitem{Bian2018} J. Bian, D. Tian, Y. Tang, and D. Tao, ~A survey on trajectory clustering analysis,
%~arXiv, 2018.
%\bibitem{Buehler2009} M. Buehler, K. Iagnemma and S. Singh, ~The DARPA Urban Challenge, ~Springer Tracts in Advanced Robotics, 2009 v. 56.
%\bibitem{Chen2015} C. Chen, A. Seff, A. Kornhauser, and J. Xiao, ~DeepDriving: Learning Affordance for Direct Perception in
%Autonomous Driving, ~in the proceedings of IEEE Int. Conf. on Computer Vision, 2015, pp. 2722-2730.
%\bibitem{Choi2006} P. P. Choi, and M. Hebert, ~Learning and Predicting Moving Object Trajectory: a piecewise trajectory segment approach,
%~Carnegie Mellon University, Aug. 2006.

%\bibitem{He2012} L. He, C. Zong, and C. Wang, ~Driving intention recognition and behaviour prediction based on a double-layer hidden Markov model,
%~published in Journal of Zhejiang University-SCIENCE C (Computers \& Electronics), 2012, 13(3):208-217.
%\bibitem{Houenou2013} A. Houenou, P. Bonnifait, V. Cherfaoui, and Y. Wen, ~Vehicle Trajectory Prediction based on Motion Model and Maneuver Recognition,
%~presented in IEEE/RSJ International Conference on Intelligent Robots and Systems, Nov.2013, pp.4363-4369.
%\bibitem{Hu2018} Y. Hu, W. Zhan, and M. Tomizuka, ~A Framework for Probabilistic Generic Traffic Scene Prediction,
%IEEE 21st Int. Conf. on Intelligent Transportation Systems, Nov. 2018, pp. 2790-2796.
%\bibitem{Krajewski2018} R. Krajewski, J. Bock, L. Kloeker, and L. Eckstein, ~The highD Dataset: A Drone Dataset
%of Naturalistic Vehicle Trajectories on German Highways for Validation of Highly Automated Driving Systems,
%~presented in IEEE International Conference on Intelligent Transportation Systems, 2018.
%\bibitem{Kim2017} B. Kim, C. M. Kang, J. Kim, S. H. Lee, C. C. Chung, and J. W. Choi,
%~Probabilistic Vehicle Trajectory Prediction over Occupancy Grid Map via Recurrent Neural Network,
%~in IEEE 20th Int. Conf. on Intelligent Transporatition Systems, Oct. 2017.

%\bibitem{Lenz2017} D. Lenz, F. Diehl, M. Truong Le, and A. Knoll, ~Deep Neural Networks for Markovian Interactive Scene Prediction in Highway Scenarios,
%~presented in IEEE Intelligent Vehicles Symposium, Jun. 2017, pp. 685-692.
%\bibitem{Petrich2014} D. Petrich, T. Dang, G. Breuel, and C. Stiller, ~Assessing Map-Based Maneuver Hypotheses using Probabilistic Methods and Evidence Theory,
%~presented in proceedings IEEE International Conference on Intelligent Transportation Systems, 2014, pp. 995-1002.
%\bibitem{Negre2014}  A. N\`egre, L. Rummelhard, and C. Laugier, ~Hybrid Sampling Bayesian Occupancy Filter.
%~presented in IEEE Intelligent Vehicles Symposium, Jun. 2014, pp. 1307-1312.
%\bibitem{Schreiber2018} M. Schreiber, S. Hoermann, and K. Dietmayer, ~Long-Term Occupancy Grid Prediction Using
%Recurrent Neural Networks, arXiv:1809.03782, 2018.
%\bibitem{Schreier2016-2} M. Schreier, V. Willert, and J. Adamy, ~An Integrated Approach to Maneuver-Based Trajectory
%Prediction and Criticality Assessment in Arbitrary Road Environments,
%~in IEEE Transactions on Intelligent Transportation Systems, Oct. 2016, vol.17, no. 10, pp. 2751-2766.
%\bibitem{Schreier2016} M. Schreier, V. Willert, and J. Adamy, ~Compact Representation of Dynamic Driving Environments for ADAS by Parametric Free Space and Dynamic Object Maps,
%~in IEEE Transactions on Intelligent Transportation Systems, Feb. 2016, vol.17, no.2, pp. 367-383.
%%\bibitem{Schreier2017} M. Schreier, ~Bayesian environment representation, prediction, and criticality assessment for driver assistance systems,
%%~published at-Automatisierungstechnik, 2017, vol. 65, no 2, p. 151-152.
%\bibitem{Schulz2018} J. Schulz, C. Hubmann, J. Löchner, and D. Burschka, ~Interaction-Aware Probabilistic Behavior Prediction
%in Urban Environments, ~at IEEE/RSJ Int. Conf. on Intelligent Robots and Systems, 2018, pp. 3999-4006.
%\bibitem{Tang2017} J. Tang, F. Liu, W. Zhang, R. Ke, and Y. Zou, ~Lane-changes pre diction base d on adaptive fuzzy neural network,
%~published in Expert Systems with Applications 91, 2018, pp. 452-463.
%\bibitem{Weiherer2013} T. Weiherer, S. Bouzouraa, and U. Hofmann, ~An interval based representation of occupancy information for driver assistance systems,
%~in proceedings of the 16th Int. IEEE Conference on Intelligent Transportation Systems, Oct. 6-9, 2013, pp. 21-27.
%\bibitem{Wiest2016} J. Wiest, ~Statistical Long-Term Motion Prediction, ~Ph.D. dissertation, ~ulm university, Germany, 2016.
%\bibitem{Woo2017} H. Woo, Y. Ji, H. Kono, Y. Tamura, Y. Kuroda, T. Sugano, Y. Yamamoto, A. Yamashita, and H. Asama,
%~Lane-Change Detection Based on Vehicle-Trajectory Prediction, ~published in IEEE Robotics and Automation Letters, 2017, vol. 2, pp. 1109-1116.
%\bibitem{Shalev-Shwartz2016} S. Shalev-Shwartz, S. Shammah, A. Shashua, ~Safe, Multi-Agent, Reinforcement Learning for Autonomous Driving,
%~arXiv, 2016.
%\bibitem{Shreier2014} M. Schreier, V. Willert, J. Adamy, ~Bayesian, Maneuver-Based, Long-Term Trajectory Prediction and Criticality Assessment for Driver Assistance Systems,
%    ~presented in IEEE International Conference on Intelligent Transportation Systems, 2014.
%\bibitem{Dimarogonas2012} D. V. Dimarogonas, E. Frazzoli, and K. H. Johansson, ~Distributed Event-Triggered Control for Multi-Agent Systems,
%~published in IEEE Transactions on Automatic Control, 2012.

\end{thebibliography}

\end{document}

### 2019-09-02 neurips_2019_ml4ad.tex
Add neurips template

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2019

% ready for submission
\usepackage{neurips_2019_ml4ad}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
% \usepackage[preprint]{neurips_2019_ml4ad}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{neurips_2019_ml4ad}

% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{neurips_2019_ml4ad}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\title{Formatting Instructions for NeurIPS 2019 Workshop on Machine Learning for Autonomous Driving}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

%Happy Horizontal People Transporter
%From the Sirius Cybernetics Corporation with a Genuine People Personality

\author{%
  Happy Horizontal People Transporter\thanks{Use footnote for providing further information
  about author (webpage, alternative address)---\emph{not} for acknowledging funding agencies.} \\
  Genuine People Personality Division\\
  Sirius Cybernetics Corporation\\
  \texttt{marvin@siriuscybernetics.com} \\
  % examples of more authors
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \AND
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}

\begin{document}

\maketitle

\begin{abstract}
  The abstract paragraph should be indented \nicefrac{1}{2}~inch (3~picas) on
  both the left- and right-hand margins. Use 10~point type, with a vertical
  spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
  bold, and in point size 12. Two line spaces precede the abstract. The abstract
  must be limited to one paragraph.
\end{abstract}

\section{Submission of papers to ML4AD 2019}

The NeurIPS 2019 workshop on Machine Learning for Autonomous Driving (ML4AD)
requires electronic submissions.  The electronic submission site is
\begin{center}
  \url{https://cmt3.research.microsoft.com/ML4AD2019}
\end{center}

Please read the instructions below carefully and follow them faithfully.

\subsection{Style}

Papers to be submitted to ML4AD 2019 must be prepared according to the
instructions presented here. Papers may be four or eight pages long,
including figures. Additional pages \emph{containing only acknowledgments and/or
  cited references} are allowed.

\subsection{Retrieval of style files}

The style files for ML4AD and other conference information are available on
the World Wide Web at
\begin{center}
  \url{https://ml4ad.github.io/#submit}
\end{center}
The file \verb+neurips_2019_ml4ad.pdf+ contains these instructions and illustrates the
various formatting requirements your NeurIPS paper must satisfy.

The only supported style file for ML4AD is \verb+neurips_2019_ml4ad.sty+,
rewritten for \LaTeXe{}.  \textbf{Previous style files for \LaTeX{} 2.09,
  Microsoft Word, and RTF are no longer supported!}

The \LaTeX{} style file contains three optional arguments: \verb+final+, which
creates a camera-ready copy, \verb+preprint+, which creates a preprint for
submission to, e.g., arXiv, and \verb+nonatbib+, which will not load the
\verb+natbib+ package for you in case of package clash.

\paragraph{Preprint option}
If you wish to post a preprint of your work online, e.g., on arXiv, using the
NeurIPS style, please use the \verb+preprint+ option. This will create a
nonanonymized version of your work with the text ``Preprint. Work in progress.''
in the footer.

At submission time, please omit the \verb+final+ and \verb+preprint+
options. This will anonymize your submission and add line numbers to aid
review. Please do \emph{not} refer to these line numbers in your paper as they
will be removed during generation of camera-ready copies.

The file \verb+neurips_2019_ml4ad.tex+ may be used as a ``shell'' for writing your
paper. All you have to do is replace the author, title, abstract, and text of
the paper with your own.

The formatting instructions contained in these style files are summarized in
Sections \ref{gen_inst}, \ref{headings}, and \ref{others} below.

\section{General formatting instructions}
\label{gen_inst}

The text must be confined within a rectangle 5.5~inches (33~picas) wide and
9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).  Use 10~point
type with a vertical spacing (leading) of 11~points.  Times New Roman is the
preferred typeface throughout, and will be selected for you by default.
Paragraphs are separated by \nicefrac{1}{2}~line space (5.5 points), with no
indentation.

The paper title should be 17~point, initial caps/lower case, bold, centered
between two horizontal rules. The top rule should be 4~points thick and the
bottom rule should be 1~point thick. Allow \nicefrac{1}{4}~inch space above and
below the title to rules. All pages should start at 1~inch (6~picas) from the
top of the page.

For the final version, authors' names are set in boldface, and each name is
centered above the corresponding address. The lead author's name is to be listed
first (left-most), and the co-authors' names (if different address) are set to
follow. If there is only one co-author, list both author and co-author side by
side.

Please pay special attention to the instructions in Section \ref{others}
regarding figures, tables, acknowledgments, and references.

\section{Headings: first level}
\label{headings}

All headings should be lower case (except for first word and proper nouns),
flush left, and bold.

First-level headings should be in 12-point type.

\subsection{Headings: second level}

Second-level headings should be in 10-point type.

\subsubsection{Headings: third level}

Third-level headings should be in 10-point type.

\paragraph{Paragraphs}

There is also a \verb+\paragraph+ command available, which sets the heading in
bold, flush left, and inline with the text, with the heading followed by 1\,em
of space.

\section{Citations, figures, tables, references}
\label{others}

These instructions apply to everyone.

\subsection{Citations within the text}

The \verb+natbib+ package will be loaded for you by default.  Citations may be
author/year or numeric, as long as you maintain internal consistency.  As to the
format of the references themselves, any style is acceptable as long as it is
used consistently.

The documentation for \verb+natbib+ may be found at
\begin{center}
  \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}
\end{center}
Of note is the command \verb+\citet+, which produces citations appropriate for
use in inline text.  For example,
\begin{verbatim}
   \citet{hasselmo} investigated\dots
\end{verbatim}
produces
\begin{quote}
  Hasselmo, et al.\ (1995) investigated\dots
\end{quote}

If you wish to load the \verb+natbib+ package with options, you may add the
following before loading the \verb+neurips_2019+ package:
\begin{verbatim}
   \PassOptionsToPackage{options}{natbib}
\end{verbatim}

If \verb+natbib+ clashes with another package you load, you can add the optional
argument \verb+nonatbib+ when loading the style file:
\begin{verbatim}
   \usepackage[nonatbib]{neurips_2019_ml4ad}
\end{verbatim}

As submission is double blind, refer to your own published work in the third
person. That is, use ``In the previous work of Jones et al.\ [4],'' not ``In our
previous work [4].'' If you cite your other papers that are not widely available
(e.g., a journal paper under review), use anonymous author names in the
citation, e.g., an author of the form ``A.\ Anonymous.''

\subsection{Footnotes}

Footnotes should be used sparingly.  If you do require a footnote, indicate
footnotes with a number\footnote{Sample of the first footnote.} in the
text. Place the footnotes at the bottom of the page on which they appear.
Precede the footnote with a horizontal rule of 2~inches (12~picas).

Note that footnotes are properly typeset \emph{after} punctuation
marks.\footnote{As in this example.}

\subsection{Figures}

\begin{figure}
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
\end{figure}

All artwork must be neat, clean, and legible. Lines should be dark enough for
purposes of reproduction. The figure number and caption always appear after the
figure. Place one line space before the figure caption and one line space after
the figure. The figure caption should be lower case (except for first word and
proper nouns); figures are numbered consecutively.

You may use color figures.  However, it is best for the figure captions and the
paper body to be legible if the paper is printed in either black/white or in
color.

\subsection{Tables}

All tables must be centered, neat, clean and legible.  The table number and
title always appear before the table.  See Table~\ref{sample-table}.

Place one line space before the table title, one line space after the
table title, and one line space after the table. The table title must
be lower case (except for first word and proper nouns); tables are
numbered consecutively.

Note that publication-quality tables \emph{do not contain vertical rules.} We
strongly suggest the use of the \verb+booktabs+ package, which allows for
typesetting high-quality, professional tables:
\begin{center}
  \url{https://www.ctan.org/pkg/booktabs}
\end{center}
This package was used to typeset Table~\ref{sample-table}.

\begin{table}
  \caption{Sample table title}
  \label{sample-table}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule(r){1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
\end{table}

\section{Final instructions}

Do not change any aspects of the formatting parameters in the style files.  In
particular, do not modify the width or length of the rectangle the text should
fit into, and do not change font sizes (except perhaps in the
\textbf{References} section; see below). Please note that pages should be
numbered.

\section{Preparing PDF files}

Please prepare submission files with paper size ``US Letter,'' and not, for
example, ``A4.''

Fonts were the main cause of problems in the past years. Your PDF file must only
contain Type 1 or Embedded TrueType fonts. Here are a few instructions to
achieve this.

\begin{itemize}

\item You should directly generate PDF files using \verb+pdflatex+.

\item You can check which fonts a PDF files uses.  In Acrobat Reader, select the
  menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
  also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
  available out-of-the-box on most Linux machines.

\item The IEEE has recommendations for generating PDF files whose fonts are also
  acceptable for NeurIPS. Please see
  \url{http://www.emfield.org/icuwb2010/downloads/IEEE-PDF-SpecV32.pdf}

\item \verb+xfig+ "patterned" shapes are implemented with bitmap fonts.  Use
  "solid" shapes instead.

\item The \verb+\bbold+ package almost always uses bitmap fonts.  You should use
  the equivalent AMS Fonts:
\begin{verbatim}
   \usepackage{amsfonts}
\end{verbatim}
followed by, e.g., \verb+\mathbb{R}+, \verb+\mathbb{N}+, or \verb+\mathbb{C}+
for $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.  You can also use the following
workaround for reals, natural and complex:
\begin{verbatim}
   \newcommand{\RR}{I\!\!R} %real numbers
   \newcommand{\Nat}{I\!\!N} %natural numbers
   \newcommand{\CC}{I\!\!\!\!C} %complex numbers
\end{verbatim}
Note that \verb+amsfonts+ is automatically loaded by the \verb+amssymb+ package.

\end{itemize}

If your file contains type 3 fonts or non embedded TrueType fonts, we will ask
you to fix it.

\subsection{Margins in \LaTeX{}}

Most of the margin problems come from figures positioned by hand using
\verb+\special+ or other commands. We suggest using the command
\verb+\includegraphics+ from the \verb+graphicx+ package. Always specify the
figure width as a multiple of the line width as in the example below:
\begin{verbatim}
   \usepackage[pdftex]{graphicx} ...
   \includegraphics[width=0.8\linewidth]{myfile.pdf}
\end{verbatim}
See Section 4.4 in the graphics bundle documentation
(\url{http://mirrors.ctan.org/macros/latex/required/graphics/grfguide.pdf})

A number of width problems arise when \LaTeX{} cannot properly hyphenate a
line. Please give LaTeX hyphenation hints using the \verb+\-+ command when
necessary.

\subsubsection*{Acknowledgments}

Use unnumbered third level headings for the acknowledgments. All acknowledgments
go at the end of the paper. Do not include acknowledgments in the anonymized
submission, only in the final paper.

\section*{References}

References follow the acknowledgments. Use unnumbered first-level heading for
the references. Any choice of citation style is acceptable as long as you are
consistent. It is permissible to reduce the font size to \verb+small+ (9 point)
when listing the references.
\medskip

\small

[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms for
connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and T.K.\ Leen
(eds.), {\it Advances in Neural Information Processing Systems 7},
pp.\ 609--616. Cambridge, MA: MIT Press.

[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS: Exploring
  Realistic Neural Models with the GEneral NEural SImulation System.}  New York:
TELOS/Springer--Verlag.

[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of learning and
recall at excitatory recurrent synapses and cholinergic modulation in rat
hippocampal region CA3. {\it Journal of Neuroscience} {\bf 15}(7):5249-5262.

\end{document}

### 2019-09-02 main.tex
Add skeletton

--- 2019-09-02 neurips_2019_ml4ad.tex
@@
 % ready for submission
-\usepackage{neurips_2019_ml4ad}
+%\usepackage{neurips_2019_ml4ad}
 
@@
 % [preprint] option:
-% \usepackage[preprint]{neurips_2019_ml4ad}
+ \usepackage[preprint]{neurips_2019_ml4ad}
 
@@
 
-\title{Formatting Instructions for NeurIPS 2019 Workshop on Machine Learning for Autonomous Driving}
+\title{Social Attention for Autonomous Decision-Making in Dense Traffic}
 
@@
 \author{%
-  Happy Horizontal People Transporter\thanks{Use footnote for providing further information
-  about author (webpage, alternative address)---\emph{not} for acknowledging funding agencies.} \\
-  Genuine People Personality Division\\
-  Sirius Cybernetics Corporation\\
-  \texttt{marvin@siriuscybernetics.com} \\
+  Edouard Leurent\thanks{Equal contribution.} \\
+  Inria, SequeL team\\
+  Renault Group\\
+  Paris, France\\
+  \texttt{edouard.leurent@renault.com} \\
   % examples of more authors
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \AND
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
+   \And
+  Jean Mercat$^*$ \\
+  Laboratoire des signaux et des syst\`emes,
+  Centrale-Sup\'elec\\
+  Renault Group\\
+  Paris, France\\
+  \texttt{jean.mercat@renault.com} \\
 }
@@
 
-\section{Submission of papers to ML4AD 2019}
+\section{Introduction}
 
-The NeurIPS 2019 workshop on Machine Learning for Autonomous Driving (ML4AD)
-requires electronic submissions.  The electronic submission site is
-\begin{center}
-  \url{https://cmt3.research.microsoft.com/ML4AD2019}
-\end{center}
+Automated Decision Making
+Modular pipeline: perception, high level representation, decision: behaviour planning, trajectory, control
+large litterature on perception
+trajectory optimization and control is well understood
 
-Please read the instructions below carefully and follow them faithfully.
+behaviour planning by rule-based systems, fsm
+but behaviour planning is difficult, complex scenes, in particular interactions with other drivers difficult to model explicitly.
+ -> leverage (deep) learning method
+Learning from demonstration: nice results for lane keeping, not so much for interaction, also require costly demonstration .
 
-\subsection{Style}
+Conversely, leverage simulation:
+Reinforcement learning:
+Framing as an MDP, choose a good representation.
+List of vehicles with their coordinates 
+but variable size, + combinatorial representation
+Occupancy grid / images : fixed size, tradeoff:  low accuracy vs exploding size
 
-Papers to be submitted to ML4AD 2019 must be prepared according to the
-instructions presented here. Papers may be four or eight pages long,
-including figures. Additional pages \emph{containing only acknowledgments and/or
-  cited references} are allowed.
+\section{Background and Related Work}
 
-\subsection{Retrieval of style files}
+\subsection{Model-Free Deep Reinforcement Learning}
 
-The style files for ML4AD and other conference information are available on
-the World Wide Web at
-\begin{center}
-  \url{https://ml4ad.github.io/#submit}
-\end{center}
-The file \verb+neurips_2019_ml4ad.pdf+ contains these instructions and illustrates the
-various formatting requirements your NeurIPS paper must satisfy.
+- MDP
+- Value function
+- DQN
 
-The only supported style file for ML4AD is \verb+neurips_2019_ml4ad.sty+,
-rewritten for \LaTeXe{}.  \textbf{Previous style files for \LaTeX{} 2.09,
-  Microsoft Word, and RTF are no longer supported!}
+\subsection{State Representation for Social Interactions}
+In the state:
+Kinematics
+Not fixed size
+Occupancy grid
 
-The \LaTeX{} style file contains three optional arguments: \verb+final+, which
-creates a camera-ready copy, \verb+preprint+, which creates a preprint for
-submission to, e.g., arXiv, and \verb+nonatbib+, which will not load the
-\verb+natbib+ package for you in case of package clash.
+In the model:
 
-\paragraph{Preprint option}
-If you wish to post a preprint of your work online, e.g., on arXiv, using the
-NeurIPS style, please use the \verb+preprint+ option. This will create a
-nonanonymized version of your work with the text ``Preprint. Work in progress.''
-in the footer.
+Weight sharing ?
 
-At submission time, please omit the \verb+final+ and \verb+preprint+
-options. This will anonymize your submission and add line numbers to aid
-review. Please do \emph{not} refer to these line numbers in your paper as they
-will be removed during generation of camera-ready copies.
+\section{Model Architecture}
 
-The file \verb+neurips_2019_ml4ad.tex+ may be used as a ``shell'' for writing your
-paper. All you have to do is replace the author, title, abstract, and text of
-the paper with your own.
+Motivation: 
+- when fed with a complex scene description, the model should be able to filter information and consider only what is relevant for decision: ie \emph{pay attention} to vehicles that are close, and conflict with the planned route
+- the model should be able to access variable number of vehicles
+- the model should be invariant to the ordering of vehicles
 
-The formatting instructions contained in these style files are summarized in
-Sections \ref{gen_inst}, \ref{headings}, and \ref{others} below.
+\section{Experiments}
+\subsection{The intersection task}
 
-\section{General formatting instructions}
-\label{gen_inst}
+We consider a situation where interactions between vehicles is important: crossing an unsignalized intersection.
 
-The text must be confined within a rectangle 5.5~inches (33~picas) wide and
-9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).  Use 10~point
-type with a vertical spacing (leading) of 11~points.  Times New Roman is the
-preferred typeface throughout, and will be selected for you by default.
-Paragraphs are separated by \nicefrac{1}{2}~line space (5.5 points), with no
-indentation.
+To that end, we use the highway-env environment and propose a novel intersection task.
 
-The paper title should be 17~point, initial caps/lower case, bold, centered
-between two horizontal rules. The top rule should be 4~points thick and the
-bottom rule should be 1~point thick. Allow \nicefrac{1}{4}~inch space above and
-below the title to rules. All pages should start at 1~inch (6~picas) from the
-top of the page.
+- The intersection: cross (North/South/East/West), single lane or 2 lane ?
+- The other vehicles: are spawned randomly, and plan a route to a random direction.
+Follow a Kinematic Bicycle Model, lane keeping with a low-level steering controller, longitudinal control with IDM
 
-For the final version, authors' names are set in boldface, and each name is
-centered above the corresponding address. The lead author's name is to be listed
-first (left-most), and the co-authors' names (if different address) are set to
-follow. If there is only one co-author, list both author and co-author side by
-side.
+\subsection{The baselines}
+\subsection{Performances}
+\subsection{Attention interpretation}
 
-Please pay special attention to the instructions in Section \ref{others}
-regarding figures, tables, acknowledgments, and references.
+Additional videos are available at \url{http://url.github.io}
 
-\section{Headings: first level}
-\label{headings}
-
-All headings should be lower case (except for first word and proper nouns),
-flush left, and bold.
-
-First-level headings should be in 12-point type.
-
-\subsection{Headings: second level}
-
-Second-level headings should be in 10-point type.
-
-\subsubsection{Headings: third level}
-
-Third-level headings should be in 10-point type.
-
-\paragraph{Paragraphs}
-
-There is also a \verb+\paragraph+ command available, which sets the heading in
-bold, flush left, and inline with the text, with the heading followed by 1\,em
-of space.
-
-\section{Citations, figures, tables, references}
-\label{others}
-
-These instructions apply to everyone.
-
-\subsection{Citations within the text}
-
-The \verb+natbib+ package will be loaded for you by default.  Citations may be
-author/year or numeric, as long as you maintain internal consistency.  As to the
-format of the references themselves, any style is acceptable as long as it is
-used consistently.
-
-The documentation for \verb+natbib+ may be found at
-\begin{center}
-  \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}
-\end{center}
-Of note is the command \verb+\citet+, which produces citations appropriate for
-use in inline text.  For example,
-\begin{verbatim}
-   \citet{hasselmo} investigated\dots
-\end{verbatim}
-produces
-\begin{quote}
-  Hasselmo, et al.\ (1995) investigated\dots
-\end{quote}
-
-If you wish to load the \verb+natbib+ package with options, you may add the
-following before loading the \verb+neurips_2019+ package:
-\begin{verbatim}
-   \PassOptionsToPackage{options}{natbib}
-\end{verbatim}
-
-If \verb+natbib+ clashes with another package you load, you can add the optional
-argument \verb+nonatbib+ when loading the style file:
-\begin{verbatim}
-   \usepackage[nonatbib]{neurips_2019_ml4ad}
-\end{verbatim}
-
-As submission is double blind, refer to your own published work in the third
-person. That is, use ``In the previous work of Jones et al.\ [4],'' not ``In our
-previous work [4].'' If you cite your other papers that are not widely available
-(e.g., a journal paper under review), use anonymous author names in the
-citation, e.g., an author of the form ``A.\ Anonymous.''
-
-\subsection{Footnotes}
-
-Footnotes should be used sparingly.  If you do require a footnote, indicate
-footnotes with a number\footnote{Sample of the first footnote.} in the
-text. Place the footnotes at the bottom of the page on which they appear.
-Precede the footnote with a horizontal rule of 2~inches (12~picas).
-
-Note that footnotes are properly typeset \emph{after} punctuation
-marks.\footnote{As in this example.}
-
-\subsection{Figures}
-
-\begin{figure}
-  \centering
-  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
-  \caption{Sample figure caption.}
-\end{figure}
-
-All artwork must be neat, clean, and legible. Lines should be dark enough for
-purposes of reproduction. The figure number and caption always appear after the
-figure. Place one line space before the figure caption and one line space after
-the figure. The figure caption should be lower case (except for first word and
-proper nouns); figures are numbered consecutively.
-
-You may use color figures.  However, it is best for the figure captions and the
-paper body to be legible if the paper is printed in either black/white or in
-color.
-
-\subsection{Tables}
-
-All tables must be centered, neat, clean and legible.  The table number and
-title always appear before the table.  See Table~\ref{sample-table}.
-
-Place one line space before the table title, one line space after the
-table title, and one line space after the table. The table title must
-be lower case (except for first word and proper nouns); tables are
-numbered consecutively.
-
-Note that publication-quality tables \emph{do not contain vertical rules.} We
-strongly suggest the use of the \verb+booktabs+ package, which allows for
-typesetting high-quality, professional tables:
-\begin{center}
-  \url{https://www.ctan.org/pkg/booktabs}
-\end{center}
-This package was used to typeset Table~\ref{sample-table}.
-
-\begin{table}
-  \caption{Sample table title}
-  \label{sample-table}
-  \centering
-  \begin{tabular}{lll}
-    \toprule
-    \multicolumn{2}{c}{Part}                   \\
-    \cmidrule(r){1-2}
-    Name     & Description     & Size ($\mu$m) \\
-    \midrule
-    Dendrite & Input terminal  & $\sim$100     \\
-    Axon     & Output terminal & $\sim$10      \\
-    Soma     & Cell body       & up to $10^6$  \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-\section{Final instructions}
-
-Do not change any aspects of the formatting parameters in the style files.  In
-particular, do not modify the width or length of the rectangle the text should
-fit into, and do not change font sizes (except perhaps in the
-\textbf{References} section; see below). Please note that pages should be
-numbered.
-
-\section{Preparing PDF files}
-
-Please prepare submission files with paper size ``US Letter,'' and not, for
-example, ``A4.''
-
-Fonts were the main cause of problems in the past years. Your PDF file must only
-contain Type 1 or Embedded TrueType fonts. Here are a few instructions to
-achieve this.
-
-\begin{itemize}
-
-\item You should directly generate PDF files using \verb+pdflatex+.
-
-\item You can check which fonts a PDF files uses.  In Acrobat Reader, select the
-  menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
-  also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
-  available out-of-the-box on most Linux machines.
-
-\item The IEEE has recommendations for generating PDF files whose fonts are also
-  acceptable for NeurIPS. Please see
-  \url{http://www.emfield.org/icuwb2010/downloads/IEEE-PDF-SpecV32.pdf}
-
-\item \verb+xfig+ "patterned" shapes are implemented with bitmap fonts.  Use
-  "solid" shapes instead.
-
-\item The \verb+\bbold+ package almost always uses bitmap fonts.  You should use
-  the equivalent AMS Fonts:
-\begin{verbatim}
-   \usepackage{amsfonts}
-\end{verbatim}
-followed by, e.g., \verb+\mathbb{R}+, \verb+\mathbb{N}+, or \verb+\mathbb{C}+
-for $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.  You can also use the following
-workaround for reals, natural and complex:
-\begin{verbatim}
-   \newcommand{\RR}{I\!\!R} %real numbers
-   \newcommand{\Nat}{I\!\!N} %natural numbers
-   \newcommand{\CC}{I\!\!\!\!C} %complex numbers
-\end{verbatim}
-Note that \verb+amsfonts+ is automatically loaded by the \verb+amssymb+ package.
-
-\end{itemize}
-
-If your file contains type 3 fonts or non embedded TrueType fonts, we will ask
-you to fix it.
-
-\subsection{Margins in \LaTeX{}}
-
-Most of the margin problems come from figures positioned by hand using
-\verb+\special+ or other commands. We suggest using the command
-\verb+\includegraphics+ from the \verb+graphicx+ package. Always specify the
-figure width as a multiple of the line width as in the example below:
-\begin{verbatim}
-   \usepackage[pdftex]{graphicx} ...
-   \includegraphics[width=0.8\linewidth]{myfile.pdf}
-\end{verbatim}
-See Section 4.4 in the graphics bundle documentation
-(\url{http://mirrors.ctan.org/macros/latex/required/graphics/grfguide.pdf})
-
-A number of width problems arise when \LaTeX{} cannot properly hyphenate a
-line. Please give LaTeX hyphenation hints using the \verb+\-+ command when
-necessary.
+\section{Conclusion}
 
@@
 Use unnumbered third level headings for the acknowledgments. All acknowledgments
-go at the end of the paper. Do not include acknowledgments in the anonymized
+Go at the end of the paper. Do not include acknowledgments in the anonymized
 submission, only in the final paper.
@@
 
-References follow the acknowledgments. Use unnumbered first-level heading for
-the references. Any choice of citation style is acceptable as long as you are
-consistent. It is permissible to reduce the font size to \verb+small+ (9 point)
-when listing the references.
-\medskip
-
-\small
-
-[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms for
-connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and T.K.\ Leen
-(eds.), {\it Advances in Neural Information Processing Systems 7},
-pp.\ 609--616. Cambridge, MA: MIT Press.
-
-[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS: Exploring
-  Realistic Neural Models with the GEneral NEural SImulation System.}  New York:
-TELOS/Springer--Verlag.
-
-[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of learning and
-recall at excitatory recurrent synapses and cholinergic modulation in rat
-hippocampal region CA3. {\it Journal of Neuroscience} {\bf 15}(7):5249-5262.
-
 \end{document}

### 2019-09-03 main.tex
Edit graphs

@@
 \usepackage{microtype}      % microtypography
+
+% Custom packages
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{graphicx}
+\usepackage{tikz}
+\usepackage{float}
+\usetikzlibrary{arrows,automata}
+\usepackage{pgfplots}
+\usetikzlibrary{intersections}
 
@@
 \begin{document}
+
+	\tikzset{
+		state/.style={
+			rectangle,
+			draw=black, very thick,
+			minimum height=2em,
+			inner sep=2pt,
+			text centered,
+		},
+		name plot/.style={every path/.style={name path global=#1}}
+	}
+
+	\pgfmathdeclarefunction{dnorm}{2}{%
+		\pgfmathparse{1/(#2*sqrt(2*pi))*exp(-((x-#1)^2)/(2*#2^2))}%
+	}
 
@@
 Motivation: 
-- when fed with a complex scene description, the model should be able to filter information and consider only what is relevant for decision: ie \emph{pay attention} to vehicles that are close, and conflict with the planned route
-- the model should be able to access variable number of vehicles
-- the model should be invariant to the ordering of vehicles
+\begin{enumerate}
+	\item Out of a complex scene description, the model should be able to filter information and consider only what is relevant for decision.
+In other words, the agent should \emph{pay attention} to vehicles that are close or conflict with the planned route;
+	\item The model should be able to admit a variable number of vehicles as inputs;
+	\item The model should be invariant to the ordering of vehicles.
+\end{enumerate}
+
+\begin{figure}[ht]
+	\centering
+	\begin{tikzpicture}
+	\node(X1){ego};
+	\node[below of=X1, node distance=0.7cm](X2){vehicle$_{1}$};
+	\node[below of=X2, node distance=0.6cm](X3){$\vdots$};
+	\node[below of=X3, node distance=0.7cm](X4){vehicle$_{n}$};
+
+	\node[draw, right of=X1, node distance=1.8cm, rectangle](ENC1){Encoder};
+	\node[draw, right of=X2, node distance=1.8cm, rectangle](ENC2){Encoder};
+	\node[below of=ENC2, node distance=0.6cm](ENC3){$\vdots$};
+	\node[draw, right of=X4, node distance=1.8cm, rectangle](ENC4){Encoder};
+
+	\path (X1) edge (ENC1);
+	\path (X2) edge (ENC2);
+	\path (X4) edge (ENC4);
+
+	\node[draw, rectangle, right of=ENC1, node distance=1.8cm, below=-0.2cm](TRANS1){\rotatebox{90}{ Ego-attention }};
+
+	\draw (ENC1.east) -| (TRANS1.west);
+	\draw (ENC2.east) -| (TRANS1.west);
+	\draw (ENC4.east) -| (TRANS1.west);
+
+	\node[draw, right of=ENC1, node distance=3.6cm, rectangle](DEC1){Decoder};
+
+	\draw (TRANS1.east) |- (DEC1.west);
+
+	\node[right of=DEC1, node distance=2.3cm](Y1){action values};
+
+	\draw (DEC1.east) -- (Y1.west);
+
+	%        \draw[decorate,decoration={brace,mirror}] (X1.north west) -- node[left=10pt] {$X$} (X3.south west);
+	\end{tikzpicture}
+	\caption{Block diagram of our model architecture.}
+	\label{sch_whole_model}
+
+\end{figure}
+
+The multi-head self-attention layers allow vehicle interactions while keeping independence from their number
+and ordering.
+This mechanism is described in~\cite{Vaswani2017} where it is applied on sentence translation.
+In this section we explain its use for vehicle interactions.
+The computations made by each attention head is represented on figure~\ref{fig:ego-attention} and are detailed below.
+
+\begin{figure}[ht]
+	\centering
+	\begin{tikzpicture}[scale=1, every node/.style={scale=1}]
+	\node(X1){ego encoding};
+	\node[below of=X1, node distance=2cm](X2){$\vdots$};
+	\node[below of=X2, node distance=2cm](X3){vehicle$_{n}$ encoding};
+
+	\coordinate[right of= X1, node distance=2cm](X1b){};
+
+	\draw (X1) -- (X1b);
+
+	\node[draw, right of=X1b, node distance=1cm](LK1){$L_{k}$};
+	\node[draw, below of=LK1, node distance=1cm](LV1){$L_{v}$};
+	\node[draw, above of=LK1, node distance=1cm](LQ1){$L_{q}$};
+
+	\draw (X1b) -- (LQ1);
+	\draw (X1b) -- (LK1);
+	\draw (X1b) -- (LV1);
+
+	\coordinate[right of=LQ1, node distance=2.6cm](Q1){};
+	\node[below of=Q1, node distance=1.5cm](Q2){$Q = \mathbf{q}_0$};
+	\node[right of=LK1, node distance=1cm](K1){$\mathbf{k}_0$};
+	\node[right of=LV1, node distance=1cm](V1){$\mathbf{v}_0$};
+
+	\draw (LQ1) -- (Q1);
+	\draw (Q1) -- (Q2);
+	\draw (LK1) -- (K1);
+	\draw (LV1) -- (V1);
+
+	\coordinate[right of= X3, node distance=2cm](X3b){};
+
+	\draw (X3) -- (X3b);
+
+	\node[draw, right of=X3b, node distance=1cm](LK3){$L_{k}$};
+	\node[draw, below of=LK3, node distance=1cm](LV3){$L_{v}$};
+
+	\draw (X3b) -- (LK3);	
+	\draw (X3b) -- (LV3);
+
+	\node[right of=LK3, node distance=1cm](K3){$\mathbf{k}_{n}$};
+	\node[right of=LV3, node distance=1cm](V3){$\mathbf{v}_{n}$};
+
+	\draw (LK3) -- (K3);
+	\draw (LV3) -- (V3);
+
+	\coordinate[right of=K1, node distance=0.3cm](TOP){};
+	\coordinate[right of=V3, node distance=0.3cm](BOT){};
+	\draw[decorate,decoration={brace}] (TOP) -- node[left=5pt]{} (BOT);
+
+	\node[right of=X2, text width=3cm, node distance=5.5cm, below=-1.0cm](EQ){\footnotesize \[K = \left( \begin{matrix}
+		\mathbf{k}_0 \\
+		\vdots \\
+		\mathbf{k}_{n}
+		\end{matrix} \right)\]
+		\\
+		\footnotesize \[ V = \left( \begin{matrix}
+		\mathbf{v}_0 \\
+		\vdots \\
+		\mathbf{v}_{n}
+		\end{matrix}\right) \]
+	};
+
+	\node[draw, right of=X2, node distance=8.3cm](EQ2){
+	\footnotesize $\sigma\left(\frac{QK^T}{\sqrt{d_k}}\right)V$};
+
+	\coordinate[left of=EQ2, node distance=1.6cm, above=2.2cm](TOP2){};
+	\coordinate[left of=EQ2, node distance=1.5cm, below=0.0cm](MID2){};
+	\coordinate[left of=EQ2, node distance=1.6cm, below=2.2cm](BOT2){};
+	\draw[decorate,decoration={brace}] (TOP2) -- node[left=5pt]{} (BOT2);
+	\draw (MID2) -- (EQ2);
+
+	\node[right of=EQ2, node distance=2cm](OUT){output};
+	\draw (EQ2) -- (OUT);
+
+	\end{tikzpicture}
+	\caption{Architecture of an ego-attention head.
+		The blocks $L_{q}$, $L_{k}$, $L_{v}$ are linear layers. The keys $K$ and values $V$ are concatenated from all vehicles, while the query $Q$ is only produced by the ego-vehicle.}
+	\label{fig:ego-attention}
+\end{figure}
+
+Each vehicle should pay attention to specific features from a selection of the other vehicles.
+This is made with four steps: pulling together specific features, identifying these feature collections,
+enquiring among identifiers, and gathering the results.
+Each head produces a different selection of features using a linear projection of
+the input tensor resulting in the value tensor $V$.
+To identify these features, a key tensor $K$ is associated to each value.
+Then, each vehicle must select which other vehicle to pay attention to.
+For that purpose a query $Q$ is produced to find a selection of keys.
+The match score between a key and a query is the dot product is scaled with the square root of the key
+dimension $\sqrt{d_k}$ and normalized with a softmax.
+This produces an attention matrix that contains coefficients close to $1$ for matching queries and keys
+and close to $0$ otherwise.
+Finally, this matrix is used to gather the values from $V$.
+Thus, the self-attention computation for each head is written:
+\begin{equation}
+\text{output}=\underbrace{\sigma\left(\frac{QK^T}{\sqrt{d_k}}\right)}_{\text{attention matrix}}V
+\label{eq_selfattention}
+\end{equation}
+The outputs from all heads are concatenated and combined with a linear layer.
+The resulting tensor is then added to the input as in residual networks.
 

### 2019-09-04 main.tex
Update content

@@
 % [preprint] option:
- \usepackage[preprint]{neurips_2019_ml4ad}
+ \usepackage[preprint, nonatbib]{neurips_2019_ml4ad}
 
@@
 % Custom packages
+\usepackage[round]{natbib}
 \usepackage{amsmath}
@@
 \usetikzlibrary{intersections}
+\usepackage{subcaption}
+
+\input{mathdef}
 
@@
   Edouard Leurent\thanks{Equal contribution.} \\
-  Inria, SequeL team\\
-  Renault Group\\
-  Paris, France\\
-  \texttt{edouard.leurent@renault.com} \\
+  SequeL team, INRIA Lille -- Nord Europe\\
+  Renault Group, France\\
+  \texttt{edouard.leurent@inria.fr} \\
   % examples of more authors
@@
   Jean Mercat$^*$ \\
-  Laboratoire des signaux et des syst\`emes,
-  Centrale-Sup\'elec\\
-  Renault Group\\
-  Paris, France\\
+  Laboratoire des signaux et des syst\`emes, Centrale-Sup\'elec\\
+  Renault Group, France\\
   \texttt{jean.mercat@renault.com} \\
@@
 
+Contributions:
+- We propose an attention-based architecture for decision-making involving social interactions, that accounts for two-way interactions between the ego-vehicle and every other traffic participant
+- We evaluate our model on a challenging intersection-crossing task involving up to 20 vehicles perceived simultaneously.
+
 \section{Background and Related Work}
 
-\subsection{Model-Free Deep Reinforcement Learning}
+\paragraph{Model-free deep reinforcement learning} Reinforcement Learning (RL) is a general framework for decision-making under uncertainty.
+It frames the learning objective as the optimal control of a Markov Decision Process  $(S, A, P, R, \gamma)$ with measurable state space $S$, action space $A$, unknown reward function $R\in\Real^{S \times A}$, and unknown dynamics $P\in \cM(S)^{S \times A}$, where $\cM(\mathcal{X})$ denotes the probability measures over a set $\mathcal{X}$.
+We seek a policy $\pi\in\cM(A)^S$ with maximal expected $\gamma$-discounted cumulative reward, called the value function $V^\pi$.
+Formally,
 
-- MDP
-- Value function
-- DQN
+\begin{align*}
+V^\pi(s) &\eqdef \expectedvalue\left[\sum_{t=0}^\infty \gamma^t R(s_t, a_t)\condbar s_0=s, a_t\sim \pi(a_t|s_t), s_{t+1}\sim P(s_{t+1}|s_t, a_t)\right]\\
+Q^\pi(s, a) &\eqdef R(s, a) + \gamma \expectedvalueover{s'\sim P(s'|s, a)} V^\pi(s')
+\end{align*}
 
-\subsection{State Representation for Social Interactions}
+The optimal action-value function $Q^* =  \max_\pi Q^\pi(s)$ satisfies the Bellman Optimality Equation:
+\begin{equation*}
+Q^*(s, a) = (\cT Q^*) (s, a) \eqdef \expectedvalueover{s'\sim P(s'|s, a)} \max_{a'\in A} \left[R(s, a) + \gamma Q^*(s', a')\right]
+\end{equation*}
+
+As $Q^*$ is a fixed-point of the Bellman Operator $\cT$ -- which is a contraction --, it can be computed by applying $\cT$ iteratively in a fixed-point iteration fashion.
+The \emph{Q-learning} algorithm \citep{Watkins} follows this procedure by applying a sampling version $\cT$ to a batch of collected experience.
+When dealing with a continuous state space $S$, we need to employ function approximation in order to generalise to nearby states.
+The \emph{Deep Q-Network} (DQN) algorithm \citep{Mnih} implements this idea by using a neural network model to represent the action-value function $Q$.
+
+\paragraph{State-representation for social interactions}{
+
 In the state:
@@
 Weight sharing ?
+Socially aware}
+
+\paragraph{Attention mechanisms} {in neural networks have been able to introduce inter-dependencies within a variable number of inputs.
+It has been used for pedestrian trajectory forecasting in~\cite{Vemula2018} with spatiotemporal graphs and
+in~\cite{Sadeghian2019CVPR} with spatial and social attention using a generative neural network.
+In~\cite{Sadeghian2018ECCV}, attention over top-view road scene images for car trajectory forecasting is used.
+Multi-head attention mechanism has been developed in~\cite{Vaswani2017} for sentence translation.
+In~\cite{Messaoud2019} a mechanism called non-local multi-head attention is developed.
+However, this is a spatial attention that does not allow vehicle-to-vehicle attention.
+
+In the present work, we use a multi-head social attention mechanism working together with
+recurrent neural networks to forecast in the form of sequences of multi-modal position probability density
+functions.}
 
@@
 	\draw (DEC1.east) -- (Y1.west);
-
-	%        \draw[decorate,decoration={brace,mirror}] (X1.north west) -- node[left=10pt] {$X$} (X3.south west);
 	\end{tikzpicture}
@@
 
-	\coordinate[right of=LQ1, node distance=2.6cm](Q1){};
-	\node[below of=Q1, node distance=1.5cm](Q2){$Q = \mathbf{q}_0$};
+	\node[right of=LQ1, node distance=1cm](Q1){$\mathbf{q}_0$};
 	\node[right of=LK1, node distance=1cm](K1){$\mathbf{k}_0$};
@@
 	\draw (LQ1) -- (Q1);
-	\draw (Q1) -- (Q2);
 	\draw (LK1) -- (K1);
@@
 
-	\coordinate[right of=K1, node distance=0.3cm](TOP){};
+	\coordinate[right of=Q1, node distance=0.3cm](TOP){};
 	\coordinate[right of=V3, node distance=0.3cm](BOT){};
@@
 
-	\node[right of=X2, text width=3cm, node distance=5.5cm, below=-1.0cm](EQ){\footnotesize \[K = \left( \begin{matrix}
+	\node[right of=X2, text width=3cm, node distance=5.5cm](EQ){
+		\footnotesize \[Q = \left( \begin{matrix}
+		\mathbf{q}_0
+		\end{matrix} \right)\]
+		\\
+		\footnotesize \[K = \left( \begin{matrix}
 		\mathbf{k}_0 \\
@@
 \section{Experiments}
-\subsection{The intersection task}
+\subsection{Description of the environment}
 
-We consider a situation where interactions between vehicles is important: crossing an unsignalized intersection.
+In this application, we use the \href{https://github.com/eleurent/highway-env}{highway-env} environment \citep{highway-env} for simulated highway driving and behavioural decision-making.
+We propose a new task where vehicles must interact significantly with each others: a four-way intersection.
+The scene, composed of two roads crossing perpendicularly, is populated with several traffic participants initialized with random positions, velocities, and destinations.
+As in \citep{highway-env}, these vehicle are simulated with the Kinematic Bicycle Model, their lateral control is achieved by a low-level steering controller and their longitudinal behaviour follows the Intelligent Driver Model.
+However, this model only considers same-lane interactions and special care was required to prevent the lateral collisions at the intersection.
+To that end, we implemented the following simplistic behaviour: each vehicle predicts the future positions of its neighbours over a three-second horizon by using a constant velocity model.
+In case of predicted collision with a neighbour, the yielding vehicle is determined randomly and must brake until the collision prediction ceases.
 
-To that end, we use the highway-env environment and propose a novel intersection task.
+In this context, the agent must drive a vehicle by controlling its acceleration chosen from a finite set of actions $A = \{\texttt{SLOWER}, \texttt{NO-OP}, \texttt{FASTER}\}$.
+The lateral control is performed automatically by a low-level controller, such that the problem complexity is focused on the high-level interactions with other vehicles, namely the decision to either yield or take way.
+The agent is rewarded by $1$ when it drives at maximum velocity, $0$ otherwise, and by $-5$ when a collision occurs.
 
-- The intersection: cross (North/South/East/West), single lane or 2 lane ?
-- The other vehicles: are spawned randomly, and plan a route to a random direction.
-Follow a Kinematic Bicycle Model, lane keeping with a low-level steering controller, longitudinal control with IDM
+\subsection{Agents}
 
-\subsection{The baselines}
-\subsection{Performances}
+\subsection{Results}
+
+\begin{table}
+	\centering
+	\begin{tabular}{lcc}
+		\toprule
+		Architecture & MLP & Ego-Attention \\
+		\midrule 
+		Layers sizes & [128, 128] & [64, 64], 64, [64, 64] \\
+		Number of parameters & \textbf{3.0e4} & 3.4e4 \\
+		Training Time (w.r.t. MLP) & \textbf{1} & 1.25 \\
+		Mean total reward & 3.6 & \textbf{6.4} \\ 
+		\bottomrule
+	\end{tabular}
+\end{table}
+
+\begin{figure}
+	\centering
+	\begin{subfigure}[t]{.65\linewidth}
+		\centering\includegraphics[width=\linewidth]{img/total_rewards}
+		\caption{Cumulative reward}
+	\end{subfigure}
+\\
+	\begin{subfigure}[t]{.65\linewidth}
+		\centering\includegraphics[width=\linewidth]{img/total_length}
+		\caption{Episode length}
+	\end{subfigure}
+\\
+	\begin{subfigure}[t]{.65\linewidth}
+		\centering\includegraphics[width=\linewidth]{img/collision}
+		\caption{Probability of collision}
+	\end{subfigure}
+\end{figure}
+
 \subsection{Attention interpretation}
@@
 
+\bibliographystyle{named}
+\bibliography{references}
+
 \end{document}

### 2019-09-05 main.tex
Add intro and background on state repr

@@
 \usepackage{subcaption}
+\usepackage[flushleft]{threeparttable}
+\usepackage{cases}
 
@@
 
-Automated Decision Making
-Modular pipeline: perception, high level representation, decision: behaviour planning, trajectory, control
-large litterature on perception
-trajectory optimization and control is well understood
+In the last decades, the problem of \emph{behavioural planning} -- that is, high-level decision-making in the context of autonomous driving -- has arguably received less attention and seen less progress than the other components of the typical robotics pipeline: perception and control \citep{architecture}.
+Indeed, the vast majority of existing systems still rely on hand-crafted rules encoded as Finite State Machines \citep{fsm}.
+As a result, only a narrow set of specified use-cases are addressed as these methods cannot scale to more complex scenes, especially when the decision-making process involves interacting with other human drivers whose behaviour is uncertain and difficult to model explicitly.
 
-behaviour planning by rule-based systems, fsm
-but behaviour planning is difficult, complex scenes, in particular interactions with other drivers difficult to model explicitly.
- -> leverage (deep) learning method
-Learning from demonstration: nice results for lane keeping, not so much for interaction, also require costly demonstration .
+This observation has led the community to turn to learning-based methods, which bear the promise of leveraging data to automatically learn a complex driving policy.
+In the \emph{imitation learning} approach, a policy can be trained in a supervised setting to imitate human driving decisions \citep{supervised, waymo}.
+Because the cost of human driving data collection at large scale can be prohibitive, another promising approach is train a policy in simulation using reinforcement learning \citep{examples}.
 
-Conversely, leverage simulation:
-Reinforcement learning:
-Framing as an MDP, choose a good representation.
-List of vehicles with their coordinates 
-but variable size, + combinatorial representation
-Occupancy grid / images : fixed size, tradeoff:  low accuracy vs exploding size
+Beyond the choice of reinforcement learning algorithm, the formalization of the problem as a \emph{Markov Decision Process} (MDP) plays an important part in the design of the system.
+In particular, the choice of state and action spaces can have a significant influence on the performance of resulting policy.
+Describing a driving scene typically involves choosing a representation the road, signalization, and vehicles.
+The simplest form: a list of vehicles with their coordinates, orientations, velocities.
+But variable size and order dependent.
+Another common format is a spatial matrix, either cartesian (images, occupancy grid) or polar (distances in angular sectors).
+Fixed size and order independent, but accuracy/size tradeoff.
+Challenge: keep a small state size, but enforce that the model is invariant to permutations.
 
@@
 
-\paragraph{Model-free deep reinforcement learning} Reinforcement Learning (RL) is a general framework for decision-making under uncertainty.
+\paragraph{Model-free deep reinforcement learning} Reinforcement Learning is a general framework for sequential decision-making under uncertainty.
 It frames the learning objective as the optimal control of a Markov Decision Process $(S, A, P, R, \gamma)$ with measurable state space $S$, action space $A$, unknown reward function $R\in\Real^{S \times A}$, and unknown dynamics $P\in \cM(S)^{S \times A}$, where $\cM(\mathcal{X})$ denotes the probability measures over a set $\mathcal{X}$.
-We seek a policy $\pi\in\cM(A)^S$ with maximal expected $\gamma$-discounted cumulative reward, called the value function $V^\pi$.
+The objective is to find a policy $\pi\in\cM(A)^S$ with maximal expected $\gamma$-discounted cumulative reward, called the value function $V^\pi$.
 Formally,
@@
 
-The optimal action-value function $Q^* =  \max_\pi Q^\pi(s)$ satisfies the Bellman Optimality Equation:
+The optimal action-value function $Q^* =  \max_\pi Q^\pi(s)$ satisfies the Bellman Optimality Equation \citep{bellman}:
 \begin{equation*}
@@
 
-In the state:
-Kinematics
-Not fixed size
-Occupancy grid
+A vehicle driving on a road can be described in the most general way by it's continuous position, heading and velocity.
+The composite state (or joint-state) of a road traffic with one ego-vehicle (denoted $X_0$) and N other vehicles can then be described by the set of the states of all vehicles.
 
-In the model:
+\begin{equation*}
+s = \left\lbrace s_k \right\rbrace_{k \in [0, N]}\qquad
+,\text{where}\qquad
+s_k = \begin{bmatrix}
+x_k & y_k & v^x_k & v^y_k & \cos\psi_k & \sin \psi_k
+\end{bmatrix}^T
+\end{equation*}
 
-Weight sharing ?
-Socially aware}
+\begin{figure}[tp]
+	\centering
+	\includegraphics[width=0.25\textwidth]{img/coordinates}
+	\includegraphics[width=0.25\textwidth]{img/map}
+	\caption{The continuous kinematics, and spatial grid representations}
+\end{figure}
+
+The size of this state-space is $\mathbb{R}^{6(N+1)}$.
+
+This representation is used in \citep{Forbes1995, Wheeler2015, Bai2015, Gindele2015, Song2016, Sunberg2017, Paxton2017, Lee2017, Shalev-Shwartz2017, Galceran2017, Chen2017, Paxton2017}.
+
+This encoding is efficient in the sense that it uses the smallest quantity of information necessary to represent the scene.
+However, it lacks two important properties.
+First, its size varies with the number of vehicles, which can be problematic for the sake of function approximation which often expects constant-sized inputs but also to because of the growing computational complexity when more vehicles are present.
+Second, we expect a driving policy to be \emph{permutation invariant}, i.e.
+not to be dependent on the order in which all vehicles in the traffic are listed.
+Ideally, this property should be enforced instead of relying on data augmentation to cover the $N!$ possible permutations of any given traffic state.
+Formally, we require that
+\begin{equation}
+\pi(\cdot|(s_1,\dotsc,s_N)) = \pi(\cdot|(s_{\sigma(1)},\dotsc,s_{\sigma(N)})) \quad\quad\quad \forall\sigma \in \mathfrak{S}_N
+\end{equation}
+
+These limitations are addressed by the spatial matrix representation, that uses a different approach for representing a quantity $z$ localized in a space $X$.
+Instead of explicitly representing spatial dimensions as variables $x$ within a state $\{s_k=(x_k,z_k)\}_{k\in[0,N]}$ indexed on the vehicles, they are represented implicitly through the layout of several variables $z_i$ organized in a grid-like structure indexed on a quantization of the space $X$.
+
+\begin{equation}
+X = \underset{i\in I}{\oplus} X_i
+\end{equation}
+
+\begin{equation}
+s_{i} = \begin{cases}
+z_k & if \exists k\in[1,N] s.t. x_k \in X_i \\
+0 & else
+\end{cases}
+\end{equation}
+
+The $z$ variable often corresponds to mere presence information (0-1) but can also include additional channels such as heading and velocity.
+The size of this state space is then $|Z|^{|I|}$.
+This representation is used in \citep{Mukadam2018, Isele2017, Fridman2018}.
+
+TODO: Can also include other layers, and top-view images
+State dimensionality is |I| = product(Size / Accuracy)
+
+Transition:
+This permutation invariance property can be implemented within the policy architecture, as done in \citep{Chen2017} or \citep{Qi2016}, but also directly in the state representation.
+Weight sharing and reduce operator that is \citep{Socially aware, point_net}.
+A particular instantiation of this idea: attention.
 
@@
 	\centering
+	\begin{threeparttable}
+    \caption{Results}
 	\begin{tabular}{lcc}
@@
 		Number of parameters & \textbf{3.0e4} & 3.4e4 \\
-		Training Time (w.r.t. MLP) & \textbf{1} & 1.25 \\
+		Inference Time$^*$ & \textbf{1} & 1.25 \\
 		Mean total reward & 3.6 & \textbf{6.4} \\ 
@@
 	\end{tabular}
+	\begin{tablenotes}
+		\small
+		\item $^*$ with respect to MLP.
+	\end{tablenotes}
+	\end{threeparttable}
 \end{table}

### 2019-09-06 main.tex
Update

--- 2019-09-04 main.tex
@@
 
-Automated Decision Making
-Modular pipeline: perception, high level representation, decision: behaviour planning, trajectory, control
-large litterature on perception
-trajectory optimization and control is well understood
+In the last decades, the problem of \emph{behavioural planning} -- that is, high-level decision-making in the context of autonomous driving -- has arguably received less attention and seen less progress than the other components of the typical robotics pipeline: perception and control \citep{Gonzalez2016}.
+Indeed, the vast majority of existing systems still rely on hand-crafted rules encoded as Finite State Machines \citep{Paden2016}.
+As a result, only a narrow set of specified use-cases are addressed as these methods cannot scale to more complex scenes, especially when the decision-making process involves interacting with other human drivers whose behaviour is uncertain and difficult to model explicitly.
 
-behaviour planning by rule-based systems, fsm
-but behaviour planning is difficult, complex scenes, in particular interactions with other drivers difficult to model explicitly.
- -> leverage (deep) learning method
-Learning from demonstration: nice results for lane keeping, not so much for interaction, also require costly demonstration .
+This observation has led the community to turn to learning-based methods, which bear the promise of leveraging data to automatically learn a complex driving policy.
+In the imitation learning approach, a policy can be trained in a supervised manner to imitate human driving decisions \citep{Pomerleau1989, Ross2011, Bojarski2016, Xu2016, Eraqi2017, Codevilla2017, Rehder2017c, Rezagholiradeh2018, Bansal2018}.
+Because the cost of human driving data collection at large scale can be prohibitive, another promising approach is train a policy in simulation using reinforcement learning \citep{Cardamone2009, Ross2011}.
 
-Conversely, leverage simulation:
-Reinforcement learning:
-Framing as an MDP, choose a good representation.
-List of vehicles with their coordinates 
-but variable size, + combinatorial representation
-Occupancy grid / images : fixed size, tradeoff:  low accuracy vs exploding size
+Beyond the choice of reinforcement learning algorithm, the formalization of the problem as a Markov Decision Process (MDP) plays an important part in the design of the system.
+In particular, the choice of state and action spaces can have a significant influence on the performance of resulting policy.
+Describing a driving scene typically involves choosing a representation the road and vehicles.
+The simplest form: a list of vehicles with their coordinates, orientations, velocities.
+But variable size and order-dependent.
+Another common format is a spatial matrix, either Cartesian (images, occupancy grid) or polar (distances in angular sectors).
+Fixed size and order independent, but accuracy/size tradeoff.
+Challenge: keep a small state size, but enforce that the model is invariant to permutations.
 
@@
 
-In the state:
-Kinematics
-Not fixed size
-Occupancy grid
+A vehicle driving on a road can be described in the most general way by it's continuous position, heading and velocity.
+The composite state (or joint-state) of a road traffic with one ego-vehicle (denoted $s_0$) and N other vehicles can then be described by the set of the states of all vehicles.
 
-In the model:
+\begin{equation*}
+s = \left\lbrace s_k \right\rbrace_{k \in [0, N]}\qquad
+\text{where}\qquad
+s_k = \begin{bmatrix}
+x_k & y_k & v^x_k & v^y_k & \cos\psi_k & \sin \psi_k
+\end{bmatrix}^T
+\end{equation*}
 
-Weight sharing ?
-Socially aware}
+\begin{figure}[tp]
+	\centering
+	\includegraphics[width=0.25\textwidth]{img/coordinates}
+	\includegraphics[width=0.25\textwidth]{img/map}
+	\caption{The coordinates (left) and spatial-grid (right) representations}
+\end{figure}
+
+This representation is used in \citep{Bai2015, Gindele2015, Song2016, Sunberg2017, Paxton2017, Galceran2017, Chen2017}.
+
+This encoding is efficient in the sense that it uses the smallest quantity of information necessary to represent the scene.
+However, it lacks two important properties.
+First, its size varies with the number of vehicles which can be problematic for the sake of function approximation which often expects constant-sized inputs, but also for concerns of varying computational complexity.
+Second, we expect a driving policy to be \emph{permutation invariant}, i.e.
+not to be dependent on the order in which other traffic participants are listed.
+Ideally, this property should be enforced and not relying on the coverage of the $N!$ possible permutations of any given traffic state.
+Formally, we require that:
+\begin{equation*}
+\pi(\cdot|(s_1,\dotsc,s_N)) = \pi(\cdot|(s_{\sigma(1)},\dotsc,s_{\sigma(N)})) \quad\quad\quad \forall\sigma \in \mathfrak{S}_N
+\end{equation*}
+
+A popular way to address this limitations is to use a spatial matrix representation.
+Instead of explicitly representing spatial information as variables $x, y$ within the state $\{s_k=(x_k,y_k,f_k)\}_{k\in[0,N]}$ indexed on the vehicles, they are instead represented implicitly through the layout of several feature variables $f_ij$ organized in a tensor, where the $i,j$ indexes refer to cells of a quantization of the 2D-space.
+Note that the size of this tensor is related to the area covered divided by the length of the quantization, which denotes a trade-off between accuracy and computational burden.
+In an occupancy grid, the $f$ features contains presence information (0-1) and additional channels such as velocity and heading, as in \citep{Isele2017, Fridman2018, Bansal2018, Rehder2017c}.
+Another example is the use of top-view RGB images \citep{Bagnell2010, Rehder2017, Rehder2017c, Liu2018}.
+
+Transition:
+This permutation invariance property can be implemented within the policy architecture, as done in \citep{Chen2017} or \citep{Qi2016}, but also directly in the state representation.
+Weight sharing and reduce operator that is invariant to permutation eg max in \citep{Chen2017} and sum in \citep{point_net}.
+A particular instantiation of this idea: attention.
 
@@
 \section{Experiments}
-\subsection{Description of the environment}
+\paragraph{The environment}
 
@@
 	\centering
-	\begin{tabular}{lcc}
+	\begin{threeparttable}
+    \caption{Results}
+	\begin{tabular}{lccc}
 		\toprule
-		Architecture & MLP & Ego-Attention \\
+		Architecture & MLP Coordinates & MLP Grid & Ego-Attention \\
 		\midrule 
-		Layers sizes & [128, 128] & [64, 64], 64, [64, 64] \\
-		Number of parameters & \textbf{3.0e4} & 3.4e4 \\
-		Training Time (w.r.t. MLP) & \textbf{1} & 1.25 \\
-		Mean total reward & 3.6 & \textbf{6.4} \\ 
+		Layers sizes & [128, 128] &  [64, 64] & [64, 64], 64, [64, 64] \\
+		Number of parameters & \textbf{3.0e4} & ? & 3.4e4 \\
+		Mean total reward & 3.7 & 4.5 & \textbf{6.4} \\ 
 		\bottomrule
 	\end{tabular}
+%	\begin{tablenotes}
+%		\small
+%		\item 
+%	\end{tablenotes}
+	\end{threeparttable}
 \end{table}
@@
 	\centering
-	\begin{subfigure}[t]{.65\linewidth}
-		\centering\includegraphics[width=\linewidth]{img/total_rewards}
-		\caption{Cumulative reward}
+	\begin{subfigure}[t]{.49\linewidth}
+		\centering\includegraphics[width=\linewidth]{img/total_reward}
 	\end{subfigure}
-\\
-	\begin{subfigure}[t]{.65\linewidth}
-		\centering\includegraphics[width=\linewidth]{img/total_length}
-		\caption{Episode length}
+	\begin{subfigure}[t]{.49\linewidth}
+		\centering\includegraphics[width=\linewidth]{img/length}
 	\end{subfigure}
-\\
-	\begin{subfigure}[t]{.65\linewidth}
-		\centering\includegraphics[width=\linewidth]{img/collision}
-		\caption{Probability of collision}
+	\begin{subfigure}[t]{.49\linewidth}
+		\centering\includegraphics[width=\linewidth]{img/velocity}
+	\end{subfigure}
+	\begin{subfigure}[t]{.49\linewidth}
+	\centering\includegraphics[width=\linewidth]{img/distance}
     \end{subfigure}

### 2019-09-09 main.tex
Update manuscript

@@
 % ready for submission
-%\usepackage{neurips_2019_ml4ad}
+\usepackage[nonatbib]{neurips_2019_ml4ad}
 
@@
 % [preprint] option:
- \usepackage[preprint, nonatbib]{neurips_2019_ml4ad}
+% \usepackage[preprint, nonatbib]{neurips_2019_ml4ad}
 
@@
 \usepackage{tikz}
+\usepackage{xspace}
 \usepackage{float}
@@
 \usepackage{cases}
+\usepackage{multirow}
 
@@
   \texttt{edouard.leurent@inria.fr} \\
-  % examples of more authors
    \And
@@
 \begin{abstract}
-  The abstract paragraph should be indented \nicefrac{1}{2}~inch (3~picas) on
-  both the left- and right-hand margins. Use 10~point type, with a vertical
-  spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
-  bold, and in point size 12. Two line spaces precede the abstract. The abstract
-  must be limited to one paragraph.
+  TODO
 \end{abstract}
@@
 - We propose an attention-based architecture for decision-making involving social interactions, that accounts for two-way interactions between the ego-vehicle and every other traffic participant
-- We evaluate our model on a challenging intersection-crossing task involving up to 20 vehicles perceived simultaneously.
+- We evaluate our model on a challenging intersection-crossing task involving up to 15 vehicles perceived simultaneously.
 
@@
 
-This representation is used in \citep{Bai2015, Gindele2015, Song2016, Sunberg2017, Paxton2017, Galceran2017, Chen2017}.
+For instance, this representation is used in \citep{Bai2015, Gindele2015, Song2016, Sunberg2017, Paxton2017, Galceran2017, Chen2017}.
 
@@
 First, its size varies with the number of vehicles which can be problematic for the sake of function approximation which often expects constant-sized inputs, but also for concerns of varying computational complexity.
-Second, we expect a driving policy to be \emph{permutation invariant}, i.e.
+Second, we expect a driving policy $\pi$ to be \emph{permutation invariant}, i.e.
 not to be dependent on the order in which other traffic participants are listed.
-Ideally, this property should be enforced and not relying on the coverage of the $N!$ possible permutations of any given traffic state.
+Ideally, this property should be enforced and not approximated by relying on the coverage of the $N!$ possible permutations $\tau$ of any given traffic state.
 Formally, we require that:
 \begin{equation*}
-\pi(\cdot|(s_1,\dotsc,s_N)) = \pi(\cdot|(s_{\sigma(1)},\dotsc,s_{\sigma(N)})) \quad\quad\quad \forall\sigma \in \mathfrak{S}_N
+\pi(\cdot|(s_0, s_1,\dotsc,s_N)) = \pi(\cdot|(s_0, s_{\tau(1)},\dotsc,s_{\tau(N)})) \quad\quad\quad \forall\tau \in \mathfrak{S}_N
 \end{equation*}
@@
 A popular way to address this limitations is to use a spatial matrix representation.
-Instead of explicitly representing spatial information as variables $x, y$ within the state $\{s_k=(x_k,y_k,f_k)\}_{k\in[0,N]}$ indexed on the vehicles, they are instead represented implicitly through the layout of several feature variables $f_ij$ organized in a tensor, where the $i,j$ indexes refer to cells of a quantization of the 2D-space.
-Note that the size of this tensor is related to the area covered divided by the length of the quantization, which denotes a trade-off between accuracy and computational burden.
+Instead of explicitly representing spatial information as variables $x, y$ along with other features $f$ directly inside the state $\{s_k=(x_k,y_k,f_k)\}_{k\in[0,N]}$ indexed on the vehicles, they are instead represented implicitly through the layout of several feature variables $f_ij$ organized in a tensor structure, where the $i,j$ indexes refer to the cells of a quantization of the 2D-space.
+Note that the size of this tensor is related to the area covered divided by the length of the quantization, which reflects a trade-off between accuracy and computational load.
 In an occupancy grid, the $f$ features contains presence information (0-1) and additional channels such as velocity and heading, as in \citep{Isele2017, Fridman2018, Bansal2018, Rehder2017c}.
@@
 
-Transition:
-This permutation invariance property can be implemented within the policy architecture, as done in \citep{Chen2017} or \citep{Qi2016}, but also directly in the state representation.
-Weight sharing and reduce operator that is invariant to permutation eg max in \citep{Chen2017} and sum in \citep{point_net}.
-A particular instantiation of this idea: attention.
+This permutation invariance property can also be implemented within the policy architecture.
+A general technique to achieve this is to treat each entity similarly in the early stages -- e.g.
+through weight sharing -- before reducing them with a projection operator that is itself invariant to permutations, for instance a max-pooling as in \citep{Chen2017} or a sum as in \citep{Qi2016}.
+A particular instantiation of this idea is the use of attention mechanisms.
 
-\paragraph{Attention mechanisms} {in neural networks have been able to introduce inter-dependencies within a variable number of inputs.
-It has been used for pedestrian trajectory forecasting in~\cite{Vemula2018} with spatiotemporal graphs and
-in~\cite{Sadeghian2019CVPR} with spatial and social attention using a generative neural network.
+\paragraph{Attention mechanisms} {The attention architecture was introduced to enable neural networks to discover inter-dependencies within a variable number of inputs.
+It has been used for pedestrian trajectory forecasting in~\cite{Vemula2018} with spatiotemporal graphs and in~\cite{Sadeghian2019CVPR} with spatial and social attention using a generative neural network.
 In~\cite{Sadeghian2018ECCV}, attention over top-view road scene images for car trajectory forecasting is used.
@@
 However, this is a spatial attention that does not allow vehicle-to-vehicle attention.
-
-In the present work, we use a multi-head social attention mechanism working together with
-recurrent neural networks to forecast in the form of sequences of multi-modal position probability density
-functions.}
+In the present work, we use a multi-head social attention mechanism...}
 
@@
 	\node[below of=X2, node distance=0.6cm](X3){$\vdots$};
-	\node[below of=X3, node distance=0.7cm](X4){vehicle$_{n}$};
+	\node[below of=X3, node distance=0.7cm](X4){vehicle$_{N}$};
 
@@
 	\node[below of=X1, node distance=2cm](X2){$\vdots$};
-	\node[below of=X2, node distance=2cm](X3){vehicle$_{n}$ encoding};
+	\node[below of=X2, node distance=2cm](X3){vehicle$_{N}$ encoding};
 
@@
 
+%Descriptive features, names keys $k_i$
+%The similarity of the query $q_i$ to the keys $k_j$ is assessed through a similarity measure implemented as a scalar product: $<q_i, k_j>$, $QK^T$.
+%These similarities are then normalized by a softmax $\sigma$ function and a division by $\sqrt{d_k}}$ where $d_k$ is the dimension of keys and queries. The final value is obtained.
+
+Self attention
+
 Each vehicle should pay attention to specific features from a selection of the other vehicles.
-This is made with four steps: pulling together specific features, identifying these feature collections,
-enquiring among identifiers, and gathering the results.
+This is made with three steps: ...
 Each head produces a different selection of features using a linear projection of
@@
 To identify these features, a key tensor $K$ is associated to each value.
-Then, each vehicle must select which other vehicle to pay attention to.
+Then, each vehicle (TODO: or just the ego) must select which other vehicle to pay attention to.
 For that purpose a query $Q$ is produced to find a selection of keys.
@@
 
+Ego attention: only the ego-vehicle produces a query, we get one output feature.
+We can easily see that the result is permutation invariant: indeed, a permutation $\tau$ will change the order of the rows of keys $K$ and values $V$ but will keep their correspondence.
+The final result is a sum of values weighted by their key-similarities, which is independent of the ordering.
+%$\sigma\left(Q\tau(K)^T/\sqrt{d_k}\right)\tau(V) = \tau\left(\sigma(QK^T/\sqrt{d_k})\right)\tau(V) = \sigma\left(QK^T/\sqrt{d_k}\right)V$
+
 \section{Experiments}
-\paragraph{The environment}
+\paragraph{Environment}
 
 In this application, we use the \href{https://github.com/eleurent/highway-env}{highway-env} environment \citep{highway-env} for simulated highway driving and behavioural decision-making.
-We propose a new task where vehicles must interact significantly with each others: a four-way intersection.
-The scene, composed of two roads crossing perpendicularly, is populated with several traffic participants initialized with random positions, velocities, and destinations.
-As in \citep{highway-env}, these vehicle are simulated with the Kinematic Bicycle Model, their lateral control is achieved by a low-level steering controller and their longitudinal behaviour follows the Intelligent Driver Model.
-However, this model only considers same-lane interactions and special care was required to prevent the lateral collisions at the intersection.
+We propose a new task where vehicle-to-vehicle interaction plays a significant part: crossing a four-way intersection.
+The scene -- composed of two roads crossing perpendicularly -- is populated with several traffic participants initialized with random positions, velocities, and destinations.
+As in \citep{highway-env}, these vehicle are simulated with the Kinematic Bicycle Model, their lateral control is achieved by a low-level steering controller tracking a target route, and their longitudinal behaviour follows the Intelligent Driver Model.
+However, this model only considers same-lane interactions and special care was required to prevent lateral collisions at the intersection.
 To that end, we implemented the following simplistic behaviour: each vehicle predicts the future positions of its neighbours over a three-second horizon by using a constant velocity model.
-In case of predicted collision with a neighbour, the yielding vehicle is determined randomly and must brake until the collision prediction ceases.
+In case of predicted collision with a neighbour, the yielding vehicle is determined based on road priorities, and must brake until the collision prediction ceases.
 
@@
 
-\subsection{Agents}
+\paragraph{Agents}
 
-\subsection{Results}
+TODO, see \autoref{tab:agents}.
 
-\begin{table}
+\begin{table}[tp]
 	\centering
 	\begin{threeparttable}
-    \caption{Results}
+		\caption{Characteristics of the agents}
+		\label{tab:agents}
 		\begin{tabular}{lccc}
 			\toprule
-		Architecture & MLP Coordinates & MLP Grid & Ego-Attention \\
+			Architecture & \MLPC & \MLPG & \EgoAtt \\
 			\midrule 
-		Layers sizes & [128, 128] &  [64, 64] & [64, 64], 64, [64, 64] \\
-		Number of parameters & \textbf{3.0e4} & ? & 3.4e4 \\
-		Mean total reward & 3.7 & 4.5 & \textbf{6.4} \\ 
+			Input sizes & [15, 7] & [?, ?, 7] & [1-15, 7] \\
+			Layers sizes & [128, 128] &  [64, 64] & Encoder: [64, 64] \\
+			& & & Attention $d_k$: 64\\
+			& & & Decoder: [64, 64]\\
+			Number of parameters & 3.0e4 & ? & 3.4e4 \\
+			Permutation invariant & No & Yes & Yes \\
 			\bottomrule
 		\end{tabular}
-%	\begin{tablenotes}
-%		\small
-%		\item 
-%	\end{tablenotes}
 	\end{threeparttable}
@@
 
-\begin{figure}
+\paragraph{Performances}
+
+We show in \autoref{fig:results} the evolution of the total reward, episode length and average velocity.
+The \MLPC agent learns to accelerate to earn short-term rewards, as shown by its high average velocity , but fails to exploit the information of other vehicles and crashes often, leading to shorter episodes = risky policy, worst performing.
+Conversely, the \MLPG benefits from its invariance to permutations and manages to learn to brake upon arrival at the intersection to avoid collisions (increased duration).
+However, it only proceeds when the intersection has been cleared (low mean velocity), resulting in an overly cautious policy, a common trait known as the freezing robot problem.
+Slight increase in performance.
+In stark contrast, the \EgoAtt policy quickly learns both when it must slow down at the intersection but also when it can exploit the gaps in the traffic and take way to vehicles that are far/slow enough.
+The overall resulting behaviour is qualitatively more nuanced and human-like.
+
+\begin{figure}[tp]
 	\centering
-	\begin{subfigure}[t]{.49\linewidth}
+	\begin{subfigure}[t]{.75\linewidth}
 		\centering\includegraphics[width=\linewidth]{img/total_reward}
 	\end{subfigure}
+\\
 	\begin{subfigure}[t]{.49\linewidth}
@@
 	\end{subfigure}
-	\begin{subfigure}[t]{.49\linewidth}
-	\centering\includegraphics[width=\linewidth]{img/distance}
-    \end{subfigure}
+%	\begin{subfigure}[t]{.49\linewidth}
+%	\centering\includegraphics[width=\linewidth]{img/distance}
+ %   \end{subfigure}
+ \caption{The episode total rewards, lengths, and average velocity (higher is better). We display the mean values -- along with their 95\% confidence interval -- averaged over 120 random seeds.}
+ \label{fig:results}
 \end{figure}
 
-\subsection{Attention interpretation}
+\paragraph{Attention interpretation}
+
+TODO: figure
 
@@
 
+\paragraph{Exploiting interaction patterns}
+The priority rules are note enforced through rewards but interactions: based on the defined road priorities, some vehicles will take way to the ego-vehicle, while others will not.
+
+By switching the priority road, the resulting policy changes: In the exact same initial state, the ego-vehicle takes way to incoming vehicles when its policy was trained on a priority road, and gives way otherwise.
+The only change between these two situations is the rules of interactions imposed on simulated vehicles, which showcases the ability of our proposed architecture to discover and exploit such interaction patterns.
+
 \section{Conclusion}
 
-\subsubsection*{Acknowledgments}
+TODO
 
-Use unnumbered third level headings for the acknowledgments. All acknowledgments
-Go at the end of the paper. Do not include acknowledgments in the anonymized
-submission, only in the final paper.
+%\subsubsection*{Acknowledgments}
+%
+%TODO
 

### 2019-09-09 main.tex
Update

@@
 \begin{abstract}
-  TODO
+  This work addresses the design of learning architectures in the context of high-level decision making for autonomous driving in a dense traffic setting.
+We focus on three challenges: dealing with a varying number of nearby vehicles, being invariant to the ordering chosen to describe the scene, and accounting for the existing interactions between the traffic participants.
 \end{abstract}
@@
 	\caption{The coordinates (left) and spatial-grid (right) representations}
+	\label{fig:representation}
 \end{figure}
 
-For instance, this representation is used in \citep{Bai2015, Gindele2015, Song2016, Sunberg2017, Paxton2017, Galceran2017, Chen2017}.
+This representation is illustrated in \autoref{fig:representation} and was used e.g. in \citep{Bai2015, Gindele2015, Song2016, Sunberg2017, Paxton2017, Galceran2017, Chen2017}.
 
@@
 			Layers sizes & [128, 128] &  [64, 64] & Encoder: [64, 64] \\
-			& & & Attention $d_k$: 64\\
+			& & & Attention: $d_k=64$\\
 			& & & Decoder: [64, 64]\\
 			Number of parameters & 3.0e4 & ? & 3.4e4 \\
+			Variable input size & No & No & Yes \\
 			Permutation invariant & No & Yes & Yes \\
@@
 
-TODO: figure
+This model is also more interpretable, as we can visualize the learnt attention matrix.
+
+TODO: figure yield vs take way
 
@@
 
+change the priority -> influences the rules of interactions -> changes the learnt behaviour.
 By switching the priority road, the resulting policy changes: In the exact same initial state, the ego-vehicle takes way to incoming vehicles when its policy was trained on a priority road, and gives way otherwise.
@@
 
-\section*{References}
-
 \bibliographystyle{named}

### 2019-09-10 main.tex
Update intro

@@
 \begin{abstract}
-  This work addresses the design of learning architectures in the context of high-level decision making for autonomous driving in a dense traffic setting.
-We focus on three challenges: dealing with a varying number of nearby vehicles, being invariant to the ordering chosen to describe the scene, and accounting for the existing interactions between the traffic participants.
+  This work addresses the design of learning architectures for behavioural planning in a dense traffic setting.
+We claim that three important requirements are: dealing with a varying number of nearby vehicles, being invariant to the ordering chosen to describe them, while staying accurate and compact.
+We observe that the two most popular representations in the literature do not fit all these criteria, and show that they perform badly on an intersection-crossing task that involves complex anticipation and negotiation skills.
+In contrast, we propose an attention-based architecture that enjoys all these properties while also accounting for the existing interactions between the traffic participants.
+We show that this architecture leads to significant improvements in performance, and that it is able to capture interactions patterns that can be visualized and qualitatively interpreted.
 \end{abstract}
@@
 Indeed, the vast majority of existing systems still rely on hand-crafted rules encoded as Finite State Machines \citep{Paden2016}.
-As a result, only a narrow set of specified use-cases are addressed as these methods cannot scale to more complex scenes, especially when the decision-making process involves interacting with other human drivers whose behaviour is uncertain and difficult to model explicitly.
+As a result, only a narrow set of specified use-cases are addressed as these methods cannot scale to more complex scenes, especially when the decision-making process involves interacting with other human drivers whose behaviours are uncertain and difficult to model explicitly.
 
@@
 
-Beyond the choice of reinforcement learning algorithm, the formalization of the problem as a Markov Decision Process (MDP) plays an important part in the design of the system.
-In particular, the choice of state and action spaces can have a significant influence on the performance of resulting policy.
-Describing a driving scene typically involves choosing a representation the road and vehicles.
-The simplest form: a list of vehicles with their coordinates, orientations, velocities.
-But variable size and order-dependent.
-Another common format is a spatial matrix, either Cartesian (images, occupancy grid) or polar (distances in angular sectors).
-Fixed size and order independent, but accuracy/size tradeoff.
-Challenge: keep a small state size, but enforce that the model is invariant to permutations.
+Beyond the choice of reinforcement learning algorithm, the formalization of the problem as a Markov Decision Process plays an important part in the design of the system.
+Indeed, the definition of the state space involves choosing a representation of the driving scene.
+In this work, we focus in on how the vehicles are represented.
+In particular, we claim that the two most-widely used representations both suffer from different drawbacks: on the one hand, the \emph{list of features} representation is compact and accurate but has a varying-size and depends on the choice of ordering.
+On the other hand, the \emph{spatial grid} representation addresses these concerns but in return suffers from an accuracy-size trade-off.
 
-Contributions:
-- We propose an attention-based architecture for decision-making involving social interactions, that accounts for two-way interactions between the ego-vehicle and every other traffic participant
+Our contributions are the following: first, we propose an attention-based architecture for decision-making involving social interactions, that deals with the varying-size and that accounts for two-way interactions between the ego-vehicle and every other traffic participant
 - We evaluate our model on a challenging intersection-crossing task involving up to 15 vehicles perceived simultaneously.
@@
 
+In order to apply a reinforcement learning algorithm such as DQN to an autonomous driving problem, a state space $S$ must first be chosen, that is, a representation of the scene.
 A vehicle driving on a road can be described in the most general way by it's continuous position, heading and velocity.
@@
 	\includegraphics[width=0.25\textwidth]{img/map}
-	\caption{The coordinates (left) and spatial-grid (right) representations}
+	\caption{The \emph{list of features} (left) and \emph{spatial grid} (right) representations}
 	\label{fig:representation}
@@
 
-This representation is illustrated in \autoref{fig:representation} and was used e.g. in \citep{Bai2015, Gindele2015, Song2016, Sunberg2017, Paxton2017, Galceran2017, Chen2017}.
+This representation, that we call \emph{list of features}, is illustrated in \autoref{fig:representation} and was used e.g.
+in \citep{Bai2015, Gindele2015, Song2016, Sunberg2017, Paxton2017, Galceran2017, Chen2017}.
 

### 2019-09-10 main.tex
Update architecture

@@
 \usepackage{graphicx}
+\usepackage{array, makecell}
 \usepackage{tikz}
@@
 This observation has led the community to turn to learning-based methods, which bear the promise of leveraging data to automatically learn a complex driving policy.
-In the imitation learning approach, a policy can be trained in a supervised manner to imitate human driving decisions \citep{Pomerleau1989, Ross2011, Bojarski2016, Xu2016, Eraqi2017, Codevilla2017, Rehder2017c, Rezagholiradeh2018, Bansal2018}.
-Because the cost of human driving data collection at large scale can be prohibitive, another promising approach is train a policy in simulation using reinforcement learning \citep{Cardamone2009, Ross2011}.
+In the imitation learning approach, a policy can be trained in a supervised manner to imitate human driving decisions \citep[e.g.][]{Pomerleau1989, Ross2011, Bojarski2016, Xu2016, Eraqi2017, Codevilla2017, Rehder2017c, Rezagholiradeh2018, Rhinehart2018, Bansal2018, Rhinehart2019}.
+Because the cost of human driving data collection at large scale can be prohibitive, another promising approach is train a policy in simulation using reinforcement learning \citep[e.g.][]{Cardamone2009, Ross2011, Mukadam2017, Chen2017, Isele_2018, Ha2018, Kendall2019}.
 
@@
 
-Our contributions are the following: first, we propose an attention-based architecture for decision-making involving social interactions, that deals with the varying-size and that accounts for two-way interactions between the ego-vehicle and every other traffic participant
-- We evaluate our model on a challenging intersection-crossing task involving up to 15 vehicles perceived simultaneously.
+Our contributions are the following: first, we propose an attention-based architecture for decision-making involving social interactions.
+This architecture allows to satisfy the variable-size and permutation invariance requirements even when using a \emph{list of features} representation.
+It also naturally accounts for interactions between the ego-vehicle and any other traffic participant.
+Second, we evaluate our model on a challenging intersection-crossing task involving up to 15 vehicles perceived simultaneously.
+We show that our proposed method provides significant quantitative improvements, and that it enables to capture interaction patterns in a way that is visually interpretable.
 
@@
 In order to apply a reinforcement learning algorithm such as DQN to an autonomous driving problem, a state space $S$ must first be chosen, that is, a representation of the scene.
+When social interactions are relevant to the decision, the state should at leat compose a description of every nearby vehicle.
 A vehicle driving on a road can be described in the most general way by it's continuous position, heading and velocity.
-The composite state (or joint-state) of a road traffic with one ego-vehicle (denoted $s_0$) and N other vehicles can then be described by the set of the states of all vehicles.
+Then, the joint state of a road traffic with one ego-vehicle denoted $s_0$ and $N$ other vehicles can be described by the (unordered) set of individual vehicle states:
 
 \begin{equation*}
-s = \left\lbrace s_k \right\rbrace_{k \in [0, N]}\qquad
+s = \left\lbrace s_i \right\rbrace_{i \in [0, N]}\qquad
 \text{where}\qquad
-s_k = \begin{bmatrix}
-x_k & y_k & v^x_k & v^y_k & \cos\psi_k & \sin \psi_k
+s_i = \begin{bmatrix}
+x_i & y_i & v^x_i & v^y_i & \cos\psi_i & \sin \psi_i
 \end{bmatrix}^T
@@
 
-This representation, that we call \emph{list of features}, is illustrated in \autoref{fig:representation} and was used e.g.
-in \citep{Bai2015, Gindele2015, Song2016, Sunberg2017, Paxton2017, Galceran2017, Chen2017}.
+This representation, that we call \emph{list of features}, is illustrated in \autoref{fig:representation} (left) and was used for instance in \citep{Bai2015, Gindele2015, Song2016, Sunberg2017, Paxton2017, Galceran2017, Chen2017}.
 
@@
 
-A popular way to address this limitations is to use a spatial matrix representation.
-Instead of explicitly representing spatial information as variables $x, y$ along with other features $f$ directly inside the state $\{s_k=(x_k,y_k,f_k)\}_{k\in[0,N]}$ indexed on the vehicles, they are instead represented implicitly through the layout of several feature variables $f_ij$ organized in a tensor structure, where the $i,j$ indexes refer to the cells of a quantization of the 2D-space.
-Note that the size of this tensor is related to the area covered divided by the length of the quantization, which reflects a trade-off between accuracy and computational load.
-In an occupancy grid, the $f$ features contains presence information (0-1) and additional channels such as velocity and heading, as in \citep{Isele2017, Fridman2018, Bansal2018, Rehder2017c}.
-Another example is the use of top-view RGB images \citep{Bagnell2010, Rehder2017, Rehder2017c, Liu2018}.
+A popular way to address this limitations is to use a \emph{spatial grid} representation.
+Instead of explicitly representing spatial information as variables $x, y$ along with other features $f$ directly inside the state $\{s_i=(x_i,y_i,f_i)\}_{i\in[0,N]}$ indexed on the vehicles, they are instead represented implicitly through the layout of several feature variables $f_{ij}$ organized in a tensor structure, where the $(i,j)$ indexes refer to the cells of a quantization of the 2D-space.
+This representation is illustrated in \autoref{fig:representation} (right).
+Note that the size of this tensor is related to the area covered divided by the length of the quantization, which reflects a trade-off between accuracy and dimensionality.
+In an occupancy grid, the $f$ features contains presence information (0-1) and additional channels such as velocity and heading, as in \citep[e.g.][]{Isele2017, Fridman2018, Bansal2018, Rehder2017c}.
+Another example is the use of top-view RGB images \citep[e.g.][]{Bagnell2010, Rehder2017, Rehder2017c, Liu2018}.
 
@@
 A general technique to achieve this is to treat each entity similarly in the early stages -- e.g.
-through weight sharing -- before reducing them with a projection operator that is itself invariant to permutations, for instance a max-pooling as in \citep{Chen2017} or a sum as in \citep{Qi2016}.
-A particular instantiation of this idea is the use of attention mechanisms.
+through weight sharing -- before reducing them with a projection operator that is itself invariant to permutations, for instance a max-pooling as in \citep{Chen2017} or an average as in \citep{Qi2016}.
+A particular instantiation of this idea is attention mechanisms.
 
@@
 However, this is a spatial attention that does not allow vehicle-to-vehicle attention.
-In the present work, we use a multi-head social attention mechanism...}
+In the present work, we use a multi-head social attention mechanism to capture vehicle-to-ego dependencies and build varying input size and permutation invariance into the policy model.}
 
@@
 
-Motivation: 
-\begin{enumerate}
-	\item Out of a complex scene description, the model should be able to filter information and consider only what is relevant for decision.
-In other words, the agent should \emph{pay attention} to vehicles that are close or conflict with the planned route;
-	\item The model should be able to admit a variable number of vehicles as inputs;
-	\item The model should be invariant to the ordering of vehicles.
-\end{enumerate}
+Out of a complex scene description, the model should be able to filter information and consider only what is relevant for decision.
+In other words, the agent should \emph{pay attention} to vehicles that are close or conflict with the planned route.
 
-\begin{figure}[ht]
+The proposed architecture is presented in \autoref{fig:architecture}.
+It is used to represent the $Q$-function that will be optimized by the DQN algorithm.
+It is composed of a first linear encoding layer whose weights are shared between all vehicles.
+At that point, the embeddings only contain individual features of size $d_x$.
+They are then fed to an ego-attention layer, composed of several heads stacked together.
+Such an attention head is illustrated in \autoref{fig:ego-attention}, and works in the following way: in order to select a subset of vehicles depending on the context, the ego-vehicle  first emits a query $Q = [q_0]\in\Real^{1 \times d_k}$, computed with a linear projection $L_q\in\Real^{d_x \times d_k}$ of its embedding.
+This query is then compared to a set of keys $K = [k_0, \dots, k_N]\in\Real^{N \times d_k}$ containing descriptive features $k_i$ for each vehicle, again computed with a shared linear projection $L_k\in\Real^{d_x \times d_k}$.
+The similarity between the query $q_0$ and any key $k_i$ is assessed by their dot product $q_0 k_i^T$.
+These similarities are then scaled by the inverse-square-root-dimension $1/\sqrt{d_k}$\footnote{This scaling is due to the fact that the dot-product of two independent random vectors with mean 0,  variance 1, and dimension $d_k$, is a random variable with mean 0 and variance $d_k$} and normalized with a softmax function $\sigma$ across vehicles.
+We obtain a stochastic matrix called the \emph{attention matrix}, which is finally used to gather a set of output value $V = [v_0, \dots, v_N]$, where each value $v_i$ is a feature computed with a shared linear projection $L_v\in\Real^{d_x \times d_v}$.
+Overall, the ego-attention computation for each head can be written as:
+\begin{equation}
+\text{output}=\underbrace{\sigma\left(\frac{QK^T}{\sqrt{d_k}}\right)}_{\text{attention matrix}}V
+\label{eq:selfattention}
+\end{equation}
+The outputs from all heads are finally combined with a linear layer, and the resulting tensor is then added to the ego encoding as in residual networks.
+We can easily see that the result is permutation invariant: indeed, a permutation $\tau$ will change the order of the rows in keys $K$ and values $V$ in \eqref{eq:selfattention} but will keep their correspondence.
+The final result is a dot product of values and key-similarities, which is independent of the ordering.
+
+\begin{figure}[tp]
 	\centering
@@
 
-	\node[draw, rectangle, right of=ENC1, node distance=1.8cm, below=-0.2cm](TRANS1){\rotatebox{90}{ Ego-attention }};
+	\node[draw, rectangle, right of=ENC1, node distance=2.0cm, below=-0.4cm, fill=white](TRANS3){\rotatebox{90}{ Ego-attention }};
+	\node[draw, rectangle, right of=ENC1, node distance=1.9cm, below=-0.3cm, fill=white](TRANS2){\rotatebox{90}{ Ego-attention }};
+	\node[draw, rectangle, right of=ENC1, node distance=1.8cm, below=-0.2cm, fill=white](TRANS1){\rotatebox{90}{ Ego-attention }};
 
@@
 	\caption{Block diagram of our model architecture.}
-	\label{sch_whole_model}
+	\label{fig:architecture}
 
@@
 
-The multi-head self-attention layers allow vehicle interactions while keeping independence from their number
-and ordering.
-This mechanism is described in~\cite{Vaswani2017} where it is applied on sentence translation.
-In this section we explain its use for vehicle interactions.
-The computations made by each attention head is represented on figure~\ref{fig:ego-attention} and are detailed below.
-
-\begin{figure}[ht]
+\begin{figure}[tp]
 	\centering
@@
 
-%Descriptive features, names keys $k_i$
-%The similarity of the query $q_i$ to the keys $k_j$ is assessed through a similarity measure implemented as a scalar product: $<q_i, k_j>$, $QK^T$.
-%These similarities are then normalized by a softmax $\sigma$ function and a division by $\sqrt{d_k}}$ where $d_k$ is the dimension of keys and queries. The final value is obtained.
-
-Self attention
-
-Each vehicle should pay attention to specific features from a selection of the other vehicles.
-This is made with three steps: ...
-Each head produces a different selection of features using a linear projection of
-the input tensor resulting in the value tensor $V$.
-To identify these features, a key tensor $K$ is associated to each value.
-Then, each vehicle (TODO: or just the ego) must select which other vehicle to pay attention to.
-For that purpose a query $Q$ is produced to find a selection of keys.
-The match score between a key and a query is the dot product is scaled with the square root of the key
-dimension $\sqrt{d_k}$ and normalized with a softmax.
-This produces an attention matrix that contains coefficients close to $1$ for matching queries and keys
-and close to $0$ otherwise.
-Finally, this matrix is used to gather the values from $V$.
-Thus, the self-attention computation for each head is written:
-\begin{equation}
-\text{output}=\underbrace{\sigma\left(\frac{QK^T}{\sqrt{d_k}}\right)}_{\text{attention matrix}}V
-\label{eq_selfattention}
-\end{equation}
-The outputs from all heads are concatenated and combined with a linear layer.
-The resulting tensor is then added to the input as in residual networks.
-
-Ego attention: only the ego-vehicle produces a query, we get one output feature.
-We can easily see that the result is permutation invariant: indeed, a permutation $\tau$ will change the order of the rows of keys $K$ and values $V$ but will keep their correspondence.
-The final result is a sum of values weighted by their key-similarities, which is independent of the ordering.
-%$\sigma\left(Q\tau(K)^T/\sqrt{d_k}\right)\tau(V) = \tau\left(\sigma(QK^T/\sqrt{d_k})\right)\tau(V) = \sigma\left(QK^T/\sqrt{d_k}\right)V$
-
 \section{Experiments}
@@
 			Input sizes & [15, 7] & [?, ?, 7] & [1-15, 7] \\
-			Layers sizes & [128, 128] &  [64, 64] & Encoder: [64, 64] \\
-			& & & Attention: $d_k=64$\\
-			& & & Decoder: [64, 64]\\
+			Layers sizes & [128, 128] &  [64, 64] & \makecell[l]{Encoder: [64, 64] \\Attention: 2 heads, $d=32$ \\ Decoder: [64, 64]} \\
 			Number of parameters & 3.0e4 & ? & 3.4e4 \\

### 2019-09-11 main.tex
Update manuscript: agents

@@
 In the imitation learning approach, a policy can be trained in a supervised manner to imitate human driving decisions \citep[e.g.][]{Pomerleau1989, Ross2011, Bojarski2016, Xu2016, Eraqi2017, Codevilla2017, Rehder2017c, Rezagholiradeh2018, Rhinehart2018, Bansal2018, Rhinehart2019}.
-Because the cost of human driving data collection at large scale can be prohibitive, another promising approach is train a policy in simulation using reinforcement learning \citep[e.g.][]{Cardamone2009, Ross2011, Mukadam2017, Chen2017, Isele_2018, Ha2018, Kendall2019}.
+Because the cost of human driving data collection at large scale can be prohibitive, another promising approach is train a policy in simulation using reinforcement learning \citep[e.g.][]{Cardamone2009, Ross2011, Mukadam2017, Chen2017, Isele2017, Isele2018, Ha2018, Kendall2019}.
 
@@
 \section{Background and Related Work}
+
+\label{sec:background}
 
@@
 When dealing with a continuous state space $S$, we need to employ function approximation in order to generalise to nearby states.
-The \emph{Deep Q-Network} (DQN) algorithm \citep{Mnih} implements this idea by using a neural network model to represent the action-value function $Q$.
+The \emph{Deep Q-Network} (DQN) algorithm \citep{Mnih2015} implements this idea by using a neural network model to represent the action-value function $Q$.
 
@@
 
-\begin{equation*}
+\begin{equation}
 s = \left\lbrace s_i \right\rbrace_{i \in [0, N]}\qquad
@@
 \end{bmatrix}^T
-\end{equation*}
+\label{eq:coordinates}
+\end{equation}
 
@@
 However, it lacks two important properties.
-First, its size varies with the number of vehicles which can be problematic for the sake of function approximation which often expects constant-sized inputs, but also for concerns of varying computational complexity.
+First, its size varies with the number of vehicles which can be problematic for the sake of function approximation which often expects constant-sized inputs.
 Second, we expect a driving policy $\pi$ to be \emph{permutation invariant}, i.e.
@@
 \section{Model Architecture}
+\label{sec:architecture}
 
@@
 We obtain a stochastic matrix called the \emph{attention matrix}, which is finally used to gather a set of output value $V = [v_0, \dots, v_N]$, where each value $v_i$ is a feature computed with a shared linear projection $L_v\in\Real^{d_x \times d_v}$.
-Overall, the ego-attention computation for each head can be written as:
+Overall, the attention computation for each head can be written as:
 \begin{equation}
@@
 The outputs from all heads are finally combined with a linear layer, and the resulting tensor is then added to the ego encoding as in residual networks.
-We can easily see that the result is permutation invariant: indeed, a permutation $\tau$ will change the order of the rows in keys $K$ and values $V$ in \eqref{eq:selfattention} but will keep their correspondence.
+We can easily see that this process is permutation invariant: indeed, a permutation $\tau$ will change the order of the rows in keys $K$ and values $V$ in \eqref{eq:selfattention} but will keep their correspondence.
 The final result is a dot product of values and key-similarities, which is independent of the ordering.
@@
 	\end{tikzpicture}
-	\caption{Block diagram of our model architecture.}
+	\caption{Block diagram of our model architecture. It is composed of several linear identical encoders, a stack of ego-attention heads, and a linear decoder.}
 	\label{fig:architecture}
@@
 
+	\draw[draw=black, dashed] (1.75cm, -5.5cm) rectangle (9.5cm,1.5cm);
 	\end{tikzpicture}
@@
 
-TODO, see \autoref{tab:agents}.
+We evaluate three different agents, whose characteristics are summarized in \autoref{tab:agents}.
+\begin{itemize}
+	\item \MLPC: a \emph{list of features} state representation is used, as described in Section \ref{sec:background}.
+The model is a simple multi-layer perceptron (MLP).
+Because this architecture requires a fixed-size input, we use zero-padding to fill the input tensor up to a maximum number $N=14$ of observed vehicles, and add an additional \emph{presence} feature to the coordinates described in \eqref{eq:coordinates} so as to identify active rows.
+	\item \MLPG: a \emph{spatial grid} representation is used, as described in Section \ref{sec:background}, with a $11 \times 11$ grid where each cell represents a $5\text{m}\times 5$m square.
+The model is a multi-layer perceptron.
+	\item \EgoAtt: a \emph{list of features} state representation is used along with the Ego-Attention architecture described in Section \ref{sec:architecture}.
+As this model supports varying-size inputs, zero-padding is not required.
+\end{itemize}
 
@@
 			\midrule 
-			Input sizes & [15, 7] & [?, ?, 7] & [1-15, 7] \\
-			Layers sizes & [128, 128] &  [64, 64] & \makecell[l]{Encoder: [64, 64] \\Attention: 2 heads, $d=32$ \\ Decoder: [64, 64]} \\
-			Number of parameters & 3.0e4 & ? & 3.4e4 \\
-			Variable input size & No & No & Yes  \\
-			Permutation invariant & No & Yes & Yes \\
+			Input sizes & [15, 7] & [11, 11, 5] & [1-15, 7] \\
+			Layers sizes & [128, 128] &  [48, 48] & \makecell[tl]{Encoder: [64, 64] \\Attention: 2 heads, $d_k=32$ \\ Decoder: [64, 64]} \\
+			Number of parameters & 3.0e4 & 3.2e4 & 3.4e4 \\
+			Variable input size & No & No &  {Yes}  \\
+			Permutation invariant & No & {Yes} &  {Yes} \\
 			\bottomrule
@@
 
+These agents are all trained with the DQN algorithm using the same hyperparameters, and their architectures are scaled to admit about the same number of trainable parameters for fair comparison.
+
 \paragraph{Performances}
 
-We show in \autoref{fig:results} the evolution of the total reward, episode length and average velocity.
-The \MLPC agent learns to accelerate to earn short-term rewards, as shown by its high average velocity , but fails to exploit the information of other vehicles and crashes often, leading to shorter episodes = risky policy, worst performing.
-Conversely, the \MLPG benefits from its invariance to permutations and manages to learn to brake upon arrival at the intersection to avoid collisions (increased duration).
-However, it only proceeds when the intersection has been cleared (low mean velocity), resulting in an overly cautious policy, a common trait known as the freezing robot problem.
-Slight increase in performance.
+We plot in \autoref{fig:results} the evolution of the total reward, episode length and average velocity during training, over 4000 episodes and repeated across 120 random seeds.
+The \MLPC agent learns to accelerate to earn short-term rewards, as shown by its high average velocity, but fails to exploit the information of other vehicles and crashes often, leading to shorter episodes.
+In short, we obtain a risky and blind policy that is the worst performing.
+Conversely, the \MLPG architecture benefits from its invariance to permutations and manages to learn to brake upon arrival at the intersection to avoid collisions, as we can see from its higher episode length.
+However, it only proceeds when the intersection has been fully cleared, as reflected by its low average velocity.
+This results in an overly cautious policy -- a common trait colloquially known as the \emph{freezing robot problem} \citep{Trautman2010} -- but a slight increase in performance.
 In stark contrast, the \EgoAtt policy quickly learns both when it must slow down at the intersection but also when it can exploit the gaps in the traffic and take way to vehicles that are far/slow enough.

### 2019-09-11 main.tex
Updates from Overleaf

--- 2019-09-10 main.tex
@@
 \begin{abstract}
-  This work addresses the design of learning architectures for behavioural planning in a dense traffic setting.
-We claim that three important requirements are: dealing with a varying number of nearby vehicles, being invariant to the ordering chosen to describe them, while staying accurate and compact.
-We observe that the two most popular representations in the literature do not fit all these criteria, and show that they perform badly on an intersection-crossing task that involves complex anticipation and negotiation skills.
-In contrast, we propose an attention-based architecture that enjoys all these properties while also accounting for the existing interactions between the traffic participants.
-We show that this architecture leads to significant improvements in performance, and that it is able to capture interactions patterns that can be visualized and qualitatively interpreted.
+  TODO
 \end{abstract}
@@
 Indeed, the vast majority of existing systems still rely on hand-crafted rules encoded as Finite State Machines \citep{Paden2016}.
-As a result, only a narrow set of specified use-cases are addressed as these methods cannot scale to more complex scenes, especially when the decision-making process involves interacting with other human drivers whose behaviours are uncertain and difficult to model explicitly.
+As a result, only a narrow set of specified use-cases are addressed as these methods cannot scale to more complex scenes, especially when the decision-making process involves interacting with other human drivers whose behaviour is uncertain and difficult to model explicitly.
 
@@
 
-Beyond the choice of reinforcement learning algorithm, the formalization of the problem as a Markov Decision Process plays an important part in the design of the system.
-Indeed, the definition of the state space involves choosing a representation of the driving scene.
-In this work, we focus in on how the vehicles are represented.
-In particular, we claim that the two most-widely used representations both suffer from different drawbacks: on the one hand, the \emph{list of features} representation is compact and accurate but has a varying-size and depends on the choice of ordering.
-On the other hand, the \emph{spatial grid} representation addresses these concerns but in return suffers from an accuracy-size trade-off.
+Beyond the choice of reinforcement learning algorithm, the formalization of the problem as a Markov Decision Process (MDP) plays an important part in the design of the system.
+In particular, the choice of state and action spaces can have a significant influence on the performance of resulting policy.
+Describing a driving scene typically involves choosing a representation the road and vehicles.
+The simplest form: a list of vehicles with their coordinates, orientations, velocities.
+But variable size and order-dependent.
+Another common format is a spatial matrix, either Cartesian (images, occupancy grid) or polar (distances in angular sectors).
+Fixed size and order independent, but accuracy/size tradeoff.
+Challenge: keep a small state size, but enforce that the model is invariant to permutations.
 
-Our contributions are the following: first, we propose an attention-based architecture for decision-making involving social interactions, that deals with the varying-size and that accounts for two-way interactions between the ego-vehicle and every other traffic participant
+Contributions:
+- We propose an attention-based architecture for decision-making involving social interactions, that accounts for two-way interactions between the ego-vehicle and every other traffic participant
 - We evaluate our model on a challenging intersection-crossing task involving up to 15 vehicles perceived simultaneously.
@@
 
-\paragraph{State-representation for social interactions}{
+\paragraph{State-representation for social interactions}
 
-In order to apply a reinforcement learning algorithm such as DQN to an autonomous driving problem, a state space $S$ must first be chosen, that is, a representation of the scene.
 A vehicle driving on a road can be described in the most general way by it's continuous position, heading and velocity.
@@
 	\includegraphics[width=0.25\textwidth]{img/map}
-	\caption{The \emph{list of features} (left) and \emph{spatial grid} (right) representations}
-	\label{fig:representation}
+	\caption{The coordinates (left) and spatial-grid (right) representations}
 \end{figure}
 
-This representation, that we call \emph{list of features}, is illustrated in \autoref{fig:representation} and was used e.g.
-in \citep{Bai2015, Gindele2015, Song2016, Sunberg2017, Paxton2017, Galceran2017, Chen2017}.
+For instance, this representation is used in \citep{Bai2015, Gindele2015, Song2016, Sunberg2017, Paxton2017, Galceran2017, Chen2017}.
 
@@
 			Layers sizes & [128, 128] &  [64, 64] & Encoder: [64, 64] \\
-			& & & Attention: $d_k=64$\\
+			& & & Attention $d_k$: 64\\
 			& & & Decoder: [64, 64]\\
 			Number of parameters & 3.0e4 & ? & 3.4e4 \\
-			Variable input size & No & No & Yes \\
 			Permutation invariant & No & Yes & Yes \\
@@
 
-\begin{figure}[tp]
+\begin{figure}[htp]
 	\centering
@@
 
-This model is also more interpretable, as we can visualize the learnt attention matrix.
-
-TODO: figure yield vs take way
+TODO: figure
 
@@
 
-change the priority -> influences the rules of interactions -> changes the learnt behaviour.
 By switching the priority road, the resulting policy changes: In the exact same initial state, the ego-vehicle takes way to incoming vehicles when its policy was trained on a priority road, and gives way otherwise.
@@
 
+\section*{References}
+
 \bibliographystyle{named}

### 2019-09-11 main.tex
Updates from Overleaf

--- 2019-09-11 main.tex
@@
 In the imitation learning approach, a policy can be trained in a supervised manner to imitate human driving decisions \citep[e.g.][]{Pomerleau1989, Ross2011, Bojarski2016, Xu2016, Eraqi2017, Codevilla2017, Rehder2017c, Rezagholiradeh2018, Rhinehart2018, Bansal2018, Rhinehart2019}.
-Because the cost of human driving data collection at large scale can be prohibitive, another promising approach is train a policy in simulation using reinforcement learning \citep[e.g.][]{Cardamone2009, Ross2011, Mukadam2017, Chen2017, Isele2017, Isele2018, Ha2018, Kendall2019}.
+Because the cost of human driving data collection at large scale can be prohibitive, another promising approach is train a policy in simulation using reinforcement learning \citep[e.g.][]{Cardamone2009, Ross2011, Mukadam2017, Chen2017, Isele2018, Ha2018, Kendall2019}.
 
@@
 
-\paragraph{State-representation for social interactions}{
+\paragraph{State-representation for social interactions}
 

### 2019-09-13 main.tex
Update attention interpretation

@@
 They are then fed to an ego-attention layer, composed of several heads stacked together.
-Such an attention head is illustrated in \autoref{fig:ego-attention}, and works in the following way: in order to select a subset of vehicles depending on the context, the ego-vehicle  first emits a query $Q = [q_0]\in\Real^{1 \times d_k}$, computed with a linear projection $L_q\in\Real^{d_x \times d_k}$ of its embedding.
+Such an attention head is illustrated in \autoref{fig:ego-attention}, and works in the following way: in order to select a subset of vehicles depending on the context, the ego-vehicle  first emits a single query $Q = [q_0]\in\Real^{1 \times d_k}$, computed with a linear projection $L_q\in\Real^{d_x \times d_k}$ of its embedding.
 This query is then compared to a set of keys $K = [k_0, \dots, k_N]\in\Real^{N \times d_k}$ containing descriptive features $k_i$ for each vehicle, again computed with a shared linear projection $L_k\in\Real^{d_x \times d_k}$.
@@
 	\centering
-	\begin{subfigure}[t]{.75\linewidth}
+	\begin{subfigure}[t]{.6\linewidth}
 		\centering\includegraphics[width=\linewidth]{img/total_reward}
@@
 
-This model is also more interpretable, as we can visualize the learnt attention matrix.
+In any given state, the attention matrix can be visualized in the following way: we connect the ego-vehicle to every each vehicle by a line with a width proportional to the corresponding attention weight.
+Since the architecture can contain several Ego-Attention heads, we use different colours to distinguish them.
+In our experiments, we used two attention heads represented in green and blue.
+We observe that both heads exhibit a common behaviour: they focus their attention on incoming vehicles that are likely to collide with the ego-vehicle, depending on their current position, heading, velocity, and ignore those that are too far or in a conflict-less situation.
+However, we can also notice that the two heads have specialized to focus on different areas: the green head is only watching the vehicles coming from the left, while the blue head restricts itself to vehicles in the front and right directions.
 
-TODO: figure yield vs take way
+\begin{figure}
+	\centering
+	\begin{minipage}{.45\textwidth}
+		\centering
+		\includegraphics[width=\linewidth]{img/head_specialization}
+		\captionof{figure}{Specialization of the attention heads}
+		\label{fig:heads}
+	\end{minipage}
+\hfill
+	\begin{minipage}{.5\textwidth}
+		\centering
+		\includegraphics[width=\linewidth]{img/distances}
+		\captionof{figure}{Generally, the attention increases as the vehicles get closer.}
+		\label{fig:distance}
+	\end{minipage}
+\end{figure}
 
-Additional videos are available at \url{http://url.github.io}
+\begin{figure}[tp]
+	\centering
+	\begin{subfigure}[t]{.4\linewidth}
+		\includegraphics[width=\linewidth]{img/watch1}
+	\end{subfigure}
+	\begin{subfigure}[t]{.4\linewidth}
+		\includegraphics[width=\linewidth]{img/watch2}
+	\end{subfigure}
+	\caption{Sensitivity to uncertainty.
+\textbf{Right}: the agent has stopped at the intersection, its attention is focused on an incoming vehicle whose destination is still uncertain.
+\textbf{Left}: as soon as the incoming vehicle changes orientation, revealing a right turn, the attention drops and the agent starts accelerating.}
+	\label{fig:sensitivity}
+\end{figure}
+
+\begin{figure}[tp]
+	\centering
+	\begin{subfigure}[t]{.32\linewidth}
+		\includegraphics[width=\linewidth]{img/episode1}
+	\end{subfigure}
+	\begin{subfigure}[t]{.32\linewidth}
+		\includegraphics[width=\linewidth]{img/episode2}
+	\end{subfigure}
+	\begin{subfigure}[t]{.32\linewidth}
+		\includegraphics[width=\linewidth]{img/episode3}
+	\end{subfigure}
+\\[0.12em]
+	\begin{subfigure}[t]{.32\linewidth}
+		\includegraphics[width=\linewidth]{img/episode4}
+	\end{subfigure}
+	\begin{subfigure}[t]{.32\linewidth}
+		\includegraphics[width=\linewidth]{img/episode5}
+	\end{subfigure}
+	\begin{subfigure}[t]{.32\linewidth}
+		\includegraphics[width=\linewidth]{img/episode6}
+	\end{subfigure}
+	\caption{A complete episode.
+From left to right: 1.
+The green and blue heads direct their attentions to the left and front vehicles, respectively.
+2.
+The left-vehicle is passing and is no longer a threat 3.
+Immediately, the green attention head switches to the next vehicle coming from the left.
+4.
+The front vehicle has now passed, and the blue attention head is now focused on the ego-vehicle.
+5.
+The ego-vehicle waits for one last vehicle coming from the left.
+6.
+The ego-vehicle can finally proceed, and its attention is focused on itself.}
+\end{figure}
 
 \paragraph{Exploiting interaction patterns}
+
 The priority rules are note enforced through rewards but interactions: based on the defined road priorities, some vehicles will take way to the ego-vehicle, while others will not.
@@
 
+TODO: figure yield vs take way
+
+\section{Discussion}
+
+We trained a policy for left-turns, which is the hardest one as it had both the most conflict points and the lowest priority level.
+We also trained individual policies for right turns and straight crossings, and obtained similar results.
+When training a joint intersection policy, it is less efficient as it has no prior information on where the ego-vehicle is going.
+A generic intersection could be trained, but additional information must be added as additional features in \eqref{eq:coordinates}.
+For instance, the destination can be encoded as a desired direction $d_x, d_y$.
+This destination feature could also be used for other vehicles to encode blinker information when available.
+This should result in a more efficient policy.
+
 \section{Conclusion}
 
-TODO
+In this work, we showed that the \emph{list of features} representation, commonly used to describe vehicles in autonomous driving literature, is not tailored for use in a function approximation setting, in particular with neural networks.
+These concerns can be addressed by the \emph{spatial grid} representation, but the latter comes at the price of an increased input size and loss of accuracy.
+In contrast, we proposed an attention-based neural network architecture to tackle the aforementioned issues without compromising either size or accuracy.
+This architecture enjoys a better performance on a simulated negotiation and intersection crossing task.
 

### 2019-09-13 main.tex
Update experiments

@@
 \usepackage{multirow}
+
+\definecolor{mylinkcolor}{HTML}{e83f6f}
+\definecolor{mycitecolor}{HTML}{2980B9}
+\definecolor{myurlcolor}{HTML}{d63230}
+% \definecolor{myalgocolor}{HTML}{32936f}
+\definecolor{myalgocolor}{HTML}{000000}
+\hypersetup{
+	linkcolor  = mylinkcolor,
+	citecolor  = mycitecolor,
+	urlcolor   = myurlcolor,
+	colorlinks = true,
+}
 
@@
 In order to apply a reinforcement learning algorithm such as DQN to an autonomous driving problem, a state space $S$ must first be chosen, that is, a representation of the scene.
-When social interactions are relevant to the decision, the state should at leat compose a description of every nearby vehicle.
+When social interactions are relevant to the decision, the state should at least compose a description of every nearby vehicle.
 A vehicle driving on a road can be described in the most general way by it's continuous position, heading and velocity.
@@
 They are then fed to an ego-attention layer, composed of several heads stacked together.
-Such an attention head is illustrated in \autoref{fig:ego-attention}, and works in the following way: in order to select a subset of vehicles depending on the context, the ego-vehicle  first emits a single query $Q = [q_0]\in\Real^{1 \times d_k}$, computed with a linear projection $L_q\in\Real^{d_x \times d_k}$ of its embedding.
+The \emph{ego} prefix highlights that it similar to a multi-head self-attention layer \citep{Vaswani2017} but with only a single output corresponding to the ego-vehicle.
+Such an ego-attention head is illustrated in \autoref{fig:ego-attention} and works in the following way: in order to select a subset of vehicles depending on the context, the ego-vehicle  first emits a single query $Q = [q_0]\in\Real^{1 \times d_k}$, computed with a linear projection $L_q\in\Real^{d_x \times d_k}$ of its embedding.
 This query is then compared to a set of keys $K = [k_0, \dots, k_N]\in\Real^{N \times d_k}$ containing descriptive features $k_i$ for each vehicle, again computed with a shared linear projection $L_k\in\Real^{d_x \times d_k}$.
@@
 However, it only proceeds when the intersection has been fully cleared, as reflected by its low average velocity.
-This results in an overly cautious policy -- a common trait colloquially known as the \emph{freezing robot problem} \citep{Trautman2010} -- but a slight increase in performance.
-In stark contrast, the \EgoAtt policy quickly learns both when it must slow down at the intersection but also when it can exploit the gaps in the traffic and take way to vehicles that are far/slow enough.
-The overall resulting behaviour is qualitatively more nuanced and human-like.
+This results in an overly cautious policy -- a common trait colloquially known as the \emph{freezing robot problem} \citep{Trautman2010} -- with a slight increase in performance.
+In stark contrast, the \EgoAtt policy quickly learns both when it must slow down at the intersection (high episode length), but also when it can exploit the gaps in the traffic and take way to vehicles that are far or slow enough (higher average velocity than \MLPG).
+This translates as a significant performance gap, and the overall resulting behaviour is qualitatively more nuanced and human-like.
 
@@
 Since the architecture can contain several Ego-Attention heads, we use different colours to distinguish them.
-In our experiments, we used two attention heads represented in green and blue.
-We observe that both heads exhibit a common behaviour: they focus their attention on incoming vehicles that are likely to collide with the ego-vehicle, depending on their current position, heading, velocity, and ignore those that are too far or in a conflict-less situation.
-However, we can also notice that the two heads have specialized to focus on different areas: the green head is only watching the vehicles coming from the left, while the blue head restricts itself to vehicles in the front and right directions.
+In our experiments, two attention heads were used and will be represented in green and blue.
+We observe in \autoref{fig:heads} that they specialized to focus on different areas: the green head is only watching the vehicles coming from the left, while the blue head restricts itself to vehicles in the front and right directions.
+However, we notice that both heads exhibit a common behaviour: they focus their attention on incoming vehicles that are likely to collide with the ego-vehicle, depending on their current position, heading, velocity, and ignore those that are too far or in a conflict-less situation.
+In particular, the attention tends to increase when vehicles get closer, as shown in \autoref{fig:distance}.
+It can also be very sensitive to small variations of the vehicle states, when
 
@@
 	\centering
-	\begin{minipage}{.45\textwidth}
+	\begin{minipage}{.44\textwidth}
 		\centering
 		\includegraphics[width=\linewidth]{img/head_specialization}
-		\captionof{figure}{Specialization of the attention heads}
+		\captionof{figure}{The attention heads specialized in different areas: left and front/right.}
 		\label{fig:heads}
@@
 \hfill
-	\begin{minipage}{.5\textwidth}
+	\begin{minipage}{.545\textwidth}
 		\centering
 		\includegraphics[width=\linewidth]{img/distances}
-		\captionof{figure}{Generally, the attention increases as the vehicles get closer.}
+		\captionof{figure}{The attention paid to a vehicle tends to increase as it gets closer.}
 		\label{fig:distance}
@@
 \textbf{Right}: the agent has stopped at the intersection, its attention is focused on an incoming vehicle whose destination is still uncertain.
-\textbf{Left}: as soon as the incoming vehicle changes orientation, revealing a right turn, the attention drops and the agent starts accelerating.}
+\textbf{Left}: as soon as the vehicle orientation changes, revealing its intention of turning right, the attention drops and the agent starts accelerating right away.}
 	\label{fig:sensitivity}
@@
 
-\section{Discussion}
+\paragraph{Goal conditioning}
 
-We trained a policy for left-turns, which is the hardest one as it had both the most conflict points and the lowest priority level.
-We also trained individual policies for right turns and straight crossings, and obtained similar results.
-When training a joint intersection policy, it is less efficient as it has no prior information on where the ego-vehicle is going.
-A generic intersection could be trained, but additional information must be added as additional features in \eqref{eq:coordinates}.
-For instance, the destination can be encoded as a desired direction $d_x, d_y$.
-This destination feature could also be used for other vehicles to encode blinker information when available.
-This should result in a more efficient policy.
+In the previous examples, we trained a policy tailored for left-turns only because it is the hardest direction with the most conflict points and the lowest priority level.
+Two individual policies tailored for right turns and driving straight can be trained as well, with similar results.
+Training a generic intersection policy would be less efficient without any prior information on where the ego-vehicle is headed.
+To remedy this problem, the destination could be added as additional features in \eqref{eq:coordinates}, for instance encoded as a desired direction $(d_x, d_y)$.
+This destination feature could also be used for other traffic participants to encode blinker information when available.
+This should result in a more efficient and generic policy.
 
@@
 In this work, we showed that the \emph{list of features} representation, commonly used to describe vehicles in autonomous driving literature, is not tailored for use in a function approximation setting, in particular with neural networks.
-These concerns can be addressed by the \emph{spatial grid} representation, but the latter comes at the price of an increased input size and loss of accuracy.
-In contrast, we proposed an attention-based neural network architecture to tackle the aforementioned issues without compromising either size or accuracy.
-This architecture enjoys a better performance on a simulated negotiation and intersection crossing task.
+These concerns can be addressed by the \emph{spatial grid} representation, but it comes at the price of an increased input size and loss of accuracy.
+In contrast, we proposed an attention-based neural network architecture to tackle the aforementioned issues of the \emph{list of features} representation without compromising either size or accuracy.
+This architecture enjoys a better performance on a simulated negotiation and intersection crossing task, and is also more interpretable thanks to the visualisation of the attention matrix.
+The resulting policy successfully learns to recognize and exploit the interaction patterns that govern the nearby traffic.
 

### 2019-09-13 paper/main.tex
Move to paper folder and add assets

--- 2019-09-13 main.tex
@@
 
+\AtBeginDocument{\def\sectionautorefname{Section}}%
+
 \input{mathdef}
@@
 
-	\pgfmathdeclarefunction{dnorm}{2}{%
-		\pgfmathparse{1/(#2*sqrt(2*pi))*exp(-((x-#1)^2)/(2*#2^2))}%
-	}
-
 \maketitle
@@
 Indeed, the vast majority of existing systems still rely on hand-crafted rules encoded as Finite State Machines \citep{Paden2016}.
-As a result, only a narrow set of specified use-cases are addressed as these methods cannot scale to more complex scenes, especially when the decision-making process involves interacting with other human drivers whose behaviours are uncertain and difficult to model explicitly.
+As a result, only a narrow set of specified use-cases are addressed and these methods cannot scale to more complex scenes, especially when the decision-making involves interacting with other human drivers whose behaviours are uncertain and difficult to model explicitly.
 
@@
 \begin{itemize}
-	\item \MLPC: a \emph{list of features} state representation is used, as described in Section \ref{sec:background}.
+	\item \MLPC: a \emph{list of features} state representation is used, as described in \autoref{sec:background}.
 The model is a simple multi-layer perceptron (MLP).
 Because this architecture requires a fixed-size input, we use zero-padding to fill the input tensor up to a maximum number $N=14$ of observed vehicles, and add an additional \emph{presence} feature to the coordinates described in \eqref{eq:coordinates} so as to identify active rows.
-	\item \MLPG: a \emph{spatial grid} representation is used, as described in Section \ref{sec:background}, with a $11 \times 11$ grid where each cell represents a $5\text{m}\times 5$m square.
+	\item \MLPG: a \emph{spatial grid} representation is used, as described in \autoref{sec:background}, with a $11 \times 11$ grid where each cell represents a $5\text{m}\times 5$m square.
 The model is a multi-layer perceptron.
-	\item \EgoAtt: a \emph{list of features} state representation is used along with the Ego-Attention architecture described in Section \ref{sec:architecture}.
+	\item \EgoAtt: a \emph{list of features} state representation is used along with the Ego-Attention architecture described in \autoref{sec:architecture}.
 As this model supports varying-size inputs, zero-padding is not required.
@@
 In stark contrast, the \EgoAtt policy quickly learns both when it must slow down at the intersection (high episode length), but also when it can exploit the gaps in the traffic and take way to vehicles that are far or slow enough (higher average velocity than \MLPG).
-This translates as a significant performance gap, and the overall resulting behaviour is qualitatively more nuanced and human-like.
+This translates as a significant performance improvement, and the overall resulting behaviour is qualitatively more nuanced and human-like.
 
@@
 
-In any given state, the attention matrix can be visualized in the following way: we connect the ego-vehicle to every each vehicle by a line with a width proportional to the corresponding attention weight.
+In any given state, the attention matrix can be visualized in the following way: we connect the ego-vehicle to every vehicle by a line of width proportional to the corresponding attention weight.
 Since the architecture can contain several Ego-Attention heads, we use different colours to distinguish them.
@@
 In particular, the attention tends to increase when vehicles get closer, as shown in \autoref{fig:distance}.
-It can also be very sensitive to small variations of the vehicle states, when
+It can also be very sensitive to small variations in the traffic state, as reflected in \autoref{fig:sensitivity}.
+A full episode showcasing interactions with several vehicles is shown in \autoref{fig:episode}.
 
-\begin{figure}
+\begin{figure}[tp]
 	\centering
@@
 	\caption{Sensitivity to uncertainty.
-\textbf{Right}: the agent has stopped at the intersection, its attention is focused on an incoming vehicle whose destination is still uncertain.
-\textbf{Left}: as soon as the vehicle orientation changes, revealing its intention of turning right, the attention drops and the agent starts accelerating right away.}
+\emph{Left}: the agent has stopped at the intersection, its attention is focused on an incoming vehicle whose destination is still uncertain.
+\emph{Right}: as soon as the vehicle orientation changes, revealing its intention of turning right, the attention drops and the agent starts accelerating right away.}
 	\label{fig:sensitivity}
@@
 	\centering
-	\begin{subfigure}[t]{.32\linewidth}
-		\includegraphics[width=\linewidth]{img/episode1}
-	\end{subfigure}
-	\begin{subfigure}[t]{.32\linewidth}
-		\includegraphics[width=\linewidth]{img/episode2}
-	\end{subfigure}
-	\begin{subfigure}[t]{.32\linewidth}
-		\includegraphics[width=\linewidth]{img/episode3}
-	\end{subfigure}
+		\includegraphics[width=.32\linewidth]{img/episode1}
+		\includegraphics[width=.32\linewidth]{img/episode2}
+		\includegraphics[width=.32\linewidth]{img/episode3}
 \\[0.12em]
-	\begin{subfigure}[t]{.32\linewidth}
-		\includegraphics[width=\linewidth]{img/episode4}
-	\end{subfigure}
-	\begin{subfigure}[t]{.32\linewidth}
-		\includegraphics[width=\linewidth]{img/episode5}
-	\end{subfigure}
-	\begin{subfigure}[t]{.32\linewidth}
-		\includegraphics[width=\linewidth]{img/episode6}
-	\end{subfigure}
+		\includegraphics[width=.32\linewidth]{img/episode4}
+		\includegraphics[width=.32\linewidth]{img/episode5}
+		\includegraphics[width=.32\linewidth]{img/episode6}
 	\caption{A complete episode.
-From left to right: 1.
+\emph{From left to right, top to bottom}: 1.
 The green and blue heads direct their attentions to the left and front vehicles, respectively.
@@
 The ego-vehicle can finally proceed, and its attention is focused on itself.}
+	\label{fig:episode}
 \end{figure}
@@
 
-The priority rules are note enforced through rewards but interactions: based on the defined road priorities, some vehicles will take way to the ego-vehicle, while others will not.
+The priority rules are not enforced through rewards but interactions: based on the defined road priorities, some vehicles will take way to the ego-vehicle while others will not.
+By changing which is a priority road, we can influence the rules of interactions which affects the learnt behaviour.
+In \autoref{fig:priority}, we compare two policies placed in the exact same initial state and observe how their decisions are affected by their model of how incoming vehicles interact with them.
+This difference showcases the ability of our proposed architecture to discover and exploit such interaction patterns.
 
-change the priority -> influences the rules of interactions -> changes the learnt behaviour.
-By switching the priority road, the resulting policy changes: In the exact same initial state, the ego-vehicle takes way to incoming vehicles when its policy was trained on a priority road, and gives way otherwise.
-The only change between these two situations is the rules of interactions imposed on simulated vehicles, which showcases the ability of our proposed architecture to discover and exploit such interaction patterns.
-
-TODO: figure yield vs take way
+ \begin{figure}[htp]
+ 	\centering
+ 	\begin{subfigure}[t]{.49\linewidth}
+ 		\includegraphics[width=\linewidth]{img/priority1}
+ 	\end{subfigure}
+ 	\begin{subfigure}[t]{.49\linewidth}
+ 		\includegraphics[width=\linewidth]{img/priority2}
+ 	\end{subfigure}
+ 	\caption{Effect of the right of way.
+\emph{Left}: when trained on a non-priority road, the agent learns to yield to incoming vehicles.
+\emph{Right}: when trained on a priority road, the agent expects other vehicles to give way and is consequently more aggressive.}
+ 	\label{fig:priority}
+ \end{figure}
 

### 2019-09-13 paper/main.tex
Add link to videos

@@
 \begin{abstract}
-  This work addresses the design of learning architectures for behavioural planning in a dense traffic setting.
-We claim that three important requirements are: dealing with a varying number of nearby vehicles, being invariant to the ordering chosen to describe them, while staying accurate and compact.
-We observe that the two most popular representations in the literature do not fit all these criteria, and show that they perform badly on an intersection-crossing task that involves complex anticipation and negotiation skills.
-In contrast, we propose an attention-based architecture that enjoys all these properties while also accounting for the existing interactions between the traffic participants.
-We show that this architecture leads to significant improvements in performance, and that it is able to capture interactions patterns that can be visualized and qualitatively interpreted.
+  We study the design of learning architectures for behavioural planning in a dense traffic setting.
+Such architectures should deal with a varying number of nearby vehicles, be invariant to the ordering chosen to describe them, while staying accurate and compact.
+We observe that the two most popular representations in the literature do not fit these criteria, and perform badly on an complex negotiation task.
+We propose an attention-based architecture that satisfies all these properties and explicitly accounts for the existing interactions between the traffic participants.
+We show that this architecture enjoys significant performance gains, and that it is able to capture interactions patterns that can be visualized and qualitatively interpreted.
+Videos and code are available at \url{https://social-attention.github.io/}
 \end{abstract}

### 2019-09-14 paper/main.tex
Update images

@@
 
-The optimal action-value function $Q^* =  \max_\pi Q^\pi(s)$ satisfies the Bellman Optimality Equation \citep{bellman}:
+The optimal action-value function $Q^* =  \max_\pi Q^\pi(s)$ satisfies the Bellman Optimality Equation:
 \begin{equation*}
@@
 
-As $Q^*$ is a fixed-point of the Bellman Operator $\cT$ -- which is a contraction --, it can be computed by applying $\cT$ iteratively in a fixed-point iteration fashion.
-The \emph{Q-learning} algorithm \citep{Watkins} follows this procedure by applying a sampling version $\cT$ to a batch of collected experience.
+As $Q^*$ is a fixed-point of the Bellman Operator $\cT$ \citep{Bellman56} -- which is a contraction --, it can be computed by applying $\cT$ in a fixed-point iteration fashion.
+The \emph{Q-learning} algorithm \citep{Watkins1992} follows this procedure by applying a sampling version $\cT$ to a batch of collected experience.
 When dealing with a continuous state space $S$, we need to employ function approximation in order to generalise to nearby states.
@@
 A popular way to address this limitations is to use a \emph{spatial grid} representation.
-Instead of explicitly representing spatial information as variables $x, y$ along with other features $f$ directly inside the state $\{s_i=(x_i,y_i,f_i)\}_{i\in[0,N]}$ indexed on the vehicles, they are instead represented implicitly through the layout of several feature variables $f_{ij}$ organized in a tensor structure, where the $(i,j)$ indexes refer to the cells of a quantization of the 2D-space.
+Instead of explicitly representing spatial information as variables $x, y$ along with other features $f$ directly inside a state $\{s_i=(x_i,y_i,f_i)\}_{i\in[0,N]}$ indexed on the vehicles, they are instead represented implicitly through the layout of several feature variables $f_{ij}$ organized in a tensor structure, where the $(i,j)$ indexes refer to a quantization of the 2D-space.
 This representation is illustrated in \autoref{fig:representation} (right).
-Note that the size of this tensor is related to the area covered divided by the length of the quantization, which reflects a trade-off between accuracy and dimensionality.
-In an occupancy grid, the $f$ features contains presence information (0-1) and additional channels such as velocity and heading, as in \citep[e.g.][]{Isele2017, Fridman2018, Bansal2018, Rehder2017c}.
+Note that the size of this tensor is related to the area covered divided by the quantization step, which reflects a trade-off between accuracy and dimensionality.
+In an occupancy grid, the $f$ features contains presence information (0-1) and additional channels such as velocity and heading, as in \citep[e.g.][]{Isele2018, Fridman2018, Bansal2018, Rehder2017c}.
 Another example is the use of top-view RGB images \citep[e.g.][]{Bagnell2010, Rehder2017, Rehder2017c, Liu2018}.
@@
 The scene -- composed of two roads crossing perpendicularly -- is populated with several traffic participants initialized with random positions, velocities, and destinations.
-As in \citep{highway-env}, these vehicle are simulated with the Kinematic Bicycle Model, their lateral control is achieved by a low-level steering controller tracking a target route, and their longitudinal behaviour follows the Intelligent Driver Model.
+As in \citep{highway-env}, these vehicle are simulated with the Kinematic Bicycle Model, their lateral control is achieved by a low-level steering controller tracking a target route, and their longitudinal behaviour follows the Intelligent Driver Model \citep{Treiber2000}.
 However, this model only considers same-lane interactions and special care was required to prevent lateral collisions at the intersection.
-To that end, we implemented the following simplistic behaviour: each vehicle predicts the future positions of its neighbours over a three-second horizon by using a constant velocity model.
-In case of predicted collision with a neighbour, the yielding vehicle is determined based on road priorities, and must brake until the collision prediction ceases.
+To that end, we implemented the following simplistic behaviour: each vehicle predicts the future positions of its neighbours over a three-seconds horizon by using a constant velocity model.
+In case of predicted collision with a neighbour, the yielding vehicle is determined based on road priorities and brakes until the collision prediction ceases.
 
@@
 \begin{itemize}
-	\item \MLPC: a \emph{list of features} state representation is used, as described in \autoref{sec:background}.
+	\item \MLPL: a \emph{list of features} state representation is used, as described in \autoref{sec:background}.
 The model is a simple multi-layer perceptron (MLP).
@@
 			\toprule
-			Architecture & \MLPC & \MLPG & \EgoAtt \\
+			Architecture & \MLPL & \MLPG & \EgoAtt \\
 			\midrule 
@@
 We plot in \autoref{fig:results} the evolution of the total reward, episode length and average velocity during training, over 4000 episodes and repeated across 120 random seeds.
-The \MLPC agent learns to accelerate to earn short-term rewards, as shown by its high average velocity, but fails to exploit the information of other vehicles and crashes often, leading to shorter episodes.
+The \MLPL agent learns to accelerate to earn short-term rewards, as shown by its high average velocity, but fails to exploit the information of other vehicles and crashes often, leading to shorter episodes.
 In short, we obtain a risky and blind policy that is the worst performing.

### 2019-09-14 paper/main.tex
Last changes

@@
 In order to apply a reinforcement learning algorithm such as DQN to an autonomous driving problem, a state space $S$ must first be chosen, that is, a representation of the scene.
-When social interactions are relevant to the decision, the state should at least compose a description of every nearby vehicle.
+When social interactions are relevant to the decision, the state should at least contain a description of every nearby vehicle.
 A vehicle driving on a road can be described in the most general way by it's continuous position, heading and velocity.
-Then, the joint state of a road traffic with one ego-vehicle denoted $s_0$ and $N$ other vehicles can be described by the (unordered) set of individual vehicle states:
+Then, the joint state of a road traffic with one ego-vehicle denoted $s_0$ and $N$ other vehicles can be described by a list of individual vehicle states:
 
 \begin{equation}
-s = \left\lbrace s_i \right\rbrace_{i \in [0, N]}\qquad
+s = \left( s_i \right)_{i \in [0, N]}\qquad
 \text{where}\qquad
@@
 Formally, we require that:
-\begin{equation*}
+\begin{equation}
+\label{eq:permutation}
 \pi(\cdot|(s_0, s_1,\dotsc,s_N)) = \pi(\cdot|(s_0, s_{\tau(1)},\dotsc,s_{\tau(N)})) \quad\quad\quad \forall\tau \in \mathfrak{S}_N
-\end{equation*}
+\end{equation}
 
@@
 
-This permutation invariance property can also be implemented within the policy architecture.
+This permutation invariance property \eqref{eq:permutation} can also be implemented within the architecture of the policy $\pi$.
 A general technique to achieve this is to treat each entity similarly in the early stages -- e.g.
 through weight sharing -- before reducing them with a projection operator that is itself invariant to permutations, for instance a max-pooling as in \citep{Chen2017} or an average as in \citep{Qi2016}.
-A particular instantiation of this idea is attention mechanisms.
+A particular instance of this idea is attention mechanisms.
 
@@
 In this context, the agent must drive a vehicle by controlling its acceleration chosen from a finite set of actions $A = \{\texttt{SLOWER}, \texttt{NO-OP}, \texttt{FASTER}\}$.
-The lateral control is performed automatically by a low-level controller, such that the problem complexity is focused on the high-level interactions with other vehicles, namely the decision to either yield or take way.
+The lateral control is performed automatically by a low-level controller, such that the problem complexity is focused on the high-level interactions with other vehicles, namely the decision to either give or take way.
 The agent is rewarded by $1$ when it drives at maximum velocity, $0$ otherwise, and by $-5$ when a collision occurs.
@@
 We plot in \autoref{fig:results} the evolution of the total reward, episode length and average velocity during training, over 4000 episodes and repeated across 120 random seeds.
-The \MLPL agent learns to accelerate to earn short-term rewards, as shown by its high average velocity, but fails to exploit the information of other vehicles and crashes often, leading to shorter episodes.
-In short, we obtain a risky and blind policy that is the worst performing.
+The \MLPL agent learns to accelerate to earn short-term rewards, as shown by its high average velocity, but fails to exploit the information of other vehicles and crashes often, leading to short episodes.
+We obtain a risky and blind policy that is the worst performing.
 Conversely, the \MLPG architecture benefits from its invariance to permutations and manages to learn to brake upon arrival at the intersection to avoid collisions, as we can see from its higher episode length.
@@
 This results in an overly cautious policy -- a common trait colloquially known as the \emph{freezing robot problem} \citep{Trautman2010} -- with a slight increase in performance.
-In stark contrast, the \EgoAtt policy quickly learns both when it must slow down at the intersection (high episode length), but also when it can exploit the gaps in the traffic and take way to vehicles that are far or slow enough (higher average velocity than \MLPG).
+In stark contrast, the \EgoAtt policy quickly learns both when it must slow down at the intersection (see the high episode length), but also when it can exploit the gaps in the traffic and take way to vehicles that are far or slow enough (see the higher average velocity than \MLPG).
 This translates as a significant performance improvement, and the overall resulting behaviour is qualitatively more nuanced and human-like.
@@
 In any given state, the attention matrix can be visualized in the following way: we connect the ego-vehicle to every vehicle by a line of width proportional to the corresponding attention weight.
-Since the architecture can contain several Ego-Attention heads, we use different colours to distinguish them.
+Since the architecture can contain several ego-attention heads, we use different colours to distinguish them.
 In our experiments, two attention heads were used and will be represented in green and blue.
 We observe in \autoref{fig:heads} that they specialized to focus on different areas: the green head is only watching the vehicles coming from the left, while the blue head restricts itself to vehicles in the front and right directions.
-However, we notice that both heads exhibit a common behaviour: they focus their attention on incoming vehicles that are likely to collide with the ego-vehicle, depending on their current position, heading, velocity, and ignore those that are too far or in a conflict-less situation.
+However, we notice that both heads exhibit a common behaviour: they direct their attention to incoming vehicles that are likely to collide with the ego-vehicle, depending on their current position, heading, velocity, and ignore those that are too far or in a conflict-less situation.
 In particular, the attention tends to increase when vehicles get closer, as shown in \autoref{fig:distance}.
@@
 
-The priority rules are not enforced through rewards but interactions: based on the defined road priorities, some vehicles will take way to the ego-vehicle while others will not.
+The agent decisions regarding right of way are not enforced through rewards but interactions: based on the defined road priorities, some vehicles will take way to the ego-vehicle while others will not.
 By changing which is a priority road, we can influence the rules of interactions which affects the learnt behaviour.
-In \autoref{fig:priority}, we compare two policies placed in the exact same initial state and observe how their decisions are affected by their model of how incoming vehicles interact with them.
+In \autoref{fig:priority}, we compare two policies placed in the exact same initial state and observe how their decisions are affected by their internal model of how incoming vehicles interact with them.
 This difference showcases the ability of our proposed architecture to discover and exploit such interaction patterns.

### 2019-11-26 paper/main.tex
Updates from Overleaf

@@
 % ready for submission
-\usepackage[nonatbib]{neurips_2019_ml4ad}
+% \usepackage[nonatbib]{neurips_2019_ml4ad}
 
@@
 % to compile a camera-ready version, add the [final] option, e.g.:
-% \usepackage[final]{neurips_2019_ml4ad}
+\usepackage[final, nonatbib]{neurips_2019_ml4ad}
 
@@
 We propose an attention-based architecture that satisfies all these properties and explicitly accounts for the existing interactions between the traffic participants.
-We show that this architecture enjoys significant performance gains, and that it is able to capture interactions patterns that can be visualized and qualitatively interpreted.
+We show that this architecture leads to significant performance gains, and is able to capture interactions patterns that can be visualised and qualitatively interpreted.
 Videos and code are available at \url{https://social-attention.github.io/}
@@
 A popular way to address this limitations is to use a \emph{spatial grid} representation.
-Instead of explicitly representing spatial information as variables $x, y$ along with other features $f$ directly inside a state $\{s_i=(x_i,y_i,f_i)\}_{i\in[0,N]}$ indexed on the vehicles, they are instead represented implicitly through the layout of several feature variables $f_{ij}$ organized in a tensor structure, where the $(i,j)$ indexes refer to a quantization of the 2D-space.
+Instead of explicitly representing spatial information as variables $x, y$ along with other features $f$ directly inside a state $\{s_i=(x_i,y_i,f_i)\}_{i\in[0,N]}$ indexed on the vehicles, they are instead represented implicitly through the layout of several feature variables $f_{ij}$ organised in a tensor structure, where the $(i,j)$ indexes refer to a quantisation of the 2D-space.
 This representation is illustrated in \autoref{fig:representation} (right).
-Note that the size of this tensor is related to the area covered divided by the quantization step, which reflects a trade-off between accuracy and dimensionality.
+Note that the size of this tensor is related to the area covered divided by the quantisation step, which reflects a trade-off between accuracy and dimensionality.
 In an occupancy grid, the $f$ features contains presence information (0-1) and additional channels such as velocity and heading, as in \citep[e.g.][]{Isele2018, Fridman2018, Bansal2018, Rehder2017c}.
@@
 The similarity between the query $q_0$ and any key $k_i$ is assessed by their dot product $q_0 k_i^T$.
-These similarities are then scaled by the inverse-square-root-dimension $1/\sqrt{d_k}$\footnote{This scaling is due to the fact that the dot-product of two independent random vectors with mean 0,  variance 1, and dimension $d_k$, is a random variable with mean 0 and variance $d_k$} and normalized with a softmax function $\sigma$ across vehicles.
+These similarities are then scaled by the inverse-square-root-dimension $1/\sqrt{d_k}$\footnote{This scaling is due to the fact that the dot-product of two independent random vectors with mean 0,  variance 1, and dimension $d_k$, is a random variable with mean 0 and variance $d_k$} and normalised with a softmax function $\sigma$ across vehicles.
 We obtain a stochastic matrix called the \emph{attention matrix}, which is finally used to gather a set of output value $V = [v_0, \dots, v_N]$, where each value $v_i$ is a feature computed with a shared linear projection $L_v\in\Real^{d_x \times d_v}$.
@@
 We propose a new task where vehicle-to-vehicle interaction plays a significant part: crossing a four-way intersection.
-The scene -- composed of two roads crossing perpendicularly -- is populated with several traffic participants initialized with random positions, velocities, and destinations.
+The scene -- composed of two roads crossing perpendicularly -- is populated with several traffic participants initialised with random positions, velocities, and destinations.
 As in \citep{highway-env}, these vehicle are simulated with the Kinematic Bicycle Model, their lateral control is achieved by a low-level steering controller tracking a target route, and their longitudinal behaviour follows the Intelligent Driver Model \citep{Treiber2000}.
@@
 
-We evaluate three different agents, whose characteristics are summarized in \autoref{tab:agents}.
+We evaluate three different agents, whose characteristics are summarised in \autoref{tab:agents}.
 \begin{itemize}
 	\item \MLPL: a \emph{list of features} state representation is used, as described in \autoref{sec:background}.
-The model is a simple multi-layer perceptron (MLP).
+The model is a simple fully-connected network (FCN).
 Because this architecture requires a fixed-size input, we use zero-padding to fill the input tensor up to a maximum number $N=14$ of observed vehicles, and add an additional \emph{presence} feature to the coordinates described in \eqref{eq:coordinates} so as to identify active rows.
-	\item \MLPG: a \emph{spatial grid} representation is used, as described in \autoref{sec:background}, with a $11 \times 11$ grid where each cell represents a $5\text{m}\times 5$m square.
-The model is a multi-layer perceptron.
+	\item \CNNG: a \emph{spatial grid} representation is used, as described in \autoref{sec:background}, with a $32 \times 32$ grid where each cell represents a $2\text{m}\times 2$m square.
+The model is a convolutional neural network (CNN).
 	\item \EgoAtt: a \emph{list of features} state representation is used along with the Ego-Attention architecture described in \autoref{sec:architecture}.
@@
 			\toprule
-			Architecture & \MLPL & \MLPG & \EgoAtt \\
+			Architecture & \MLPL & \CNNG & \EgoAtt \\
 			\midrule 
-			Input sizes & [15, 7] & [11, 11, 5] & [1-15, 7] \\
-			Layers sizes & [128, 128] &  [48, 48] & \makecell[tl]{Encoder: [64, 64] \\Attention: 2 heads, $d_k=32$ \\ Decoder: [64, 64]} \\
+			Input sizes & [15, 7] & [11, 11, 7] & [1-15, 7] \\
+			Layers sizes & [128, 128] &  \makecell[tc]{Convolutional layers: 3 \\ Kernel Size: 2 \\
+			Stride: 2 \\ Head: [20]} & \makecell[tl]{Encoder: [64, 64] \\Attention: 2 heads, $d_k=32$ \\ Decoder: [64, 64]} \\
 			Number of parameters & 3.0e4 & 3.2e4 & 3.4e4 \\
@@
 We obtain a risky and blind policy that is the worst performing.
-Conversely, the \MLPG architecture benefits from its invariance to permutations and manages to learn to brake upon arrival at the intersection to avoid collisions, as we can see from its higher episode length.
+Conversely, the \CNNG architecture benefits from its invariance to permutations and manages to learn to brake upon arrival at the intersection to avoid collisions, as we can see from its higher episode length.
 However, it only proceeds when the intersection has been fully cleared, as reflected by its low average velocity.
 This results in an overly cautious policy -- a common trait colloquially known as the \emph{freezing robot problem} \citep{Trautman2010} -- with a slight increase in performance.
-In stark contrast, the \EgoAtt policy quickly learns both when it must slow down at the intersection (see the high episode length), but also when it can exploit the gaps in the traffic and take way to vehicles that are far or slow enough (see the higher average velocity than \MLPG).
+In stark contrast, the \EgoAtt policy quickly learns both when it must slow down at the intersection (see the high episode length), but also when it can exploit the gaps in the traffic and take way to vehicles that are far or slow enough (see the higher average velocity than \CNNG).
 This translates as a significant performance improvement, and the overall resulting behaviour is qualitatively more nuanced and human-like.
@@
 
-In any given state, the attention matrix can be visualized in the following way: we connect the ego-vehicle to every vehicle by a line of width proportional to the corresponding attention weight.
+In any given state, the attention matrix can be visualised in the following way: we connect the ego-vehicle to every vehicle by a line of width proportional to the corresponding attention weight.
 Since the architecture can contain several ego-attention heads, we use different colours to distinguish them.
 In our experiments, two attention heads were used and will be represented in green and blue.
-We observe in \autoref{fig:heads} that they specialized to focus on different areas: the green head is only watching the vehicles coming from the left, while the blue head restricts itself to vehicles in the front and right directions.
+We observe in \autoref{fig:heads} that they specialised to focus on different areas: the green head is only watching the vehicles coming from the left, while the blue head restricts itself to vehicles in the front and right directions.
 However, we notice that both heads exhibit a common behaviour: they direct their attention to incoming vehicles that are likely to collide with the ego-vehicle, depending on their current position, heading, velocity, and ignore those that are too far or in a conflict-less situation.
@@
 		\includegraphics[width=\linewidth]{img/head_specialization}
-		\captionof{figure}{The attention heads specialized in different areas: left and front/right.}
+		\captionof{figure}{The attention heads specialised in different areas: left and front/right.}
 		\label{fig:heads}
@@
 This architecture enjoys a better performance on a simulated negotiation and intersection crossing task, and is also more interpretable thanks to the visualisation of the attention matrix.
-The resulting policy successfully learns to recognize and exploit the interaction patterns that govern the nearby traffic.
+The resulting policy successfully learns to recognise and exploit the interaction patterns that govern the nearby traffic.
 

### 2019-11-26 paper/poster/poster.tex
Add first version of poster

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% baposter Landscape Poster
% LaTeX Template
% Version 1.0 (15/5/13)
%
% Created by:
% Brian Amberg (baposter@brian-amberg.de)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[a0paper,portrait,fontscale=0.3, margin=2cm]{baposter}

\usepackage[utf8]{inputenc}
\usepackage[font=normalsize,labelfont=bf]{caption} % Required for specifying captions to tables and figures
\captionsetup{justification=centering}
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{lmodern}
\usepackage{relsize} % Used for making text smaller in some places

% Other packages
\usepackage{tabularx}
\def\tabularxcolumn#1{m{#1}}
\def\imagetop#1{\vtop{\null\hbox{#1}}}
\usepackage{adjustbox}
\usepackage{subfig}
\usepackage{floatrow} % sidecapfloat
\usepackage{setspace}
\usepackage{xspace}
\usepackage{noReferences}
\usepackage{caption}
\usepackage{tikz}
\usepackage[customcolors]{hf-tikz}
%\usepackage{subcaption}
\usepackage{epsf}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{dsfont}
\usepackage{enumitem}
\usepackage{float}
\usepackage{cases}
\usepackage{verbatim}
\usepackage{array}
\usepackage{graphicx}
\usepackage{graphics}
\usepackage{multirow}
\usepackage{listings}
%\usepackage[authoryear]{natbib}
\usepackage{authblk}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{eqparbox}
\usepackage{xcolor}
\usepackage[backend=biber, citestyle=authoryear, maxcitenames=2, maxbibnames=2]{biblatex}
\usepackage{lipsum}
\usepackage{svg}
\usepackage{cleveref}
\usepackage{tikz}
\usepackage{xspace}
\usetikzlibrary{arrows,automata}
\usepackage{array, makecell}
\usepackage{pgfplots}
\usetikzlibrary{intersections}

\bibliography{../budgeted_rl.bib}

\input{style}
\input{../mathdef}

\begin{document}

%\background{ % Set the background to an image (background.pdf)
%\begin{tikzpicture}[remember picture,overlay]
%\draw (current page.north west)+(-2em,2em) node[anchor=north west]
%{\includegraphics[height=1.1\textheight]{background.png}};
%\end{tikzpicture}
%}
\begin{poster}{
grid=false,
borderColor=bordercol, % Border color of content boxes
headerColorOne=headercol1, % Background color for the header in the content boxes (left side)
headerColorTwo=headercol2, % Background color for the header in the content boxes (right side)
headerFontColor=headerfontcol, % Text color for the header text in the content boxes
boxColorOne=boxcolor, % Background color for the content in the content boxes
headershape=roundedright, % Specify the rounded corner in the content box headers
%headerfont=\Large\sf\bf, % Font modifiers for the text in the content box headers
headerfont=\Large\bf\textsc, %Sans Serif
textborder=rectangle,
background=none,
headerborder=open, % Change to closed for a line under the content box headers
boxshade=plain,
textfont={\setlength{\parindent}{0.0em}\sffamily},
headerheight={0.05\textheight},
eyecatcher=true
%columns=5
}
%
%----------------------------------------------------------------------------------------
%	Title and authors
%----------------------------------------------------------------------------------------
%
{
\includegraphics[width=15em]{./img/companies}
}
{
Social Attention for Autonomous Decision-Making in Dense Traffic
}
{
Edouard Leurent, Jean Mercat
\vspace{-2\baselineskip}
}
{
\includegraphics[width=12em]{./img/inria_sc}
}

\setlength{\colheight}{0.92\textheight}

%----------------------------------------------------------------------------------------
%	Motivation
%----------------------------------------------------------------------------------------

\headerbox{\textsc{Motivation}}{name=motivation,span=1,column=0,row=0}{

\textbf{Behavioural Planning}

\textbf{Reinforcement Learning}

\begin{align*}
V^\pi(s) &\eqdef \expectedvalue\left[\sum_{t=0}^\infty \gamma^t R(s_t, a_t)\right]\\
Q^\pi(s, a) &\eqdef R(s, a) + \gamma \expectedvalueover{s'\sim P(s'|s, a)} V^\pi(s')
\end{align*}

The optimal action-value function $Q^* =  \max_\pi Q^\pi(s)$ satisfies the Bellman Optimality Equation:
\begin{equation*}
Q^*(s, a) = \expectedvalueover{s'\sim P(s'|s, a)} \max_{a'\in A} \left[R(s, a) + \gamma Q^*(s', a')\right]
\end{equation*}

}

%----------------------------------------------------------------------------------------
%	Setting
%----------------------------------------------------------------------------------------
\headerbox{\textsc{Setting}}{name=setting,span=1,column=0,row=0,below=motivation}{

\textbf{State representations for social interactions}

\begin{align*}
s = \left( s_i \right)_{i \in [0, N]}
\end{align*}
where
\begin{align*}
s_i = \begin{bmatrix}
x_i & y_i & v^x_i & v^y_i & \cos\psi_i & \sin \psi_i
\end{bmatrix}^T
\label{eq:coordinates}
\end{align*}

\begin{center}
	\includegraphics[width=0.35\textwidth]{../img/coordinates}

	List of features representation
\end{center}

\textbf{Requirements for function approximation}

\begin{enumerate}
	\item Permutation Invariance

\begin{align*}
\forall\tau \in \mathfrak{S}_N, \qquad &\pi(\cdot|(s_0, s_1,\dotsc,s_N)) = \\ &\pi(\cdot|(s_0, s_{\tau(1)},\dotsc,s_{\tau(N)}))
\end{align*}

\item Variable size

\end{enumerate}

\begin{center}
	\includegraphics[width=0.35\textwidth]{../img/map}

	Occupancy grid representation
\end{center}

}

%----------------------------------------------------------------------------------------
%	Model Architecture
%----------------------------------------------------------------------------------------
\headerbox{Model Architecture}{name=archi,column=1,span=2}{

\textbf{Self-Attention}
\begin{equation}
\text{output}=\underbrace{\sigma\left(\frac{QK^T}{\sqrt{d_k}}\right)}_{\text{attention matrix}}V
\label{eq:selfattention}
\end{equation}

\begin{center}

\begin{tikzpicture}
	\node(X1){ego};
	\node[below of=X1, node distance=0.7cm](X2){vehicle$_{1}$};
	\node[below of=X2, node distance=0.6cm](X3){$\vdots$};
	\node[below of=X3, node distance=0.7cm](X4){vehicle$_{N}$};

	\node[draw, right of=X1, node distance=1.6cm, rectangle](ENC1){Encoder};
	\node[draw, right of=X2, node distance=1.6cm, rectangle](ENC2){Encoder};
	\node[below of=ENC2, node distance=0.6cm](ENC3){$\vdots$};
	\node[draw, right of=X4, node distance=1.6cm, rectangle](ENC4){Encoder};

	\path (X1) edge (ENC1);
	\path (X2) edge (ENC2);
	\path (X4) edge (ENC4);

	\node[draw, rectangle, right of=ENC1, node distance=1.6cm, below=-0.4cm, fill=white](TRANS3){\rotatebox{90}{ Ego-attention }};
	\node[draw, rectangle, right of=ENC1, node distance=1.5cm, below=-0.3cm, fill=white](TRANS2){\rotatebox{90}{ Ego-attention }};
	\node[draw, rectangle, right of=ENC1, node distance=1.4cm, below=-0.2cm, fill=white](TRANS1){\rotatebox{90}{ Ego-attention }};

	\draw (ENC1.east) -| (TRANS1.west);
	\draw (ENC2.east) -| (TRANS1.west);
	\draw (ENC4.east) -| (TRANS1.west);

	\node[draw, right of=ENC1, node distance=2.8cm, rectangle](DEC1){Decoder};

	\draw (TRANS1.east) |- (DEC1.west);

	\node[right of=DEC1, node distance=1.2cm](Y1){Q};

	\draw (DEC1.east) -- (Y1.west);
\end{tikzpicture}
\begin{tikzpicture}[scale=0.7, every node/.style={scale=0.7}]
\node(X1){ego encoding};
\node[below of=X1, node distance=2cm](X2){$\vdots$};
\node[below of=X2, node distance=2cm](X3){vehicle$_{N}$ encoding};

\coordinate[right of= X1, node distance=2cm](X1b){};

\draw (X1) -- (X1b);

\node[draw, right of=X1b, node distance=1cm](LK1){$L_{k}$};
\node[draw, below of=LK1, node distance=1cm](LV1){$L_{v}$};
\node[draw, above of=LK1, node distance=1cm](LQ1){$L_{q}$};

\draw (X1b) -- (LQ1);
\draw (X1b) -- (LK1);
\draw (X1b) -- (LV1);

\node[right of=LQ1, node distance=1cm](Q1){$\mathbf{q}_0$};
\node[right of=LK1, node distance=1cm](K1){$\mathbf{k}_0$};
\node[right of=LV1, node distance=1cm](V1){$\mathbf{v}_0$};

\draw (LQ1) -- (Q1);
\draw (LK1) -- (K1);
\draw (LV1) -- (V1);

\coordinate[right of= X3, node distance=2cm](X3b){};

\draw (X3) -- (X3b);

\node[draw, right of=X3b, node distance=1cm](LK3){$L_{k}$};
\node[draw, below of=LK3, node distance=1cm](LV3){$L_{v}$};

\draw (X3b) -- (LK3);	
\draw (X3b) -- (LV3);

\node[right of=LK3, node distance=1cm](K3){$\mathbf{k}_{n}$};
\node[right of=LV3, node distance=1cm](V3){$\mathbf{v}_{n}$};

\draw (LK3) -- (K3);
\draw (LV3) -- (V3);

\coordinate[right of=Q1, node distance=0.3cm](TOP){};
\coordinate[right of=V3, node distance=0.3cm](BOT){};
\draw[decorate,decoration={brace}] (TOP) -- node[left=5pt]{} (BOT);

\node[right of=X2, text width=3cm, node distance=5.5cm](EQ){
};

\node[draw, right of=X2, node distance=8.3cm](EQ2){
	\footnotesize $\sigma\left(\frac{QK^T}{\sqrt{d_k}}\right)V$};

\coordinate[left of=EQ2, node distance=1.6cm, above=2.2cm](TOP2){};
\coordinate[left of=EQ2, node distance=1.5cm, below=0.0cm](MID2){};
\coordinate[left of=EQ2, node distance=1.6cm, below=2.2cm](BOT2){};
\draw[decorate,decoration={brace}] (TOP2) -- node[left=5pt]{} (BOT2);
\draw (MID2) -- (EQ2);

\node[right of=EQ2, node distance=2cm](OUT){output};
\draw (EQ2) -- (OUT);

\draw[draw=black, dashed] (1.75cm, -5.5cm) rectangle (9.5cm,1.5cm);
\end{tikzpicture}

\end{center}

}

%
%%----------------------------------------------------------------------------------------
%% Experiments
%%----------------------------------------------------------------------------------------
%
\headerbox{Experiments}{name=experiments,span=2,column=1,below=archi, above=bottom}{ 
\begin{center}
	\textbf{Agents}

	\begin{tabular}{lccc}
		\toprule
		Agent & \MLPL & \CNNG & \EgoAtt \\
		\midrule 
		Input sizes & [15, 7] & [11, 11, 7] & [1-15, 7] \\
		Layers sizes & [128, 128] &  \makecell[tc]{Convolutional layers: 3 \\ Kernel Size: 2 \\
			Stride: 2 \\ Head: [20]} & \makecell[tl]{Encoder: [64, 64] \\Attention: 2 heads, $d_k=32$ \\ Decoder: [64, 64]} \\
		Number of parameters & 3.0e4 & 3.2e4 & 3.4e4 \\
		Variable input size & No & No &  {Yes}  \\
		Permutation invariant & No & {Yes} &  {Yes} \\
		\bottomrule
	\end{tabular}
\end{center}

\begin{center}
	\textbf{Performances}

	\includegraphics[width=0.32\linewidth]{../img/total_reward}
	\includegraphics[width=0.32\linewidth]{../img/length}
	\includegraphics[width=0.32\linewidth]{../img/velocity}
\end{center}

\begin{center}
\textbf{Interpretation}

\includegraphics[width=0.26\linewidth]{../img/head_specialization}
\includegraphics[width=0.28\linewidth]{../img/distances}
\includegraphics[width=0.21\linewidth]{../img/watch1}
\includegraphics[width=0.21\linewidth]{../img/watch2}

Head specialisation \qquad\qquad\qquad Distance \qquad\qquad\qquad\qquad\qquad Sensitivity to uncertainty \qquad\qquad\qquad\qquad

\includegraphics[width=.159\linewidth]{../img/episode1}
\includegraphics[width=.159\linewidth]{../img/episode2}
\includegraphics[width=.159\linewidth]{../img/episode3}
\includegraphics[width=.159\linewidth]{../img/episode4}
\includegraphics[width=.159\linewidth]{../img/episode5}
\includegraphics[width=.159\linewidth]{../img/episode6}

A full episode

\includegraphics[width=0.2\linewidth]{../img/priority1}
\includegraphics[width=0.2\linewidth]{../img/priority2}

Exploiting interaction patterns
\end{center}

}

%----------------------------------------------------------------------------------------
%	Acknowledgements
%----------------------------------------------------------------------------------------
\headerbox{Acknowledgements}{name=ack,column=0,span=1,below=setting}{
This work has been supported by CPER Nord-Pas de Calais/FEDER DATA Advanced data science and technologies 2015-2020, the French Ministry of Higher Education and Research, INRIA, and the French Agence Nationale de la Recherche (ANR).
}

%----------------------------------------------------------------------------------------
%	References
%----------------------------------------------------------------------------------------
\headerbox{References}{name=refs,column=0,span=1,below=ack, above=bottom}{
    {
    \AtNextBibliography{\footnotesize}
    \setlength{\bibitemsep}{3pt}
    \printbibliography[heading=none]
    }
}

\end{poster}

\end{document}

### 2019-11-27 paper/main.tex
Update table

@@
 			\midrule 
-			Input sizes & [15, 7] & [11, 11, 7] & [1-15, 7] \\
+			Input sizes & [15, 7] & [32, 32, 7] & [~$\boldsymbol{\cdot}$~, 7] \\
 			Layers sizes & [128, 128] &  \makecell[tc]{Convolutional layers: 3 \\ Kernel Size: 2 \\
-			Stride: 2 \\ Head: [20]} & \makecell[tl]{Encoder: [64, 64] \\Attention: 2 heads, $d_k=32$ \\ Decoder: [64, 64]} \\
+			Stride: 2 \\ Head: [20]} & \makecell[tl]{Encoder: [64, 64] \\Attention: 2 heads\\\phantom{Attention: }$d_k=32$ \\ Decoder: [64, 64]} \\
 			Number of parameters & 3.0e4 & 3.2e4 & 3.4e4 \\

### 2019-11-27 paper/poster/img/architecture.tex
Update poster

--- 2019-11-26 paper/main.tex
@@
 
-\maketitle
-
-\begin{abstract}
-  We study the design of learning architectures for behavioural planning in a dense traffic setting.
-Such architectures should deal with a varying number of nearby vehicles, be invariant to the ordering chosen to describe them, while staying accurate and compact.
-We observe that the two most popular representations in the literature do not fit these criteria, and perform badly on an complex negotiation task.
-We propose an attention-based architecture that satisfies all these properties and explicitly accounts for the existing interactions between the traffic participants.
-We show that this architecture leads to significant performance gains, and is able to capture interactions patterns that can be visualised and qualitatively interpreted.
-Videos and code are available at \url{https://social-attention.github.io/}
-\end{abstract}
-
-\section{Introduction}
-
-In the last decades, the problem of \emph{behavioural planning} -- that is, high-level decision-making in the context of autonomous driving -- has arguably received less attention and seen less progress than the other components of the typical robotics pipeline: perception and control \citep{Gonzalez2016}.
-Indeed, the vast majority of existing systems still rely on hand-crafted rules encoded as Finite State Machines \citep{Paden2016}.
-As a result, only a narrow set of specified use-cases are addressed and these methods cannot scale to more complex scenes, especially when the decision-making involves interacting with other human drivers whose behaviours are uncertain and difficult to model explicitly.
-
-This observation has led the community to turn to learning-based methods, which bear the promise of leveraging data to automatically learn a complex driving policy.
-In the imitation learning approach, a policy can be trained in a supervised manner to imitate human driving decisions \citep[e.g.][]{Pomerleau1989, Ross2011, Bojarski2016, Xu2016, Eraqi2017, Codevilla2017, Rehder2017c, Rezagholiradeh2018, Rhinehart2018, Bansal2018, Rhinehart2019}.
-Because the cost of human driving data collection at large scale can be prohibitive, another promising approach is train a policy in simulation using reinforcement learning \citep[e.g.][]{Cardamone2009, Ross2011, Mukadam2017, Chen2017, Isele2018, Ha2018, Kendall2019}.
-
-Beyond the choice of reinforcement learning algorithm, the formalization of the problem as a Markov Decision Process plays an important part in the design of the system.
-Indeed, the definition of the state space involves choosing a representation of the driving scene.
-In this work, we focus in on how the vehicles are represented.
-In particular, we claim that the two most-widely used representations both suffer from different drawbacks: on the one hand, the \emph{list of features} representation is compact and accurate but has a varying-size and depends on the choice of ordering.
-On the other hand, the \emph{spatial grid} representation addresses these concerns but in return suffers from an accuracy-size trade-off.
-
-Our contributions are the following: first, we propose an attention-based architecture for decision-making involving social interactions.
-This architecture allows to satisfy the variable-size and permutation invariance requirements even when using a \emph{list of features} representation.
-It also naturally accounts for interactions between the ego-vehicle and any other traffic participant.
-Second, we evaluate our model on a challenging intersection-crossing task involving up to 15 vehicles perceived simultaneously.
-We show that our proposed method provides significant quantitative improvements, and that it enables to capture interaction patterns in a way that is visually interpretable.
-
-\section{Background and Related Work}
-
-\label{sec:background}
-
-\paragraph{Model-free deep reinforcement learning} Reinforcement Learning is a general framework for sequential decision-making under uncertainty.
-It frames the learning objective as the optimal control of a Markov Decision Process $(S, A, P, R, \gamma)$ with measurable state space $S$, action space $A$, unknown reward function $R\in\Real^{S \times A}$, and unknown dynamics $P\in \cM(S)^{S \times A}$, where $\cM(\mathcal{X})$ denotes the probability measures over a set $\mathcal{X}$.
-The objective is to find a policy $\pi\in\cM(A)^S$ with maximal expected $\gamma$-discounted cumulative reward, called the value function $V^\pi$.
-Formally,
-
-\begin{align*}
-V^\pi(s) &\eqdef \expectedvalue\left[\sum_{t=0}^\infty \gamma^t R(s_t, a_t)\condbar s_0=s, a_t\sim \pi(a_t|s_t), s_{t+1}\sim P(s_{t+1}|s_t, a_t)\right]\\
-Q^\pi(s, a) &\eqdef R(s, a) + \gamma \expectedvalueover{s'\sim P(s'|s, a)} V^\pi(s')
-\end{align*}
-
-The optimal action-value function $Q^* =  \max_\pi Q^\pi(s)$ satisfies the Bellman Optimality Equation:
-\begin{equation*}
-Q^*(s, a) = (\cT Q^*) (s, a) \eqdef \expectedvalueover{s'\sim P(s'|s, a)} \max_{a'\in A} \left[R(s, a) + \gamma Q^*(s', a')\right]
-\end{equation*}
-
-As $Q^*$ is a fixed-point of the Bellman Operator $\cT$ \citep{Bellman56} -- which is a contraction --, it can be computed by applying $\cT$ in a fixed-point iteration fashion.
-The \emph{Q-learning} algorithm \citep{Watkins1992} follows this procedure by applying a sampling version $\cT$ to a batch of collected experience.
-When dealing with a continuous state space $S$, we need to employ function approximation in order to generalise to nearby states.
-The \emph{Deep Q-Network} (DQN) algorithm \citep{Mnih2015} implements this idea by using a neural network model to represent the action-value function $Q$.
-
-\paragraph{State-representation for social interactions}
-
-In order to apply a reinforcement learning algorithm such as DQN to an autonomous driving problem, a state space $S$ must first be chosen, that is, a representation of the scene.
-When social interactions are relevant to the decision, the state should at least contain a description of every nearby vehicle.
-A vehicle driving on a road can be described in the most general way by it's continuous position, heading and velocity.
-Then, the joint state of a road traffic with one ego-vehicle denoted $s_0$ and $N$ other vehicles can be described by a list of individual vehicle states:
-
-\begin{equation}
-s = \left( s_i \right)_{i \in [0, N]}\qquad
-\text{where}\qquad
-s_i = \begin{bmatrix}
-x_i & y_i & v^x_i & v^y_i & \cos\psi_i & \sin \psi_i
-\end{bmatrix}^T
-\label{eq:coordinates}
-\end{equation}
-
-\begin{figure}[tp]
-	\centering
-	\includegraphics[width=0.25\textwidth]{img/coordinates}
-	\includegraphics[width=0.25\textwidth]{img/map}
-	\caption{The \emph{list of features} (left) and \emph{spatial grid} (right) representations}
-	\label{fig:representation}
-\end{figure}
-
-This representation, that we call \emph{list of features}, is illustrated in \autoref{fig:representation} (left) and was used for instance in \citep{Bai2015, Gindele2015, Song2016, Sunberg2017, Paxton2017, Galceran2017, Chen2017}.
-
-This encoding is efficient in the sense that it uses the smallest quantity of information necessary to represent the scene.
-However, it lacks two important properties.
-First, its size varies with the number of vehicles which can be problematic for the sake of function approximation which often expects constant-sized inputs.
-Second, we expect a driving policy $\pi$ to be \emph{permutation invariant}, i.e.
-not to be dependent on the order in which other traffic participants are listed.
-Ideally, this property should be enforced and not approximated by relying on the coverage of the $N!$ possible permutations $\tau$ of any given traffic state.
-Formally, we require that:
-\begin{equation}
-\label{eq:permutation}
-\pi(\cdot|(s_0, s_1,\dotsc,s_N)) = \pi(\cdot|(s_0, s_{\tau(1)},\dotsc,s_{\tau(N)})) \quad\quad\quad \forall\tau \in \mathfrak{S}_N
-\end{equation}
-
-A popular way to address this limitations is to use a \emph{spatial grid} representation.
-Instead of explicitly representing spatial information as variables $x, y$ along with other features $f$ directly inside a state $\{s_i=(x_i,y_i,f_i)\}_{i\in[0,N]}$ indexed on the vehicles, they are instead represented implicitly through the layout of several feature variables $f_{ij}$ organised in a tensor structure, where the $(i,j)$ indexes refer to a quantisation of the 2D-space.
-This representation is illustrated in \autoref{fig:representation} (right).
-Note that the size of this tensor is related to the area covered divided by the quantisation step, which reflects a trade-off between accuracy and dimensionality.
-In an occupancy grid, the $f$ features contains presence information (0-1) and additional channels such as velocity and heading, as in \citep[e.g.][]{Isele2018, Fridman2018, Bansal2018, Rehder2017c}.
-Another example is the use of top-view RGB images \citep[e.g.][]{Bagnell2010, Rehder2017, Rehder2017c, Liu2018}.
-
-This permutation invariance property \eqref{eq:permutation} can also be implemented within the architecture of the policy $\pi$.
-A general technique to achieve this is to treat each entity similarly in the early stages -- e.g.
-through weight sharing -- before reducing them with a projection operator that is itself invariant to permutations, for instance a max-pooling as in \citep{Chen2017} or an average as in \citep{Qi2016}.
-A particular instance of this idea is attention mechanisms.
-
-\paragraph{Attention mechanisms} {The attention architecture was introduced to enable neural networks to discover inter-dependencies within a variable number of inputs.
-It has been used for pedestrian trajectory forecasting in~\cite{Vemula2018} with spatiotemporal graphs and in~\cite{Sadeghian2019CVPR} with spatial and social attention using a generative neural network.
-In~\cite{Sadeghian2018ECCV}, attention over top-view road scene images for car trajectory forecasting is used.
-Multi-head attention mechanism has been developed in~\cite{Vaswani2017} for sentence translation.
-In~\cite{Messaoud2019} a mechanism called non-local multi-head attention is developed.
-However, this is a spatial attention that does not allow vehicle-to-vehicle attention.
-In the present work, we use a multi-head social attention mechanism to capture vehicle-to-ego dependencies and build varying input size and permutation invariance into the policy model.}
-
-\section{Model Architecture}
-\label{sec:architecture}
-
-Out of a complex scene description, the model should be able to filter information and consider only what is relevant for decision.
-In other words, the agent should \emph{pay attention} to vehicles that are close or conflict with the planned route.
-
-The proposed architecture is presented in \autoref{fig:architecture}.
-It is used to represent the $Q$-function that will be optimized by the DQN algorithm.
-It is composed of a first linear encoding layer whose weights are shared between all vehicles.
-At that point, the embeddings only contain individual features of size $d_x$.
-They are then fed to an ego-attention layer, composed of several heads stacked together.
-The \emph{ego} prefix highlights that it similar to a multi-head self-attention layer \citep{Vaswani2017} but with only a single output corresponding to the ego-vehicle.
-Such an ego-attention head is illustrated in \autoref{fig:ego-attention} and works in the following way: in order to select a subset of vehicles depending on the context, the ego-vehicle  first emits a single query $Q = [q_0]\in\Real^{1 \times d_k}$, computed with a linear projection $L_q\in\Real^{d_x \times d_k}$ of its embedding.
-This query is then compared to a set of keys $K = [k_0, \dots, k_N]\in\Real^{N \times d_k}$ containing descriptive features $k_i$ for each vehicle, again computed with a shared linear projection $L_k\in\Real^{d_x \times d_k}$.
-The similarity between the query $q_0$ and any key $k_i$ is assessed by their dot product $q_0 k_i^T$.
-These similarities are then scaled by the inverse-square-root-dimension $1/\sqrt{d_k}$\footnote{This scaling is due to the fact that the dot-product of two independent random vectors with mean 0,  variance 1, and dimension $d_k$, is a random variable with mean 0 and variance $d_k$} and normalised with a softmax function $\sigma$ across vehicles.
-We obtain a stochastic matrix called the \emph{attention matrix}, which is finally used to gather a set of output value $V = [v_0, \dots, v_N]$, where each value $v_i$ is a feature computed with a shared linear projection $L_v\in\Real^{d_x \times d_v}$.
-Overall, the attention computation for each head can be written as:
-\begin{equation}
-\text{output}=\underbrace{\sigma\left(\frac{QK^T}{\sqrt{d_k}}\right)}_{\text{attention matrix}}V
-\label{eq:selfattention}
-\end{equation}
-The outputs from all heads are finally combined with a linear layer, and the resulting tensor is then added to the ego encoding as in residual networks.
-We can easily see that this process is permutation invariant: indeed, a permutation $\tau$ will change the order of the rows in keys $K$ and values $V$ in \eqref{eq:selfattention} but will keep their correspondence.
-The final result is a dot product of values and key-similarities, which is independent of the ordering.
-
-\begin{figure}[tp]
-	\centering
 		\begin{tikzpicture}
@@
 
-	\node[right of=DEC1, node distance=2.3cm](Y1){action values};
+		\node[right of=DEC1, node distance=1.5cm](Y1){$Q$};
 
@@
 		\end{tikzpicture}
-	\caption{Block diagram of our model architecture. It is composed of several linear identical encoders, a stack of ego-attention heads, and a linear decoder.}
-	\label{fig:architecture}
 
-\end{figure}
-
-\begin{figure}[tp]
-	\centering
+		\clearpage
 		\begin{tikzpicture}[scale=1, every node/.style={scale=1}]
@@
 
-	\node[draw, right of=X2, node distance=8.3cm](EQ2){
-	\footnotesize $\sigma\left(\frac{QK^T}{\sqrt{d_k}}\right)V$};
+		\node[draw, right of=X2, node distance=8.1cm](EQ2){
+			\footnotesize $\underbrace{\sigma\left(\frac{QK^T}{\sqrt{d_k}}\right)}_{\text{attention matrix}}V$};
 
@@
 		\end{tikzpicture}
-	\caption{Architecture of an ego-attention head.
-		The blocks $L_{q}$, $L_{k}$, $L_{v}$ are linear layers. The keys $K$ and values $V$ are concatenated from all vehicles, while the query $Q$ is only produced by the ego-vehicle.}
-	\label{fig:ego-attention}
-\end{figure}
-
-\section{Experiments}
-\paragraph{Environment}
-
-In this application, we use the \href{https://github.com/eleurent/highway-env}{highway-env} environment \citep{highway-env} for simulated highway driving and behavioural decision-making.
-We propose a new task where vehicle-to-vehicle interaction plays a significant part: crossing a four-way intersection.
-The scene -- composed of two roads crossing perpendicularly -- is populated with several traffic participants initialised with random positions, velocities, and destinations.
-As in \citep{highway-env}, these vehicle are simulated with the Kinematic Bicycle Model, their lateral control is achieved by a low-level steering controller tracking a target route, and their longitudinal behaviour follows the Intelligent Driver Model \citep{Treiber2000}.
-However, this model only considers same-lane interactions and special care was required to prevent lateral collisions at the intersection.
-To that end, we implemented the following simplistic behaviour: each vehicle predicts the future positions of its neighbours over a three-seconds horizon by using a constant velocity model.
-In case of predicted collision with a neighbour, the yielding vehicle is determined based on road priorities and brakes until the collision prediction ceases.
-
-In this context, the agent must drive a vehicle by controlling its acceleration chosen from a finite set of actions $A = \{\texttt{SLOWER}, \texttt{NO-OP}, \texttt{FASTER}\}$.
-The lateral control is performed automatically by a low-level controller, such that the problem complexity is focused on the high-level interactions with other vehicles, namely the decision to either give or take way.
-The agent is rewarded by $1$ when it drives at maximum velocity, $0$ otherwise, and by $-5$ when a collision occurs.
-
-\paragraph{Agents}
-
-We evaluate three different agents, whose characteristics are summarised in \autoref{tab:agents}.
-\begin{itemize}
-	\item \MLPL: a \emph{list of features} state representation is used, as described in \autoref{sec:background}.
-The model is a simple fully-connected network (FCN).
-Because this architecture requires a fixed-size input, we use zero-padding to fill the input tensor up to a maximum number $N=14$ of observed vehicles, and add an additional \emph{presence} feature to the coordinates described in \eqref{eq:coordinates} so as to identify active rows.
-	\item \CNNG: a \emph{spatial grid} representation is used, as described in \autoref{sec:background}, with a $32 \times 32$ grid where each cell represents a $2\text{m}\times 2$m square.
-The model is a convolutional neural network (CNN).
-	\item \EgoAtt: a \emph{list of features} state representation is used along with the Ego-Attention architecture described in \autoref{sec:architecture}.
-As this model supports varying-size inputs, zero-padding is not required.
-\end{itemize}
-
-\begin{table}[tp]
-	\centering
-	\begin{threeparttable}
-		\caption{Characteristics of the agents}
-		\label{tab:agents}
-		\begin{tabular}{lccc}
-			\toprule
-			Architecture & \MLPL & \CNNG & \EgoAtt \\
-			\midrule 
-			Input sizes & [15, 7] & [11, 11, 7] & [1-15, 7] \\
-			Layers sizes & [128, 128] &  \makecell[tc]{Convolutional layers: 3 \\ Kernel Size: 2 \\
-			Stride: 2 \\ Head: [20]} & \makecell[tl]{Encoder: [64, 64] \\Attention: 2 heads, $d_k=32$ \\ Decoder: [64, 64]} \\
-			Number of parameters & 3.0e4 & 3.2e4 & 3.4e4 \\
-			Variable input size & No & No &  {Yes}  \\
-			Permutation invariant & No & {Yes} &  {Yes} \\
-			\bottomrule
-		\end{tabular}
-	\end{threeparttable}
-\end{table}
-
-These agents are all trained with the DQN algorithm using the same hyperparameters, and their architectures are scaled to admit about the same number of trainable parameters for fair comparison.
-
-\paragraph{Performances}
-
-We plot in \autoref{fig:results} the evolution of the total reward, episode length and average velocity during training, over 4000 episodes and repeated across 120 random seeds.
-The \MLPL agent learns to accelerate to earn short-term rewards, as shown by its high average velocity, but fails to exploit the information of other vehicles and crashes often, leading to short episodes.
-We obtain a risky and blind policy that is the worst performing.
-Conversely, the \CNNG architecture benefits from its invariance to permutations and manages to learn to brake upon arrival at the intersection to avoid collisions, as we can see from its higher episode length.
-However, it only proceeds when the intersection has been fully cleared, as reflected by its low average velocity.
-This results in an overly cautious policy -- a common trait colloquially known as the \emph{freezing robot problem} \citep{Trautman2010} -- with a slight increase in performance.
-In stark contrast, the \EgoAtt policy quickly learns both when it must slow down at the intersection (see the high episode length), but also when it can exploit the gaps in the traffic and take way to vehicles that are far or slow enough (see the higher average velocity than \CNNG).
-This translates as a significant performance improvement, and the overall resulting behaviour is qualitatively more nuanced and human-like.
-
-\begin{figure}[htp]
-	\centering
-	\begin{subfigure}[t]{.6\linewidth}
-		\centering\includegraphics[width=\linewidth]{img/total_reward}
-	\end{subfigure}
-\\
-	\begin{subfigure}[t]{.49\linewidth}
-		\centering\includegraphics[width=\linewidth]{img/length}
-	\end{subfigure}
-	\begin{subfigure}[t]{.49\linewidth}
-		\centering\includegraphics[width=\linewidth]{img/velocity}
-	\end{subfigure}
-%	\begin{subfigure}[t]{.49\linewidth}
-%	\centering\includegraphics[width=\linewidth]{img/distance}
- %   \end{subfigure}
- \caption{The episode total rewards, lengths, and average velocity (higher is better). We display the mean values -- along with their 95\% confidence interval -- averaged over 120 random seeds.}
- \label{fig:results}
-\end{figure}
-
-\paragraph{Attention interpretation}
-
-In any given state, the attention matrix can be visualised in the following way: we connect the ego-vehicle to every vehicle by a line of width proportional to the corresponding attention weight.
-Since the architecture can contain several ego-attention heads, we use different colours to distinguish them.
-In our experiments, two attention heads were used and will be represented in green and blue.
-We observe in \autoref{fig:heads} that they specialised to focus on different areas: the green head is only watching the vehicles coming from the left, while the blue head restricts itself to vehicles in the front and right directions.
-However, we notice that both heads exhibit a common behaviour: they direct their attention to incoming vehicles that are likely to collide with the ego-vehicle, depending on their current position, heading, velocity, and ignore those that are too far or in a conflict-less situation.
-In particular, the attention tends to increase when vehicles get closer, as shown in \autoref{fig:distance}.
-It can also be very sensitive to small variations in the traffic state, as reflected in \autoref{fig:sensitivity}.
-A full episode showcasing interactions with several vehicles is shown in \autoref{fig:episode}.
-
-\begin{figure}[tp]
-	\centering
-	\begin{minipage}{.44\textwidth}
-		\centering
-		\includegraphics[width=\linewidth]{img/head_specialization}
-		\captionof{figure}{The attention heads specialised in different areas: left and front/right.}
-		\label{fig:heads}
-	\end{minipage}
-\hfill
-	\begin{minipage}{.545\textwidth}
-		\centering
-		\includegraphics[width=\linewidth]{img/distances}
-		\captionof{figure}{The attention paid to a vehicle tends to increase as it gets closer.}
-		\label{fig:distance}
-	\end{minipage}
-\end{figure}
-
-\begin{figure}[tp]
-	\centering
-	\begin{subfigure}[t]{.4\linewidth}
-		\includegraphics[width=\linewidth]{img/watch1}
-	\end{subfigure}
-	\begin{subfigure}[t]{.4\linewidth}
-		\includegraphics[width=\linewidth]{img/watch2}
-	\end{subfigure}
-	\caption{Sensitivity to uncertainty.
-\emph{Left}: the agent has stopped at the intersection, its attention is focused on an incoming vehicle whose destination is still uncertain.
-\emph{Right}: as soon as the vehicle orientation changes, revealing its intention of turning right, the attention drops and the agent starts accelerating right away.}
-	\label{fig:sensitivity}
-\end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=.32\linewidth]{img/episode1}
-		\includegraphics[width=.32\linewidth]{img/episode2}
-		\includegraphics[width=.32\linewidth]{img/episode3}
-\\[0.12em]
-		\includegraphics[width=.32\linewidth]{img/episode4}
-		\includegraphics[width=.32\linewidth]{img/episode5}
-		\includegraphics[width=.32\linewidth]{img/episode6}
-	\caption{A complete episode.
-\emph{From left to right, top to bottom}: 1.
-The green and blue heads direct their attentions to the left and front vehicles, respectively.
-2.
-The left-vehicle is passing and is no longer a threat 3.
-Immediately, the green attention head switches to the next vehicle coming from the left.
-4.
-The front vehicle has now passed, and the blue attention head is now focused on the ego-vehicle.
-5.
-The ego-vehicle waits for one last vehicle coming from the left.
-6.
-The ego-vehicle can finally proceed, and its attention is focused on itself.}
-	\label{fig:episode}
-\end{figure}
-
-\paragraph{Exploiting interaction patterns}
-
-The agent decisions regarding right of way are not enforced through rewards but interactions: based on the defined road priorities, some vehicles will take way to the ego-vehicle while others will not.
-By changing which is a priority road, we can influence the rules of interactions which affects the learnt behaviour.
-In \autoref{fig:priority}, we compare two policies placed in the exact same initial state and observe how their decisions are affected by their internal model of how incoming vehicles interact with them.
-This difference showcases the ability of our proposed architecture to discover and exploit such interaction patterns.
-
- \begin{figure}[htp]
- 	\centering
- 	\begin{subfigure}[t]{.49\linewidth}
- 		\includegraphics[width=\linewidth]{img/priority1}
- 	\end{subfigure}
- 	\begin{subfigure}[t]{.49\linewidth}
- 		\includegraphics[width=\linewidth]{img/priority2}
- 	\end{subfigure}
- 	\caption{Effect of the right of way.
-\emph{Left}: when trained on a non-priority road, the agent learns to yield to incoming vehicles.
-\emph{Right}: when trained on a priority road, the agent expects other vehicles to give way and is consequently more aggressive.}
- 	\label{fig:priority}
- \end{figure}
-
-\paragraph{Goal conditioning}
-
-In the previous examples, we trained a policy tailored for left-turns only because it is the hardest direction with the most conflict points and the lowest priority level.
-Two individual policies tailored for right turns and driving straight can be trained as well, with similar results.
-Training a generic intersection policy would be less efficient without any prior information on where the ego-vehicle is headed.
-To remedy this problem, the destination could be added as additional features in \eqref{eq:coordinates}, for instance encoded as a desired direction $(d_x, d_y)$.
-This destination feature could also be used for other traffic participants to encode blinker information when available.
-This should result in a more efficient and generic policy.
-
-\section{Conclusion}
-
-In this work, we showed that the \emph{list of features} representation, commonly used to describe vehicles in autonomous driving literature, is not tailored for use in a function approximation setting, in particular with neural networks.
-These concerns can be addressed by the \emph{spatial grid} representation, but it comes at the price of an increased input size and loss of accuracy.
-In contrast, we proposed an attention-based neural network architecture to tackle the aforementioned issues of the \emph{list of features} representation without compromising either size or accuracy.
-This architecture enjoys a better performance on a simulated negotiation and intersection crossing task, and is also more interpretable thanks to the visualisation of the attention matrix.
-The resulting policy successfully learns to recognise and exploit the interaction patterns that govern the nearby traffic.
-
-%\subsubsection*{Acknowledgments}
-%
-%TODO
-
-\bibliographystyle{named}
-\bibliography{references}
 

### 2019-11-27 paper/poster/poster.tex
Update poster

@@
 \captionsetup{justification=centering}
+
 \usepackage{booktabs} % Horizontal rules in tables
@@
 \usepackage{noReferences}
-\usepackage{caption}
 \usepackage{tikz}
 \usepackage[customcolors]{hf-tikz}
-%\usepackage{subcaption}
 \usepackage{epsf}
@@
 \usetikzlibrary{intersections}
+\usepackage{pifont}
 
@@
 {
-\includegraphics[width=15em]{./img/companies}
+\includegraphics[width=10em]{./img/companies}
 }
@@
 Edouard Leurent, Jean Mercat
-\vspace{-2\baselineskip}
+\vspace{-4\baselineskip}
 }
@@
 
+\begin{itemize}
+	\item \hlg{High-level} decisions: lane, stop or go
+	\item In \hlr{interaction} with other drivers
+\end{itemize}
+
 \textbf{Reinforcement Learning}
 
-\begin{align*}
-V^\pi(s) &\eqdef \expectedvalue\left[\sum_{t=0}^\infty \gamma^t R(s_t, a_t)\right]\\
-Q^\pi(s, a) &\eqdef R(s, a) + \gamma \expectedvalueover{s'\sim P(s'|s, a)} V^\pi(s')
-\end{align*}
+\begin{equation*}
+Q^*(s, a) \eqdef \max_\pi \expectedvalue\left[\sum_{t=0}^\infty \gamma^t R(s_t, a_t)\condbar \substack{
+s_0=s\\a_0=a
+}
+\right]
+\end{equation*}
 
-The optimal action-value function $Q^* =  \max_\pi Q^\pi(s)$ satisfies the Bellman Optimality Equation:
+Bellman Optimality Equation:
 \begin{equation*}
@@
 \end{equation*}
-
 }
@@
 %----------------------------------------------------------------------------------------
-\headerbox{\textsc{Setting}}{name=setting,span=1,column=0,row=0,below=motivation}{
+\headerbox{\textsc{State Representation}}{name=setting,span=1,column=0,row=0,below=motivation}{
 
-\textbf{State representations for social interactions}
+\textbf{Representing social interactions}
 
+A joint state $s$ of $N+1$ observed vehicles
 \begin{align*}
-s = \left( s_i \right)_{i \in [0, N]}
-\end{align*}
-where
-\begin{align*}
-s_i = \begin{bmatrix}
+s &= \left( s_i \right)_{i \in [0, N]}\\
+s_i &= \begin{bmatrix}
 x_i & y_i & v^x_i & v^y_i & \cos\psi_i & \sin \psi_i
@@
 	\includegraphics[width=0.35\textwidth]{../img/coordinates}
-
-	List of features representation
+	\captionof*{figure}{List of features representation}
 \end{center}
 
-\textbf{Requirements for function approximation}
+\textbf{Issues related to function approximation}
 
 \begin{enumerate}
-	\item Permutation Invariance
+	\item \hlg{Variable} size
+    \begin{itemize}
+		\item[\incarrow] usual models accept \hlr{fixed-size} inputs
+	\end{itemize}
 
+	\item \hlr{Sensitivity} to the ordering
+	\begin{itemize}
+		\item we want the policy to be \hlg{permutation-invariant}:
+	\end{itemize}
 \begin{align*}
@@
 \end{align*}
-
-\item Variable size
-
 \end{enumerate}
 
+\textbf{A common solution}
 \begin{center}
 	\includegraphics[width=0.35\textwidth]{../img/map}
+	\captionof*{figure}{Occupancy grid representation}
+\end{center}
 
-	Occupancy grid representation
-\end{center}
+\begin{itemize}[nolistsep]
+	\item[\hlg{\checkmark}] Fixed-size
+	\item[\hlg{\checkmark}] Does not depend on an ordering
+	\item[\hlr{\xmark}] Suffers from an \hlr{accuracy / size tradeoff}
+\end{itemize}
 
@@
 
-\textbf{Self-Attention}
-\begin{equation}
-\text{output}=\underbrace{\sigma\left(\frac{QK^T}{\sqrt{d_k}}\right)}_{\text{attention matrix}}V
-\label{eq:selfattention}
-\end{equation}
-
 \begin{center}
-
-\begin{tikzpicture}
-	\node(X1){ego};
-	\node[below of=X1, node distance=0.7cm](X2){vehicle$_{1}$};
-	\node[below of=X2, node distance=0.6cm](X3){$\vdots$};
-	\node[below of=X3, node distance=0.7cm](X4){vehicle$_{N}$};
-
-	\node[draw, right of=X1, node distance=1.6cm, rectangle](ENC1){Encoder};
-	\node[draw, right of=X2, node distance=1.6cm, rectangle](ENC2){Encoder};
-	\node[below of=ENC2, node distance=0.6cm](ENC3){$\vdots$};
-	\node[draw, right of=X4, node distance=1.6cm, rectangle](ENC4){Encoder};
-
-	\path (X1) edge (ENC1);
-	\path (X2) edge (ENC2);
-	\path (X4) edge (ENC4);
-
-	\node[draw, rectangle, right of=ENC1, node distance=1.6cm, below=-0.4cm, fill=white](TRANS3){\rotatebox{90}{ Ego-attention }};
-	\node[draw, rectangle, right of=ENC1, node distance=1.5cm, below=-0.3cm, fill=white](TRANS2){\rotatebox{90}{ Ego-attention }};
-	\node[draw, rectangle, right of=ENC1, node distance=1.4cm, below=-0.2cm, fill=white](TRANS1){\rotatebox{90}{ Ego-attention }};
-
-	\draw (ENC1.east) -| (TRANS1.west);
-	\draw (ENC2.east) -| (TRANS1.west);
-	\draw (ENC4.east) -| (TRANS1.west);
-
-	\node[draw, right of=ENC1, node distance=2.8cm, rectangle](DEC1){Decoder};
-
-	\draw (TRANS1.east) |- (DEC1.west);
-
-	\node[right of=DEC1, node distance=1.2cm](Y1){Q};
-
-	\draw (DEC1.east) -- (Y1.west);
-\end{tikzpicture}
-\begin{tikzpicture}[scale=0.7, every node/.style={scale=0.7}]
-\node(X1){ego encoding};
-\node[below of=X1, node distance=2cm](X2){$\vdots$};
-\node[below of=X2, node distance=2cm](X3){vehicle$_{N}$ encoding};
-
-\coordinate[right of= X1, node distance=2cm](X1b){};
-
-\draw (X1) -- (X1b);
-
-\node[draw, right of=X1b, node distance=1cm](LK1){$L_{k}$};
-\node[draw, below of=LK1, node distance=1cm](LV1){$L_{v}$};
-\node[draw, above of=LK1, node distance=1cm](LQ1){$L_{q}$};
-
-\draw (X1b) -- (LQ1);
-\draw (X1b) -- (LK1);
-\draw (X1b) -- (LV1);
-
-\node[right of=LQ1, node distance=1cm](Q1){$\mathbf{q}_0$};
-\node[right of=LK1, node distance=1cm](K1){$\mathbf{k}_0$};
-\node[right of=LV1, node distance=1cm](V1){$\mathbf{v}_0$};
-
-\draw (LQ1) -- (Q1);
-\draw (LK1) -- (K1);
-\draw (LV1) -- (V1);
-
-\coordinate[right of= X3, node distance=2cm](X3b){};
-
-\draw (X3) -- (X3b);
-
-\node[draw, right of=X3b, node distance=1cm](LK3){$L_{k}$};
-\node[draw, below of=LK3, node distance=1cm](LV3){$L_{v}$};
-
-\draw (X3b) -- (LK3);	
-\draw (X3b) -- (LV3);
-
-\node[right of=LK3, node distance=1cm](K3){$\mathbf{k}_{n}$};
-\node[right of=LV3, node distance=1cm](V3){$\mathbf{v}_{n}$};
-
-\draw (LK3) -- (K3);
-\draw (LV3) -- (V3);
-
-\coordinate[right of=Q1, node distance=0.3cm](TOP){};
-\coordinate[right of=V3, node distance=0.3cm](BOT){};
-\draw[decorate,decoration={brace}] (TOP) -- node[left=5pt]{} (BOT);
-
-\node[right of=X2, text width=3cm, node distance=5.5cm](EQ){
-};
-
-\node[draw, right of=X2, node distance=8.3cm](EQ2){
-	\footnotesize $\sigma\left(\frac{QK^T}{\sqrt{d_k}}\right)V$};
-
-\coordinate[left of=EQ2, node distance=1.6cm, above=2.2cm](TOP2){};
-\coordinate[left of=EQ2, node distance=1.5cm, below=0.0cm](MID2){};
-\coordinate[left of=EQ2, node distance=1.6cm, below=2.2cm](BOT2){};
-\draw[decorate,decoration={brace}] (TOP2) -- node[left=5pt]{} (BOT2);
-\draw (MID2) -- (EQ2);
-
-\node[right of=EQ2, node distance=2cm](OUT){output};
-\draw (EQ2) -- (OUT);
-
-\draw[draw=black, dashed] (1.75cm, -5.5cm) rectangle (9.5cm,1.5cm);
-\end{tikzpicture}
+	\begin{minipage}{0.4\textwidth}
+		\includegraphics[trim={3.8cm 22.5cm 10cm 2.5cm}, clip, page=1, width=\linewidth]{img/architecture}
+		\captionof*{figure}{Model architecture}
+		\begin{itemize}
+			\item[\hlg{\checkmark}] Inputs can have a {variable size}
+			\item[\hlg{\checkmark}] Based on a dot product
+			\begin{itemize}
+				\item[$\incarrow$] {permutation-invariant}
+			\end{itemize}
+			\item[\hlg{\checkmark}] Compact size with no accuracy loss
+		\end{itemize}
+	\end{minipage}
+	\begin{minipage}{0.59\textwidth}
+		\includegraphics[trim={3.8cm 18.3cm 5.8cm 2.5cm}, clip, page=2, width=\linewidth]{img/architecture}
+		\captionof*{figure}{Ego-attention block}
+	\end{minipage}
 
@@
 \begin{center}
-	\textbf{Agents}
-
+	\captionof*{table}{\textbf{Agents}}
 	\begin{tabular}{lccc}
@@
 		\midrule 
-		Input sizes & [15, 7] & [11, 11, 7] & [1-15, 7] \\
+		Input sizes & [15, 7] & [32, 32, 7] & [~$\boldsymbol{\cdot}$~, 7] \\
 		Layers sizes & [128, 128] &  \makecell[tc]{Convolutional layers: 3 \\ Kernel Size: 2 \\
-			Stride: 2 \\ Head: [20]} & \makecell[tl]{Encoder: [64, 64] \\Attention: 2 heads, $d_k=32$ \\ Decoder: [64, 64]} \\
+			Stride: 2 \\ Head: [20]} & \makecell[tl]{Encoder: [64, 64] \\Attention: 2 heads\\\phantom{Attention: }$d_k=32$ \\ Decoder: [64, 64]} \\
 		Number of parameters & 3.0e4 & 3.2e4 & 3.4e4 \\
-		Variable input size & No & No &  {Yes}  \\
-		Permutation invariant & No & {Yes} &  {Yes} \\
+		Variable input size & \hlr{No} & \hlr{No} &  \hlg{Yes}  \\
+		Permutation invariant & \hlr{No} & \hlg{Yes} &  \hlg{Yes} \\
 		\bottomrule
@@
 \begin{center}
-	\textbf{Performances}
-
+	\captionof*{figure}{\textbf{Performances}}
 	\includegraphics[width=0.32\linewidth]{../img/total_reward}
+	\includegraphics[width=0.32\linewidth]{../img/velocity}
 	\includegraphics[width=0.32\linewidth]{../img/length}
-	\includegraphics[width=0.32\linewidth]{../img/velocity}
 \end{center}
@@
 \begin{center}
-\textbf{Interpretation}
+\textbf{Attention visualisation}
 
-\includegraphics[width=0.26\linewidth]{../img/head_specialization}
-\includegraphics[width=0.28\linewidth]{../img/distances}
-\includegraphics[width=0.21\linewidth]{../img/watch1}
-\includegraphics[width=0.21\linewidth]{../img/watch2}
+\begin{minipage}[b]{0.26\textwidth}
+\includegraphics[width=\linewidth]{../img/head_specialization}
+\captionof*{figure}{Head specialisation}
+\end{minipage}
+\begin{minipage}[b]{0.28\textwidth}
+	\includegraphics[width=\linewidth]{../img/distances}
+\captionof*{figure}{Distance}
+\end{minipage}
+\begin{minipage}[b]{0.42\textwidth}
+	\includegraphics[width=0.49\linewidth]{../img/watch1}
+	\includegraphics[width=0.49\linewidth]{../img/watch2}
+	\captionof*{figure}{Sensitivity to uncertainty}
+\end{minipage}
 
-Head specialisation \qquad\qquad\qquad Distance \qquad\qquad\qquad\qquad\qquad Sensitivity to uncertainty \qquad\qquad\qquad\qquad
-
+\begin{center}
 \includegraphics[width=.159\linewidth]{../img/episode1}
@@
 \includegraphics[width=.159\linewidth]{../img/episode6}
+\captionof*{figure}{A full episode}
+\end{center}
 
-A full episode
-
-\includegraphics[width=0.2\linewidth]{../img/priority1}
-\includegraphics[width=0.2\linewidth]{../img/priority2}
+\includegraphics[width=0.18\linewidth]{../img/priority1}
+\includegraphics[width=0.18\linewidth]{../img/priority2}
 
@@
 %----------------------------------------------------------------------------------------
-\headerbox{Acknowledgements}{name=ack,column=0,span=1,below=setting}{
+\headerbox{Acknowledgements}{name=ack,column=0,span=1,below=setting, above=bottom}{
 This work has been supported by CPER Nord-Pas de Calais/FEDER DATA Advanced data science and technologies 2015-2020, the French Ministry of Higher Education and Research, INRIA, and the French Agence Nationale de la Recherche (ANR).
@@
 %----------------------------------------------------------------------------------------
-\headerbox{References}{name=refs,column=0,span=1,below=ack, above=bottom}{
-    {
-    \AtNextBibliography{\footnotesize}
-    \setlength{\bibitemsep}{3pt}
-    \printbibliography[heading=none]
-    }
-}
+%\headerbox{References}{name=refs,column=0,span=1,below=ack, above=bottom}{
+%    {
+%    \AtNextBibliography{\footnotesize}
+%    \setlength{\bibitemsep}{3pt}
+%    \printbibliography[heading=none]
+%    }
+%}
 

### 2019-11-27 paper/main.tex
Updates from Overleaf

--- 2019-11-26 paper/main.tex
@@
 			\midrule 
-			Input sizes & [15, 7] & [11, 11, 7] & [1-15, 7] \\
+			Input sizes & [15, 7] & [32, 32, 7] & [1-15, 7] \\
 			Layers sizes & [128, 128] &  \makecell[tc]{Convolutional layers: 3 \\ Kernel Size: 2 \\

### 2020-07-28 paper/poster/poster.tex
Poster adjustements

@@
 
-\documentclass[a0paper,portrait,fontscale=0.3, margin=2cm]{baposter}
+\documentclass[paperwidth=36in,paperheight=48in,portrait,fontscale=0.3, margin=2cm]{baposter}
 
@@
 
-\bibliography{../budgeted_rl.bib}
+\bibliography{../references.bib}
 
@@
 textfont={\setlength{\parindent}{0.0em}\sffamily},
-headerheight={0.05\textheight},
+headerheight={0.06\textheight},
 eyecatcher=true
@@
 {
-\includegraphics[width=10em]{./img/companies}
+\includegraphics[height=1.7cm]{./img/companies}
 }
@@
 {
-Edouard Leurent, Jean Mercat
-\vspace{-4\baselineskip}
+Edouard Leurent$^*$, Jean Mercat$^*$
 }
 {
-\includegraphics[width=12em]{./img/inria_sc}
+\includegraphics[height=2cm]{./img/inria_sc}
 }
-
-\setlength{\colheight}{0.92\textheight}
 
@@
 
+Optimal Control Ojective
 \begin{equation*}
@@
 
-Bellman Optimality Equation:
+Bellman Optimality Equation and DQN Algorithm
 \begin{equation*}
@@
 	\begin{itemize}
-		\item we want the policy to be \hlg{permutation-invariant}:
+		\item[\incarrow] we want the policy to be \hlg{permutation-invariant}:
 	\end{itemize}
@@
 
-\begin{itemize}[nolistsep]
+\begin{itemize}[]
 	\item[\hlg{\checkmark}] Fixed-size
